### 2018-07-02 Manuscript/ManuscriptMarkdown.Rmd
Manuscript Update

---
title: "Working title: Regression models for Cylindrical data."
author: |
  | Jolien Cremers^[Corresponding author: j.cremers@uu.nl] $^1$, Helena J.M. Pennings$^{2,3}$,  Christophe Ley $^{4}$
  | $^1$Department of Methodology and Statistics, Utrecht University
  | $^2$TNO
  | $^3$Department of Education, Utrecht University
  | $^4$Department of Applied Mathematics, Computer Science and Statistics, Ghent University

bibliography: CircularData.bib
csl: apa.csl
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: yes
  html_document: default
  word_document: default
fontsize: 11pt
geometry: margin=1in
header-includes:
  - \usepackage{multirow}
  - \usepackage{appendix}
  - \usepackage{color}
  - \usepackage{hyperref}
  - \usepackage{subcaption}
keywords: bla vla
documentclass: article
pandoc_args: --natbib
abstract: |
  This is the abstract
---
```{r inlude = FALSE, echo = FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r, echo = FALSE}
library(circular)
library(bpnreg)
library(haven)
library(tikzDevice)
library(plotrix)
library(tidyr)
library(dplyr) #sample_n()
library(MASS) #mvrnorm()
library(kableExtra) #latex tables
```

\section{Introduction}\label{Introduction}

Cylindrical data are data that consist of a linear component and a directional 
component (a set of angles). In this paper, the directional component is 
circular meaning that it consists of a single angle instead of a set of angles. 
Figure \ref{} \textcolor{red}{NOG MAKEN} shows the difference between a circular variable (left) and a 
linear variable (right). The most obviuous difference is that on a circular 
scale the datapoints 0$^\circ$ and 360$^\circ$ are connected and in fact 
represent the same number while on a linear scale the two ends, $-\infty$ and 
$\infty$ are not connected. Cylindrical data occur in several fields of 
research, such as for instance in meteorology [@garcia2013exploring], ecology 
[@garcia2014test] or marine research [@lagona2015hidden]. Several types of data 
in psychological research are also of a cylindrical nature. For example, in 
research on human navigation in the field of cognitive psychology both distance,
a linear variable, and direction, a circular variable, of movement are of 
interest [@chrastil2017rotational]. In eye-tracking research, which is used for 
investigating cognitive processes we can also speak of cylindrical data. When 
measuring eye-movements we speak of the eye-movements themselves, the saccades, 
and fixations, the periods of time between movements when the eyes are still and
looking at one point. Of the saccades both the direction, a circular variable, 
and the duration, a linear variable, are of interest (for a review of 
eye-tracking research see @rayner200935th). Data from circumplex measurement
instruments, e.g. the interpersonal circumplex as used in personality
psychology, are also of a cylindrical nature (see Section \ref{Example} for a
more detailed explanation). In this paper we will discuss how a correct
statistical treatment of such cylindrical data can lead to new insights. In
particular we will show how cylindrical models pave the way for circular-linear
and linear-circular regression. We will do this for a motivating example from
the field of educational psychology. We will refer to this example data as the
teacher data. In this example, apart from modelling the relation between the
linear and circular component of a cylindrical variable we would also like to
predict the two components from a set of covariates in a regression model. The
teacher data will be further introduced in Sections \ref{Example} and
\ref{DataAnalysis}.

As is the case for circular data, the analysis of cylindrical data requires 
special methods. For an overview of methods for circular data see
@mardia2000directional and @jammalamadaka2001topics and for modern treatments
@ley2017modern. In the literature several methods have been put forward to
model the relation between the linear and circular component of a cylindrical 
variable. Some of these are based on regressing the linear component onto the 
circular component using the following type of relation: $$y = \beta_0 + 
\beta_1*\cos(\theta) + \beta_2*\sin(\theta)+ \epsilon,$$ where $y$ is the linear
component and $\theta$ the circular component [@mardia1978model;
@johnson1978some; @mastrantonio2015bayesian]. Others model the relation in a 
different way, e.g. by specifying a multivariate model for several linear and 
circular variables and modelling their covariance matrix 
[@mastrantonio2018joint] or by proposing a joint cylindrical distribution. 
For example, @abe2017tractable introduce a cylindrical distribution  based on a 
Weibull distribution for the linear component and a sine-skewed von Mises 
distribution for the circular component and link these through their shape and 
concentration parameters respectively. However, none of the methods that have 
been proposed thus far include additional covariates on which both the circular 
and linear component are regressed.

Our aim in this paper is to fill this gap in the literature by  adapting four of
the existing methods for the analysis of cylindrical data in such a way that
they include a regression of both the linear and circular component of a
cylindrical variable onto a set of covariates. First however, we will introduce
the teacher data in Section \ref{Example}. In Section \ref{Models} we introduce 
the four modified models for cylindrical data that we use to analyze the data from
the motivating example. We also choose a model selection criterion to compare
the models. The teacher data will be analysed in Section \ref{DataAnalysis}. The
paper will be concluded with a discussion in Section \ref{Discussion}. The
Appendix contains technical details.

\section{Teacher data}\label{Example}

The motivating example for this article comes from the field of educational 
psychology and was collected for the studies of @pennings2017interpersonal,
@Claessens2016side and @vanderWant2015role. These studies concern research on 
classroom climate. An indicator of classroom climate is the students'
perceptions of their teachers' interpersonal behavior. These interpersonal
perceptions, both in educational psychology as well as in other areas of
psychology, can be measured using circumplex measurement instruments (see
@horowitz2010handbook for an overview of many such instruments).

The Questionnaire on Teacher Interaction (QTI) [@wubbels2006interpersonal] is
one such circumplex measurement instrument that can be used to study student
perceptions of their teachers' interpersonal behavior. This instrument contains
items that load on two dimensions: Agency and Communion. Agency refers to the
degree of power or control a teacher exerts in interaction with his/her
students. Communion refers to the degree of friendliness or affiliation a
teacher conveys in interaction with his/her students. The loadings on the two 
dimensions of the QTI can be placed in a two-dimensional space formed by Agency
(vertical) and Communion (horizontal), see Figure \ref{QTI}. Different parts of
this space are characterized by different teacher behavior, e.g. 'helpful' or
'uncertain'. We call the two-dimensional space the interpersonal circle (IPC).
The idea is that the IPC is ``a continuous order with no beginning or end''
[@gurtman2009exploring, p. 2]. We call such ordering a circumplex ordering and
the IPC is therefore often called the interpersonal circumplex. The ordering
also implies that scores on the IPC could be viewed as a circular variable.

\begin{figure}
\centering
\includegraphics[width = 0.8\textwidth]{IPC-T.png}
\caption{The interpersonal circle for teachers (IPC-T). The words presented in
the circumference of the circle are anchor words to describe the type of
behavior located in each part of the IPC.}
\label{QTI}
\end{figure}

@Cremers2018Assessing show how data from the IPC can be
considered circular data and analyzed as such using a regression model. The 
two-dimension scores Agency and Communion can be converted to a circular score 
using the two-argument arctangent function in Equation \eqref{PredVal}, where 
$A$ represents a score on the Agency dimension and $C$ represents a score on the
Communion dimension. The resulting circular variable $\theta$ can then be
modeled and for example be regressed onto covariates in a circular regression 
model. However, when two-dimensional data are converted to the circle we lose 
some information, namely the length of the two-dimensional vector $(A, C)$, 
\emph{i.e.}, its euclidean norm $\mid\mid (A, C) \mid\mid$. This length 
represents the strength of the type of interpersonal behavior a teacher shows 
towards his/her students. In a cylindrical model we are able to incorporate this
information, and model a circular variable $\theta$ together with a linear 
variable corresponding to $\mid\mid (A, C) \mid\mid$. This will lead to an 
improved analysis of data from the IPC.  In the next section we will introduce 
several models that we can use for a more accurate and more informative regression
analysis on the teacher data. More information on the teacher data will be given
in Section \ref{DataAnalysis}.

\begin{equation}\label{PredVal}
\theta          = \text{atan2}\left(A, \: C\right)  =
\left|{\begin{array}{lcl}
                                                                       \arctan\left(\frac{A}{C}\right) & \text{if}  \quad&C > 0 \\
\arctan\left(\frac{A}{C}\right) + \pi & \text{if}  \quad& C  <  0  \:\: \&\:\: A \geq 0\\
 \arctan\left(\frac{A}{C}\right) - \pi & \text{if}  \quad&C  <  0 \:\:  \&\:\:A  < 0\\
 +\frac{\pi}{2} & \text{if}  \quad& C  =  0  \:\: \&\:\:A > 0\\
 -\frac{\pi}{2} & \text{if}  \quad& C =  0  \:\: \&\:\:A < 0\\
 \text{undefined} & \text{if} \quad& C =  0   \:\: \&\:\:A = 0.
 \end{array}}
\right.
\end{equation}

```{r loaddata, return = FALSE, echo = FALSE}
Dat <- read_spss("Data Heleen/mergeddata.sav")

#only select the first measurement occasion
Dat <- subset(Dat, Time == 0)

#sort on class size
Dat <- Dat[order(-Dat$nklas), ]

#Now remove duplicates 
#(because dataframe is sorted on class size the smallest classes are removed)

duplicates <- duplicated(Dat[,c("docnr", "Time")])
Dat <- Dat[!duplicates,]

#center and remove missings for self-efficacy
Dat <- Dat %>% drop_na(Efficacy_SE)
Dat$SEc <- Dat$Efficacy_SE- mean(Dat$Efficacy_SE)

set.seed(50)

#Create holdout and test sample
Dat.hold <- sample_n(Dat, 15)
Dat.test <- Dat[-as.numeric(rownames(Dat.hold)),]

#Compute circular and linear outcome
theta.test <- atan2(Dat.test$lAgency, Dat.test$lcommunion)
y.test     <- sqrt(Dat.test$lAgency^2 + Dat.test$lcommunion^2)

theta.hold <- atan2(Dat.hold$lAgency, Dat.hold$lcommunion)
y.hold     <- sqrt(Dat.hold$lAgency^2 + Dat.hold$lcommunion^2)

```

\section{Four cylindrical regression models}\label{Models}

In this section we will introduce four cylindrical regression models. We will 
extend the models from @mastrantonio2018joint and @abe2017tractable to contain 
predictors for the linear and circular outcomes in the cylindrical model. 
Additionally we introduce two models where both variates $\Theta$ and $Y$ are
predicted by covariates and that model the relation between the linear and
circular outcome as follows (following @wang2015joint and
@mastrantonio2015bayesian):

\begin{equation}\label{circlinlink}
y = \gamma_0 + \gamma_{cos}*\cos(\theta)*r + \gamma_{sin}*\sin(\theta)*r + \gamma_1*x_1 + \dots + \gamma_q*x_q +  \epsilon,
\end{equation}

where $r$ will be introduced in Section \ref{CL-(G)PN}, the error term $\epsilon
\sim N(0, \sigma)$, $\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, 
\gamma_q$ are the intercept and regression coefficients and   $x_1, \dots, x_q$
are the $q$ covariate values. In this model $Y$ follows a normal distribution
and $\Theta$ a projected normal (PN) or general projected normal (GPN)
distribution on the circle.

\subsection{The modified CL-PN and modified CL-GPN  models}\label{CL-(G)PN}

In both of these models the relation between the circular variable $\Theta \in
[0, 2\pi)$ and linear variable $Y \in (-\infty, + \infty)$ is specified as in
Equation \ref{circlinlink}. The distribution of $Y$ conditional on $\Theta$ is
then given as:

\begin{equation}\label{ycondtheta}
f(y \mid \theta) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp\left[\frac{c^2 + (y - (\gamma_0 + \gamma_1x_1 + \dots + \gamma_qx_q))^{2} - 2c(y - (\gamma_0 + \gamma_1x_1 + \dots + \gamma_qx_q))}{2\sigma^2}\right],
\end{equation}

where $c = \begin{bmatrix} r \cos \theta \\  r\sin \theta \end{bmatrix}^t 
\begin{bmatrix} \gamma_{cos} \\ \gamma_{sin} \end{bmatrix}$, $r \geq 0$, 
$\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q$ are the
intercept and regression coefficients and $\sigma^2 \geq 0$ is the error
variance. The linear outcome thus has a normal distribution, conditional on
$\Theta$.

For the circular outcome we assume either a projected normal (PN) or a
general projected normal (GPN) distribution. These distributions arise from the
radial projection of a distribution defined on the plane onto the circle. The
relation between a bivariate variate $\boldsymbol{S}$ in the plane and the
circular outcome $\Theta$ is defined as follows:

\begin{equation}\label{projection}
\boldsymbol{S} = \begin{bmatrix} S^{I} \\ S^{II} \end{bmatrix} = R\boldsymbol{u} = \begin{bmatrix} R \cos \Theta \\  R\sin \Theta \end{bmatrix},
\end{equation}

where $R = \mid\mid S \mid\mid$, the euclidean norm of the bivariate vector 
$\boldsymbol{S}$. In the PN distribution we assume $\boldsymbol{S} \sim 
N_2(\boldsymbol{\mu}, \boldsymbol{I})$ and in the GPN we assume $\boldsymbol{S} 
\sim N_2(\boldsymbol{\mu}, \boldsymbol{\Sigma})$ where $\boldsymbol{\Sigma} = 
\begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1 \end{bmatrix}$, $\rho \in 
(-\infty, +\infty)$ and $\tau^2 \geq 0$ (as in @hernandez2016general).

Following @nunez2011bayesian, the joint density of $\Theta$ and $R$ for the PN
distribution in a regression set-up equals:

\begin{equation}\label{pnreg}
f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = [2\pi]^{-1} \exp\left[ \frac{-r^2 - \boldsymbol{\mu}^2\boldsymbol{\mu} + 2r\boldsymbol{u}^t\boldsymbol{\mu}}{2}\right],
\end{equation}

In a regression setup the outcome $\theta_i$ for each individual $i = 1,
\dots, n$, where $n$ is the sample size, is generated independently from
Equation (\ref{pnreg}). The mean vector $\boldsymbol{\mu}_i \in \mathbb{R}^2$ is 
defined as $\boldsymbol{\mu}_i = \boldsymbol{z}_i\boldsymbol{B}$. The vector 
$\boldsymbol{z}_i$ is a a vector of $p$ covariate values and $\boldsymbol{B} = 
(\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$ contain the regression 
coefficients. Note however that the dimensions of $\boldsymbol{\beta}^{I}$ and 
$\boldsymbol{\beta }^{I}$ need not necessarily be the same and we are thus 
allowed to have a different set of predictor variables and vectors 
$\boldsymbol{z}_i^I$ and $\boldsymbol{z}_i^{II}$ for the two components of 
$\boldsymbol{\mu}_i$.

Following @wang2012directional and @hernandez2016general the joint density of
$r$ and $\theta$ for the GPN distribution equals:

\begin{equation}\label{gpnreg}
f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma}) = r(2\pi\tau)^{-1} \exp\left[ -0.5 \sigma^2(r\boldsymbol{u}-\boldsymbol{\mu})^{t}\boldsymbol{\Sigma}^{-1}(r\boldsymbol{u}-\boldsymbol{\mu})\right],
\end{equation}

where $\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2  & \rho\\ \rho & 1 
\end{bmatrix}$, $\boldsymbol{u}= \begin{bmatrix} \cos \theta \\ \sin \theta 
\end{bmatrix}$. In a regression setup the outcome $\theta_i$ for each individual
$i = 1, \dots, n$, where $n$ is the sample size, is generated independently
from Equation (\ref{gpnreg}). The mean vector $\boldsymbol{\mu}_i \in 
\mathbb{R}^2$ is defined as $\boldsymbol{\mu}_i = 
\boldsymbol{z}_i(\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$. The vector 
$\boldsymbol{z}_i$ is a a vector of $p$ covariate values and each 
$\boldsymbol{\beta}^{k}$ is a vector with two regression coefficients, one for 
each of the two components of $\boldsymbol{\mu}_i$. Note that for the CL-GPN 
model we do need to have the same predictors for both components of 
$\boldsymbol{\mu}_i$.

Both cylindrical models introduced in this section are estimated using MCMC 
methods. These methods are based on @nunez2011bayesian, @wang2012directional and
@hernandez2016general for the regression of the circular outcome. A detailed
description of the Bayesian estimation and MCMC samplers can be found in
Appendices \ref{A1} and \ref{A2}.

\subsection{The modified Abe-Ley model}\label{WeiSSVM}

This model is an extension of the cylindrical model introduced by
@abe2017tractable to the regression context. The density of the model for the
pair of the circular variable $\theta \in [0, 2\pi)$ and linear variable $y$ on
the positive real half-line $[0, + \infty)$ is:

\begin{equation}\label{WeiSSVMdensity}
f(\theta, y) = \frac{\alpha(\beta)^\alpha}{2\pi\cosh(\kappa)}
                 (1 +\lambda\sin(\theta - \mu))
                 y^{\alpha-1}
                 \exp[-((\beta y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu)))],
\end{equation}

In a regression setup the outcome vector $(\theta_i, y_i)^t$ for each individual
$i = 1, \dots, n$, where $n$ is the sample size, is generated independently from
Equation (\ref{WeiSSVMdensity}). The parameter $\alpha > 0$ is a linear shape 
parameter, $\beta_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\nu}) > 0$ is a linear 
scale parameter, $\mu_i = \eta_0 + 
2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}) \in [0, 2\pi)$ is a circular 
location parameter, $\kappa > 0$ is a cicular concentration parameter and 
$\lambda \in [-1, 1]$ is a circular skewness parameter. The parameter 
$\boldsymbol{\nu}$ is a vector of $q$ regression coefficients and intercept
$\nu_j \in (-\infty, +\infty)$ where $j = 0, \dots, q$, for the prediction of
$y$. The parameter $\eta_0 \in [0, 2\pi)$ is the intercept and
$\boldsymbol{\eta}$, is a vector of $p$ regression coefficients $\eta_j \in 
(-\infty, +\infty)$, where $j = 1, \dots, p$, for the prediction of $\theta$. 
The vector $\boldsymbol{x}_i$ is a vector of predictor values for the prediction
of $y$ and $\boldsymbol{z}_i$ is a vector of predictor values for the prediction
of $\theta$.

As in @abe2017tractable, the conditional distribution of $y$ given $\theta$ is a
Weibull distribution with scale $\alpha$ and shape
$\beta(1-\tanh(\kappa)\cos(\theta - \mu))^{1-\alpha}$ and the conditional 
distribution of $\theta$ given $y$ is a sine skewed von Mises distribution with 
location parameter $\mu$ and concentration parameter $(\beta 
y)^\alpha\tanh(\kappa)$.

The log-likelihood for this model equals:

\begin{align}\label{WeiSSVMLikelihood}
l(\alpha, \boldsymbol{\nu}, \lambda, \kappa, \boldsymbol{\eta}) 
   &= n[\ln(\alpha) - \ln(2\pi\cosh(\kappa))] + \alpha \sum^{n}_{i = 1} \ln(\exp(\boldsymbol{x}_i^t\boldsymbol{\nu})) \nonumber\\
   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - \eta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}))) 
   +(\alpha-1)\sum^{n}_{i = 1} \ln(y_i) \nonumber\\
   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\nu})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - \eta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta})))
\end{align}

We can use numerical optimization (Nelder-Mead) to find solutions for the
maximum likelihood (ML) estimates for the parameters of the model.

\subsection{Modified joint projected and skew normal (GPN-SSN)}\label{CL-GPN_multivariate}

This model is an extension of the cylindrical model introduced by
@mastrantonio2018joint to the regression context. The model contains $p$
circular outcomes and $q$ linear outcomes. The circular outcomes 
$\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots, \boldsymbol{\Theta}_p)$ 
are modeled together by a multivariate GPN distribution and the linear outcomes 
$\boldsymbol{Y} = (\boldsymbol{Y}_1, \dots, \boldsymbol{Y}_q)$  are modeled 
together by a multivariate skew normal distribution [@sahu2003new]. Because the
GPN distribution is modelled using a so-called augmented representation (as in
(\ref{projection}) and (\ref{gpnreg})) it is convenient to use a similar tactic
for modelling the multivariate skew normal distribution. As in
@mastrantonio2018joint we use the following representation of the linear
outcomes:

$$\boldsymbol{Y} = \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D} + \boldsymbol{H},$$
where where $\boldsymbol{\mu}_y$ is a mean vector for the linear outcome 
$\boldsymbol{Y}$, $\boldsymbol{\Lambda} = \text{diag}(\boldsymbol{\lambda})$ is a $q 
\times q$ diagonal matrix with diagonal elements $\boldsymbol{\lambda} = 
\lambda_1, \dots, \lambda_q$ (the parameters that indicate the skewness of the 
linear outcomes), $\boldsymbol{D} \sim HN_q(\boldsymbol{0}_q, 
\boldsymbol{I}_q)$, a q-dimensional half normal distribution
[@olmos2012extension] and $\boldsymbol{H} \sim N_q(\boldsymbol{0}_q,
\boldsymbol{\Sigma}_y)$. This means that conditional on the auxiliary data
$\boldsymbol{D}$, $\boldsymbol{Y}$ is normally distributed with mean
$\boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D}$ and covariance matrix
$\boldsymbol{\Sigma}_y$. The joint density for $(\boldsymbol{Y},
\boldsymbol{D})^t$ is defined as follows:

\begin{equation}\label{YDjoint}
f(\boldsymbol{y}, \boldsymbol{d}) = 2^q\phi_q(\boldsymbol{y} \mid \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{d}, \boldsymbol{\Sigma}_y) \phi_1(\boldsymbol{d} \mid \boldsymbol{0}_q, \boldsymbol{I}_q).
\end{equation}

As in @mastrantonio2018joint dependence between the linear and circular outcome
is created by modelling the augmented representations of $\boldsymbol{\Theta}$,
and $\boldsymbol{Y}$ together in a $2 \times p + q$ dimensional normal
distribution. The joint density of the model is then represented by:

\begin{equation}\label{YDThetarjoint}
f(\boldsymbol{\theta}, \boldsymbol{r}, \boldsymbol{y}, \boldsymbol{d}) = 2^q\phi_{2p+q}((\boldsymbol{s}, \boldsymbol{y})^t \mid \boldsymbol{\mu} + (\boldsymbol{0}_{2p}, diag(\boldsymbol{\lambda})\boldsymbol{d})^t, \boldsymbol{\Sigma}) \phi_q(\boldsymbol{d} \mid \boldsymbol{0}_q, \boldsymbol{I}_q) \prod_{j = 1}^{p}r_j,
\end{equation}

where the mean vector $\boldsymbol{\mu} = (\boldsymbol{\mu}_s,
\boldsymbol{\mu}_y)^t$ and $\boldsymbol{\Sigma} = \left ( \begin{matrix}
\boldsymbol{\Sigma}_s & \boldsymbol{\Sigma}_{sy} \\ \boldsymbol{\Sigma}_{sy^t} &
\boldsymbol{\Sigma}_y \\ \end{matrix} \right )$. The matrix
$\boldsymbol{\Sigma}_s$ is the covariance matrix for the variances of and
covariances between the augmented representations of the circular outcome and
the matrix $\boldsymbol{\Sigma}_{sy}$ contains covariances between the augmented
representations of the circular outcome and the linear outcome.

In our regression extension we have $i = 1, \dots, n$ observations of $p$ 
circular outcomes, $q$ linear outcomes and $k$ covariates. The mean in the 
density in (\ref{YDThetarjoint}) then becomes $\boldsymbol{\mu}_i = 
\boldsymbol{x}_i\boldsymbol{B}$ where $\boldsymbol{B}$ is a $k \times (2 
\times p + q)$ matrix with regression coeffients. We estimate the model using 
MCMC methods. A detailed description of these methods is given in Appendix
\ref{A3}.

\subsection{Model fit criterion}\label{Modelfit}

For the four cylindrical models we focus on their out-of-sample predictive 
performance to determine the fit of the model. To do so we split our data in a
training and holdout set (10 $\%$ of the sample). A proper criterion to compare
out-of-sample predictive performance is the the Predictive Log Scoring Loss
(PLSL) [@gneiting2007strictly]. The lower the value of this criterion, the
better the predictive performance of the model. Using ML estimates this
criterion can be computed as follows:

\begin{equation}\label{PLSLML}
PLSL = -2 \sum_{i = 1}^{M}\log l(x_i \mid \hat{\boldsymbol{\vartheta}}),
\end{equation}

where $l$ is the model likelihood, $M$ is the sample size of the holdout set, 
$x_i$ is the $i^{th}$ datapoint from the holdout set and 
$\hat{\boldsymbol{\vartheta}}$ are the ML estimates of the model parameters. 
Using posterior samples the criterion is similar to the log pointwise predictive
density (lppd) as outlined in @BDA and can be computed as:

\begin{equation}\label{PLSLBayes}
PLSL = -2 \frac{1}{B} \sum_{j = 1}^{B}\sum_{i = 1}^{M} \log l(x_i \mid \boldsymbol{\vartheta}^{(j)}),
\end{equation}

where $B$ is the amount of posterior samples and $\boldsymbol{\vartheta}^{(j)}$
are the posterior estimates of the model parameters for the $j^{th}$ iteration.
Note that although we fit the CL-PN, CL-GPN and joint GPN-SSN models using
Bayesian statistics, we do not take prior information (only the likelihood) into
account when assessing model fit. According to @BDA this is not necessary since
we are assessing the fit of a model to data only \textcolor{red}{AFMAKEN!}. For 
each of the cylindrical models we can then compute a PLSL for the circular and 
linear outcome by using the conditional log-likelihoods of the respective 
outcome.

\section{Data Analysis}\label{DataAnalysis}

In this section we will analyze the teacher dataset with the help of
the four cylindrical models introduced in Section \ref{Models}. We will however
first give a more detailed description of the dataset.

```{r FitCLPN, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}

source("Posterior Sampling informative Prior CL-PN.R")

its <- 20000
set.seed(101)

ZI.PN.test  <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
ZII.PN.test <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
X.PN.test   <- as.matrix(cbind(rep(1, length(theta.test)),
                               cos(theta.test),
                               sin(theta.test),
                               Dat.test$SEc))

ZI.PN.hold  <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
ZII.PN.hold <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
X.PN.hold   <- as.matrix(cbind(rep(1, length(theta.hold)),
                               cos(theta.hold),
                               sin(theta.hold),
                               Dat.hold$SEc))

res_CLPN <- CLPN(theta.test, y.test, X.PN.test, ZI.PN.test, ZII.PN.test, its,
                 theta.hold, y.hold, X.PN.hold, ZI.PN.hold, ZII.PN.hold)
```

```{r FitCLGPN, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}

source("Posterior Sampling CL-GPN.R")

its <- 20000
set.seed(101)

Z.GPN.test <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
X.GPN.test <- as.matrix(cbind(rep(1, length(theta.test)),
                              cos(theta.test),
                              sin(theta.test),
                              Dat.test$SEc))

Z.GPN.hold <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
X.GPN.hold <- as.matrix(cbind(rep(1, length(theta.hold)),
                              cos(theta.hold),
                              sin(theta.hold),
                              Dat.hold$SEc))

res_CLGPN <- CLGPN(theta.test, y.test, X.GPN.test, Z.GPN.test, its, p = 2, 
                   theta.hold, y.hold, X.GPN.hold, Z.GPN.hold)

```

```{r FitCLGPNM, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}

source("Posterior Sampling CL-GPN_multivariate.R")

its <- 20000
set.seed(101)

X.MGPN.test <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))

X.MGPN.hold <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))

res_CLGPNM <- CLGPNM(theta.test, y.test, X.MGPN.test, its, p = 1, q = 1, 
                     theta.hold, y.hold, X.MGPN.hold)

```

```{r FitAbeLey, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}

source("WeisSVM optimization.R")

set.seed(101)

Z.test     <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))
X.test    <- as.matrix(cbind(rep(1, length(theta.test)), Dat.test$SEc))

Z.hold     <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))
X.hold     <- as.matrix(cbind(rep(1, length(theta.hold)), Dat.hold$SEc))

colnames(X.test) <- c("ax", "bx")
colnames(Z.test) <- c("az", "bz")
colnames(X.hold) <- c("ax", "bx")
colnames(Z.hold) <- c("az", "bz")

dat.t <- as.data.frame(cbind(theta.test, y.test, X.test, Z.test))
dat.h <- as.data.frame(cbind(theta.hold, y.hold, X.hold, Z.hold))

ui.reg = rbind(c(1, 0, 0, 0, 0, 0, 0),
               c(-1, 0, 0, 0, 0, 0, 0),
               c(0, 0, 0, 0, 1, 0, 0),
               c(0, 0, 0, 0, 0, 1, 0),
               c(0, 0, 0, 0, 0, 0, 1),
               c(0, 0, 0, 0, 0, 0, -1))

ci.reg = c(-pi, -pi, 0, 0, -1, -1)

param <- c(1,1,1,1,1,1,0)*0.5

ui.reg %*% param - ci.reg

resAL <- constrOptim(param, func.regII, ui = ui.reg, ci = ci.reg, method = "Nelder-Mead",
                     control = list(maxit = 1000000, fnscale = -1), data = dat.t)

```

\subsection{Data description}\label{DataDescriptives}

The teacher data was collected between 2010 and 2015 and contains several repeated
measures on the IPC of 161 teachers gathered for the studies of
@vanderWant2015role, @Claessens2016side and @pennings2017interpersonal. 
The measurements were obtained using the QTI and taken in different years and in
different classes. For this paper we only consider one measurement, namely the
one for the first measurement occasion (2010) and the one for the largest class
if data for multiple classes were available. In addition to a variable `IPC`
containing the circular outcome and the 'length' of the IPC, the linear outcome,
a teachers' self-efficacy (\verb|SE|) will be used as covariate in the analysis.
After the removal of missing cases we end up with a sample of 148 teachers Table
\ref{Tableteacherdescriptives} shows descriptives for the dataset.

```{r descripitves, echo = FALSE, results = FALSE}
#sample size
nrow(Dat)

#summary statistics
summary(Dat$Efficacy_SE)
summary(sqrt(Dat$lAgency^2 + Dat$lcommunion^2))
summary(as.circular(atan2(Dat$lAgency, Dat$lcommunion)%%(2*pi)))

sd(Dat$Efficacy_SE)
sd(sqrt(Dat$lAgency^2 + Dat$lcommunion^2))

```

\begin{table}
\centering
\caption{Descriptives for the teacher dataset} 
\begin{tabular}{lrrrl}
  \noalign{\smallskip}\hline\noalign{\smallskip}
Variable & mean/$\bar{\theta}$ & sd/$\hat{\rho}$ & Range & Type \\ \hline\noalign{\smallskip}
IPC &33.23$^\circ$& 0.76 & - & Circular\\
length IPC & 0.43 & 0.15 & 0.08 - 0.80 & Linear\\
SE & 4.72 & 0.87 & 2 - 6.5 & Linear\\
   \hline
\end{tabular}
\label{Tableteacherdescriptives}
\caption*{Note that $\hat{\rho}$ is an sample estimate for the circular concentration where a value of 0 means that the data is not concentrated at all, i.e. spread over the entire circle, and a value of 1 means that all data is concentrated at a single point on the circle. }
\end{table}

\subsection{Models}\label{DataModels}

The regression equations for the linear and the circular outcome of the four
cylindrical models fit to the teacher dataset are as follows:

\begin{itemize}
\item For the modified CL-PN and CL-GPN models:

$\hat{\boldsymbol{\mu}}_{i} = \begin{pmatrix}
  \mu_{i}^{I}  \\
\mu_{i}^{II}
 \end{pmatrix}=\begin{pmatrix}
  \beta_0^{I} + \beta_1^{I}\text{SE}_i  \\
  \beta_0^{II} + \beta_1^{II}\text{SE}_i
 \end{pmatrix},$

$\hat{y}_i = \gamma_0 + \gamma_{cos}\cos\theta_ir_i + \gamma_{sin}\sin\theta_ir_i + \gamma_1\text{SE}_i.$

\item For the modified Abe-Ley model:

$\hat{\mu}_{i} = \eta_0 + 2 * \tan^{-1}(\eta_1\text{SE}_i),$

$\hat{\beta}_{i} = \exp(\nu_0 + \nu_1\text{SE}_i).$

\item For the modified joint projected and skew normal model:

$\hat{\boldsymbol{\mu}}_{i} = \boldsymbol{\beta}_0 + \boldsymbol{\beta}_1\text{SE}_i$, where $\boldsymbol{\mu}_i = (\boldsymbol{\mu}_{s_i}, \boldsymbol{\mu}_{y_i})^t$, $\boldsymbol{\beta}_0 = (\beta_{0_s^{I}}, \beta_{0_s^{II}},\beta_{0_y})$ and $\boldsymbol{\beta}_1 = (\beta_{1_s^{I}}, \beta_{1_s^{II}},\beta_{1_y})$.
\end{itemize}

The variable \verb|SE| is thus used as a predictor for the linear and circular 
outcomes in all four cylindrical models. We will use the loglikelihoods of the
following conditional densities \textcolor{red}{CHRISTOPHE, MAG HIER WEL $\sim$ GEBRUIKT WORDEN?}for the computation of the PLSL criterion to
evaluate the model fit:

\begin{itemize}
\item for the modified CL-PN model:

$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use Equation (\ref{pnreg}).

\item for the modified CL-GPN model:

$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use Equation (\ref{gpnreg}).

\item for the modified Abe-Ley model:

$y_i \mid \boldsymbol{\theta}_i, \beta_i, \mu_i, \kappa, \alpha \sim W\left(\beta_i(1-\tanh(\kappa)\cos(\theta_i - \mu_i))^{1/\alpha}, \alpha\right)$, a Weibull distribution.

$\theta_i \mid y_i, \beta_i, \mu_i, \kappa, \alpha \lambda \sim SSVM\left(\mu_i, (\beta_iy_i)^{\alpha}(\tanh{\kappa})\right)$, a sine-skewed von Mises distribution.

\item for the modified joint projected and skew normal model:

$y_i \mid \boldsymbol{\mu}_i, \boldsymbol{\Sigma}, \boldsymbol{\theta}_i, r_i \sim SSN(\mu_{i_y} + \lambda d_i + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_w^{-1}(\boldsymbol{s}_i - \boldsymbol{\mu}_{i_s}), \boldsymbol{\Sigma}_y + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_s\boldsymbol{\Sigma}_{sy}),$

$\theta_i \mid \boldsymbol{\mu}_i, \boldsymbol{\Sigma}, y_i, d_i \sim GPN(\boldsymbol{\mu}_{i_s} + \boldsymbol{\Sigma}_{sy}\boldsymbol{\Sigma}_y^{-1} (y_i - \mu_{i_y} - \lambda d_i), \boldsymbol{\Sigma}_s + \boldsymbol{\Sigma}_{sy}\boldsymbol{\Sigma}_y^{-1}\boldsymbol{\Sigma}_{sy}^t)$

where $SSN$ is the skew normal distribution.

\end{itemize}

\subsection{Results \& Analysis}\label{DataResults}

Before analysis there were a couple of settings that we had to specify for the 
cylindrical models. Starting values for the WeiSSVM model were the following 
$\eta_0 = 0.5, \eta_1 = 0.5, \nu_0 = 0.5, \nu_1 = 0.5, \kappa = 0.5, \alpha = 
0.5, \lambda = 0$. The initial amount of iterations for the three MCMC samplers 
was set to 2000. After we checked convergence via traceplots we concluded that 
some of the parameters of the multivariate GPN model did not converge. Therefore
we set the amount of iterations of the MCMC models to 20,000 and subtracted a 
burn-in of 5000. With this specification the MCMC chains for all parameters 
converge. Note that we choose the same amount of iterations for all three 
Bayesian cylindrical models to make their comparison via the PLSL as fair as 
possible. Lastly, the predictor \verb|SE| was centered before inclusion in the
analysis.

```{r estCLPN, echo = FALSE}

CLPN.Gamma.m   <- apply(res_CLPN$Gamma[5000:20000,], 2, mode_est)
CLPN.Gamma.hpd <- apply(res_CLPN$Gamma[5000:20000,], 2, hpd_est)

CLPN.BI.m   <- apply(res_CLPN$BI[5000:20000,], 2, mode_est)
CLPN.BI.hpd <- apply(res_CLPN$BI[5000:20000,], 2, hpd_est)

CLPN.BII.m <- apply(res_CLPN$BII[5000:20000,], 2, mode_est)
CLPN.BII.hpd <- apply(res_CLPN$BII[5000:20000,], 2, hpd_est)

CLPN.Sigma.m   <- mode_est(res_CLPN$Sigma[5000:20000])
CLPN.Sigma.hpd <- hpd_est(res_CLPN$Sigma[5000:20000])

modes.PN <- c(CLPN.BI.m, CLPN.BII.m, CLPN.Gamma.m, CLPN.Sigma.m, NA, NA, NA)
hpd.LB.PN <- c(CLPN.BI.hpd[1,], CLPN.BII.hpd[1,], CLPN.Gamma.hpd[1,], CLPN.Sigma.hpd[1], NA, NA, NA)
hpd.UB.PN <- c(CLPN.BI.hpd[2,], CLPN.BII.hpd[2,], CLPN.Gamma.hpd[2,], CLPN.Sigma.hpd[2], NA, NA, NA)

CLPNres.tab <- cbind(modes.PN, hpd.LB.PN, hpd.UB.PN)

```

```{r estCLGPN, echo = FALSE}

CLGPN.Sig.m   <- mode_est(res_CLGPN$Sig)
CLGPN.Sig.hpd <- hpd_est(res_CLGPN$Sig)

CLGPN.Gamma.m   <- apply(res_CLGPN$Gamma[5000:20000,], 2, mode_est)
CLGPN.Gamma.hpd <- apply(res_CLGPN$Gamma[5000:20000,], 2, hpd_est)

CLGPN.BI.m   <- apply(res_CLGPN$BI[5000:20000,], 2, mode_est)
CLGPN.BI.hpd <- apply(res_CLGPN$BI[5000:20000,], 2, hpd_est)

CLGPN.BII.m   <- apply(res_CLGPN$BII[5000:20000,], 2, mode_est)
CLGPN.BII.hpd <- apply(res_CLGPN$BII[5000:20000,], 2, hpd_est)

CLGPN.Sigma11.m   <- mode_est(res_CLGPN$Sigma[1,1,5000:20000])
CLGPN.Sigma11.hpd <- hpd_est(res_CLGPN$Sigma[1,1,5000:20000])
CLGPN.Sigma22.m   <- mode_est(res_CLGPN$Sigma[2,2,5000:20000])
CLGPN.Sigma22.hpd <- hpd_est(res_CLGPN$Sigma[2,2,5000:20000])
CLGPN.Sigma12.m   <- mode_est(res_CLGPN$Sigma[1,2,5000:20000])
CLGPN.Sigma12.hpd <- hpd_est(res_CLGPN$Sigma[1,2,5000:20000])

modes.GPN  <- c(CLGPN.BI.m, CLGPN.BII.m,
                CLGPN.Gamma.m, CLGPN.Sig.m,
                CLGPN.Sigma11.m, CLGPN.Sigma12.m, CLGPN.Sigma22.m)
hpd.LB.GPN <- c(CLGPN.BI.hpd[1,], CLGPN.BII.hpd[1,],
               CLGPN.Gamma.hpd[1,], CLGPN.Sig.hpd[1],
               CLGPN.Sigma11.hpd[1], CLGPN.Sigma12.hpd[1], CLGPN.Sigma22.hpd[1])
hpd.UB.GPN <- c(CLGPN.BI.hpd[2,], CLGPN.BII.hpd[2,], 
               CLGPN.Gamma.hpd[2,], CLGPN.Sig.hpd[2],
               CLGPN.Sigma11.hpd[2], CLGPN.Sigma12.hpd[2], CLGPN.Sigma22.hpd[2])

CLGPNres.tab <- cbind(modes.GPN, hpd.LB.GPN, hpd.UB.GPN)
CLPNGPNres.tab <- round(cbind(CLPNres.tab, CLGPNres.tab), 2)

rownames(CLPNGPNres.tab) <- c("$\\beta_0^{I}$", "$\\beta_0^{I}$",
                              "$\\beta_0^{II}$", "$\\beta_1^{II}$",
                              "$\\gamma_0$", "$\\gamma_{cos}$",
                              "$\\gamma_{sin}$", "$\\gamma_1$",
                              "$\\sigma$", "$\\sum_{1,1}$",
                              "$\\sum_{1,2}$", "$\\sum_{2,2}$")
colnames(CLPNGPNres.tab) <- c("Mode", "HPD LB", "HPD UB", "Mode", "HPD LB", "HPD UB")

kable(CLPNGPNres.tab, "latex", booktabs = TRUE, escape = FALSE,
      caption = "Results for the modified CL-PN and CL-GPN model")%>%
add_header_above(c("Parameter" = 1, "CL-PN" = 3, "CL-GPN" = 3))

```

```{r estSL, echo = FALSE}

ALres.tab <- matrix(round(c(resAL$par[1:4], resAL$par[6], resAL$par[5], resAL$par[7]), 2),7,1)
rownames(ALres.tab) <- c("$\\eta_0$", "$\\eta_1$",
                         "$\\nu_0$", "$\\nu_1$",
                         "$\\alpha$", "$\\kappa$", "$\\lambda$")

kable(ALres.tab, "latex", booktabs = T, escape = FALSE,
      caption = "Results for the modified Abe-Ley model")%>%
add_header_above(c("Parameter" = 1, "ML-estimate" = 1))

```

```{r estCLMGPN, echo = FALSE}

CLGPNM.lambda.m   <- mode_est(res_CLGPNM$lambda[5000:20000, ])
CLGPNM.lambda.hpd <- hpd_est(res_CLGPNM$lambda[5000:20000, ])

CLGPNM.Beta11.m <- mode_est(res_CLGPNM$Beta[1,1,5000:20000])
CLGPNM.Beta21.m <- mode_est(res_CLGPNM$Beta[2,1,5000:20000])
CLGPNM.Beta12.m <- mode_est(res_CLGPNM$Beta[1,2,5000:20000])
CLGPNM.Beta22.m <- mode_est(res_CLGPNM$Beta[2,2,5000:20000])
CLGPNM.Beta13.m <- mode_est(res_CLGPNM$Beta[1,3,5000:20000])
CLGPNM.Beta23.m <- mode_est(res_CLGPNM$Beta[2,3,5000:20000])

CLGPNM.Beta11.hpd <- hpd_est(res_CLGPNM$Beta[1,1,5000:20000])
CLGPNM.Beta21.hpd <- hpd_est(res_CLGPNM$Beta[2,1,5000:20000])
CLGPNM.Beta12.hpd <- hpd_est(res_CLGPNM$Beta[1,2,5000:20000])
CLGPNM.Beta22.hpd <- hpd_est(res_CLGPNM$Beta[2,2,5000:20000])
CLGPNM.Beta13.hpd <- hpd_est(res_CLGPNM$Beta[1,3,5000:20000])
CLGPNM.Beta23.hpd <- hpd_est(res_CLGPNM$Beta[2,3,5000:20000])

CLGPNM.Betacon11.m <- mode_est(res_CLGPNM$Betacon[1,1,5000:20000])
CLGPNM.Betacon21.m <- mode_est(res_CLGPNM$Betacon[2,1,5000:20000])
CLGPNM.Betacon12.m <- mode_est(res_CLGPNM$Betacon[1,2,5000:20000])
CLGPNM.Betacon22.m <- mode_est(res_CLGPNM$Betacon[2,2,5000:20000])
CLGPNM.Betacon13.m <- mode_est(res_CLGPNM$Betacon[1,3,5000:20000])
CLGPNM.Betacon23.m <- mode_est(res_CLGPNM$Betacon[2,3,5000:20000])

CLGPNM.Betacon11.hpd <- hpd_est(res_CLGPNM$Betacon[1,1,5000:20000])
CLGPNM.Betacon21.hpd <- hpd_est(res_CLGPNM$Betacon[2,1,5000:20000])
CLGPNM.Betacon12.hpd <- hpd_est(res_CLGPNM$Betacon[1,2,5000:20000])
CLGPNM.Betacon22.hpd <- hpd_est(res_CLGPNM$Betacon[2,2,5000:20000])
CLGPNM.Betacon13.hpd <- hpd_est(res_CLGPNM$Betacon[1,3,5000:20000])
CLGPNM.Betacon23.hpd <- hpd_est(res_CLGPNM$Betacon[2,3,5000:20000])

CLGPNM.Sigma11.m <- mode_est(res_CLGPNM$Sigma[1,1,5000:20000])
CLGPNM.Sigma12.m <- mode_est(res_CLGPNM$Sigma[1,2,5000:20000])
CLGPNM.Sigma13.m <- mode_est(res_CLGPNM$Sigma[1,3,5000:20000])
CLGPNM.Sigma22.m <- mode_est(res_CLGPNM$Sigma[2,2,5000:20000])
CLGPNM.Sigma23.m <- mode_est(res_CLGPNM$Sigma[2,3,5000:20000])
CLGPNM.Sigma33.m <- mode_est(res_CLGPNM$Sigma[3,3,5000:20000])

CLGPNM.Sigma11.hpd <- hpd_est(res_CLGPNM$Sigma[1,1,5000:20000])
CLGPNM.Sigma12.hpd <- hpd_est(res_CLGPNM$Sigma[1,2,5000:20000])
CLGPNM.Sigma13.hpd <- hpd_est(res_CLGPNM$Sigma[1,3,5000:20000])
CLGPNM.Sigma22.hpd <- hpd_est(res_CLGPNM$Sigma[2,2,5000:20000])
CLGPNM.Sigma23.hpd <- hpd_est(res_CLGPNM$Sigma[2,3,5000:20000])
CLGPNM.Sigma33.hpd <- hpd_est(res_CLGPNM$Sigma[3,3,5000:20000])

CLGPNM.Sigmacon11.m <- mode_est(res_CLGPNM$Sigmacon[1,1,5000:20000])
CLGPNM.Sigmacon12.m <- mode_est(res_CLGPNM$Sigmacon[1,2,5000:20000])
CLGPNM.Sigmacon13.m <- mode_est(res_CLGPNM$Sigmacon[1,3,5000:20000])
CLGPNM.Sigmacon22.m <- mode_est(res_CLGPNM$Sigmacon[2,2,5000:20000])
CLGPNM.Sigmacon23.m <- mode_est(res_CLGPNM$Sigmacon[2,3,5000:20000])
CLGPNM.Sigmacon33.m <- mode_est(res_CLGPNM$Sigmacon[3,3,5000:20000])

CLGPNM.Sigmacon11.hpd <- hpd_est(res_CLGPNM$Sigmacon[1,1,5000:20000])
CLGPNM.Sigmacon12.hpd <- hpd_est(res_CLGPNM$Sigmacon[1,2,5000:20000])
CLGPNM.Sigmacon13.hpd <- hpd_est(res_CLGPNM$Sigmacon[1,3,5000:20000])
CLGPNM.Sigmacon22.hpd <- hpd_est(res_CLGPNM$Sigmacon[2,2,5000:20000])
CLGPNM.Sigmacon23.hpd <- hpd_est(res_CLGPNM$Sigmacon[2,3,5000:20000])
CLGPNM.Sigmacon33.hpd <- hpd_est(res_CLGPNM$Sigmacon[3,3,5000:20000])

modes.GPN  <- c(CLGPNM.Beta11.m, CLGPNM.Beta12.m, CLGPNM.Beta13.m,
                CLGPNM.Beta21.m, CLGPNM.Beta22.m, CLGPNM.Beta23.m,
                CLGPNM.Sigma11.m, CLGPNM.Sigma22.m, CLGPNM.Sigma33.m,
                CLGPNM.Sigma12.m, CLGPNM.Sigma13.m, CLGPNM.Sigma23.m,
                CLGPNM.lambda.m)
hpd.LB.GPN <- c(CLGPNM.Beta11.hpd[1], CLGPNM.Beta12.hpd[1], CLGPNM.Beta13.hpd[1],
                CLGPNM.Beta21.hpd[1], CLGPNM.Beta22.hpd[1], CLGPNM.Beta23.hpd[1],
                CLGPNM.Sigma11.hpd[1], CLGPNM.Sigma22.hpd[1], CLGPNM.Sigma33.hpd[1],
                CLGPNM.Sigma12.hpd[1], CLGPNM.Sigma13.hpd[1], CLGPNM.Sigma23.hpd[1],
                CLGPNM.lambda.hpd[1])
hpd.UB.GPN <- c(CLGPNM.Beta11.hpd[2], CLGPNM.Beta12.hpd[2], CLGPNM.Beta13.hpd[2],
                CLGPNM.Beta21.hpd[2], CLGPNM.Beta22.hpd[2], CLGPNM.Beta23.hpd[2],
                CLGPNM.Sigma11.hpd[2], CLGPNM.Sigma22.hpd[2], CLGPNM.Sigma33.hpd[2],
                CLGPNM.Sigma12.hpd[2], CLGPNM.Sigma13.hpd[2], CLGPNM.Sigma23.hpd[2],
                CLGPNM.lambda.hpd[2])

modescon.GPN  <- c(CLGPNM.Betacon11.m, CLGPNM.Betacon12.m, CLGPNM.Betacon13.m,
                   CLGPNM.Betacon21.m, CLGPNM.Betacon22.m, CLGPNM.Betacon23.m,
                   CLGPNM.Sigmacon11.m, CLGPNM.Sigmacon22.m, CLGPNM.Sigmacon33.m,
                   CLGPNM.Sigmacon12.m, CLGPNM.Sigmacon13.m, CLGPNM.Sigmacon23.m,
                   CLGPNM.lambda.m)
hpdcon.LB.GPN <- c(CLGPNM.Betacon11.hpd[1], CLGPNM.Betacon12.hpd[1], CLGPNM.Betacon13.hpd[1],
                   CLGPNM.Betacon21.hpd[1], CLGPNM.Betacon22.hpd[1], CLGPNM.Betacon23.hpd[1],
                   CLGPNM.Sigmacon11.hpd[1], CLGPNM.Sigmacon22.hpd[1], CLGPNM.Sigmacon33.hpd[1],
                   CLGPNM.Sigmacon12.hpd[1], CLGPNM.Sigmacon13.hpd[1], CLGPNM.Sigmacon23.hpd[1],
                   CLGPNM.lambda.hpd[1])
hpdcon.UB.GPN <- c(CLGPNM.Betacon11.hpd[2], CLGPNM.Betacon12.hpd[2], CLGPNM.Betacon13.hpd[2],
                   CLGPNM.Betacon21.hpd[2], CLGPNM.Betacon22.hpd[2], CLGPNM.Betacon23.hpd[2],
                   CLGPNM.Sigmacon11.hpd[2], CLGPNM.Sigmacon22.hpd[2], CLGPNM.Sigmacon33.hpd[2],
                   CLGPNM.Sigmacon12.hpd[2], CLGPNM.Sigmacon13.hpd[2], CLGPNM.Sigmacon23.hpd[2],
                   CLGPNM.lambda.hpd[2])

CLGPNMres.tab <- round(cbind(modes.GPN, hpd.LB.GPN, hpd.UB.GPN, modescon.GPN, hpdcon.LB.GPN, hpdcon.UB.GPN),2)

rownames(CLGPNMres.tab) <- c("$\\beta_{0_s^{I}}$", "$\\beta_{0_s^{II}}$",
                             "$\\beta_{0_y}$", "$\\beta_{1_s^{I}}$",
                             "$\\beta_{0_s^{II}}$", "$\\beta_{1_y}$",
                             "$\\sum_{s_{1,1}}$", "$\\sum_{s_{2,2}}$",
                             "$\\sum_{y_{3,3}}$", "$\\sum_{s_{1,2}}$",
                             "$\\sum_{sy_{1,3}}$", "$\\sum_{sy_{2,3}}$",
                             "$\\lambda$")
colnames(CLGPNMres.tab) <- c("Mode", "HPD LB", "HPD UB", "Mode", "HPD LB", "HPD UB")

kable(CLGPNMres.tab, "latex", booktabs = TRUE, escape = FALSE, caption = "Results for the modified joint projected and skew normal model")%>%
add_header_above(c("Parameter" = 1, "Unconstrained" = 3, "Constrained" = 3))

```

Tables 2, 3 and 4 show results from the four models that were fit to the teacher
dataset. There are several parameters that we can compare between these results.
Firstly, we have the estimated mean of the circular outcome. In Table 5 we see 
the posterior (predictive) distributions for the estimated circular means of the
CL-PN, CL-GPN and joint GPN-SSN  model. For the CL-PN model we can actually also
compute this means from the estimates in Table 2 as follows: $\mu_{circ} = 
atan2(\beta_0^{II}, \beta_0^{I})$. This leads to an estimate of 31.53$^\circ$. 
The estimates for the CL-PN, CL-GPN and joint GPN-SSN  model are about equal and
correspond quite well to the actual data mean of 33.23$^\circ$ in Table
\ref{Tableteacherdescriptives}. The estimate from the Abe-Ley model which is
0.37 radians or 21.20$^\circ$ is different. This difference could be caused by
the fact that the densities for the circular ouctome, projected normal or
sine-skewed von Mises, differ between the models.

```{r means, echo = FALSE}
means <- round(rbind(mean_circ(res_CLPN$theta_pred[,5000:20000]),
                     mean_circ(res_CLGPN$theta_pred[,5000:20000]),
                     mean_circ(res_CLGPNM$theta_pred[,5000:20000]))*(180/pi), 2)

rownames(means) <- c("Cl-PN", "CL-GPN", "GPN-SSN")
colnames(means) <- "Mean (degrees)"
kable(means, "latex", booktabs = T, escape = FALSE, caption = "Posterior estimates for the circular mean in the CL-PN, CL-GPN and joint GPN-SSN models")%>%
   kable_styling() %>%
   add_footnote("Note that these means are based on the posterior predictive distribution for the intercepts following (Wang & Gelfand, 2013).", notation="alphabet")

```

We can also compare the effect of \verb|SE| on the circular outcome. For the 
CL-PN model and Abe-Ley models we can get estimates of a circular regression 
coefficient of this effect. For the Abe-Ley model this estimate is the parameter
$\nu_1 = -0.01$. This means that for each unit increate in self-efficacy, at the
inflection point of the circular regression line, the score of the teacher on 
the IPC decreases with $0.01*(180/\pi) = 0.57^\circ$. For the CL-PN model we 
can use methods from @CremersMulderKlugkist2017. The estimated posterior mode of
$b_c$, a parameter that is comparable to $\nu_1$ in the Abe-Ley model, equals 
$1.13$ and its 95\% HPD interval is $(-26.55; 25.46)$. This means that for each 
unit increate in self-efficacy, at the inflection point of the circular 
regression line, the score of the teacher on the IPC increases with 
$1.13*(180/\pi) = 74.74^\circ$. Values of $b_c$ and $\nu_1$ are quite 
different even though they should represent the same effect. However, the slope 
at the inflection point, which $b_c$ and $\nu_1$ describe is not necessarily 
representative of the effect of \verb|SE| in the data range. In Figure \ref{} we
see that the inflection point (square) for the CL-PN model lies just ouside of 
the data range and that for the Abe-Ley model even further (outside the x-axis 
range). The slope of both regression lines in the data range is much more alike
(even though one is positive and the other negative). The slope at the mean
Self-efficicay, which has a value of 0 because it was centered, $SAM$ is
estimated at $0.02 (-0.09; 0.15)$ The $SAM$ can be interpreted as the amount of
change on the circle for a unit increase at the mean of the predictor variable.
However, for the CL-PN model the 95\% HPD interval of $SAM$ for the effect of
\verb|SE| includes 0 meaning that the value is not different from 0. For the
Abe-Ley model standard errors of the parameters are not known so we cannot
formally test whether $\nu_1$ differs from 0. Neither is there a known way of
computing the effect at predictor values other than the one of the inflection
point.

For the CL-GPN and multivariate 
CL-GPN models we cannot compute circular regression coefficients. Instead, we 
will compute posterior predictive distributions for the predicted circular
outcome of individuals scoring 1 unit above and 1 unit below the average
self-efficacy.

```{r circreg, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}

tikz("regline.tex", standAlone =TRUE, height = 4, width = 6, pointsize = 12, engine = "pdftex")

a1 <- mode_est(res_CLPN$BI[5000:20000, 1])
b1 <- mode_est(res_CLPN$BI[5000:20000, 2])
a2 <- mode_est(res_CLPN$BII[5000:20000, 1])
b2 <- mode_est(res_CLPN$BII[5000:20000, 2])

x <- seq(-5, 5, by = 0.1)

pred.AL <- 0.37 + 2*atan(-0.01*x)
pred.PN <- atan2(a2 + b2*x, a1 + b1*x)

plot(Dat.test$SEc, theta.test, bty = "n",
     xlim = c(-5, 2), ylab = "$\\theta_{test}$", xlab = "Self-efficacy")
points(x, pred.AL, type = "l", lty = 1)
points(x, pred.PN, type = "l", lty = 2)

dev.off()
tools::texi2dvi("regline.tex", pdf=TRUE)

```

```{r reg.coef, cache = TRUE, echo = FALSE, results = FALSE}

a1 <- res_CLPN$BI[5000:20000, 1]
b1 <- res_CLPN$BI[5000:20000, 2]
a2 <- res_CLPN$BII[5000:20000, 1]
b2 <- res_CLPN$BII[5000:20000, 2]

ax <- -(a1*b1 + a2*b2)/(b1^2 + b2^2)
ac <- atan2(a2 + b2*ax, a1 + b1*ax)
bc <- -tan(atan2(a2, a1)-ac)/ax

mode_est(ax)

mode_est(bc)
hpd_est(bc)

mode_est(bc/(1 + (bc*(-ax))^2))
hpd_est(bc/(1 + (bc*(-ax))^2))

```

\begin{figure}
\centering
\includegraphics[width = 0.8\textwidth]{regline.pdf}
\caption{Plot showing circular regresion lines as predicted by the Abe-Ley (solid line) and CL-PN model (dashed line).}
\label{regline}
\end{figure}

Next, we compare the intercepts and regression coefficients for the linear outcome.

\begin{itemize}
\item How to interpret weibull regression for Abe-Ley model?
\item Comment on different distributions for linear and circular outcome ($\lambda$ in multivariate CL-GPN), integrate with model fit does this correspond. 
\item Comment on different estimation methods, in CL-PN and CL-GPN the regression of $\theta$ is independent of $y$ but not vice-versa. In the Abe-Ley and Multivariate method dependence between $\theta$ and $y$ is modelled both ways. 
\end{itemize}

\subsubsection{Model fit}

```{r ModelFit, echo = FALSE}

CLGPNM.ll.circ <- -2*mean(res_CLGPNM$ll.circ[5000:20000])
CLGPNM.ll.lin  <- -2*mean(res_CLGPNM$ll.lin[5000:20000])

CLGPN.ll.circ <- -2*mean(res_CLGPN$ll.circ[5000:20000])
CLGPN.ll.lin  <- -2*mean(res_CLGPN$ll.lin[5000:20000])

CLPN.ll.circ <- -2*mean(res_CLPN$ll.circ[5000:20000])
CLPN.ll.lin  <- -2*mean(res_CLPN$ll.lin[5000:20000])

AL.ll.circ <- -2*func.reg.cond.circ(resAL$par, dat.h)
AL.ll.lin <- -2*func.reg.cond.lin(resAL$par, dat.h)

PLSL <- round(rbind(c(CLPN.ll.circ, CLGPN.ll.circ, AL.ll.circ, CLGPNM.ll.circ),
                    c(CLPN.ll.lin, CLGPN.ll.lin, AL.ll.lin, CLGPNM.ll.lin)), 2)
rownames(PLSL) <- c("circular", "linear")
colnames(PLSL) <- c("CL-PN", "CL-GPN", "Abe-Ley", "Joint GPN-SSN")

kable(PLSL, "latex", booktabs = T, escape = FALSE, caption = "PLSL criteria for the circular and linear outcome in the four cylindrical models")

```

```{r preddist, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}

source("Sample Abe-Ley.R")

ALdy <- matrix(NA, length(y.test), 1000)
ALdth <- matrix(NA, length(theta.test), 1000)

ALdy.h <- matrix(NA, length(y.hold), 1000)
ALdth.h <- matrix(NA, length(theta.hold), 1000)

tikz("preddist.tex", standAlone =TRUE, height = 6, width = 6, pointsize = 12, engine = "pdftex")

layout(matrix(c(1,2,3, 4,5,5), 3, 2, byrow = TRUE),heights=c(2,2,1))

par(mar = c(5,5,1,1))

x.h <- Dat.hold$SEc
x <- Dat.test$SEc

for(i in 1:length(theta.test)){
  beta <- exp(resAL$par[3] + resAL$par[4]*x[i])
  mu <- resAL$par[1] + 2*atan(resAL$par[2]*x[i])

  dat <- rweiSSVM(1000, beta = beta, alpha = resAL$par[6], mu = mu, kappa = resAL$par[5], lambda = resAL$par[7])
  ALdy[i,] <- dat$x
  ALdth[i,] <- dat$theta
}

for(i in 1:length(theta.hold)){
  beta <- exp(resAL$par[3] + resAL$par[4]*x.h[i])
  mu <- resAL$par[1] + 2*atan(resAL$par[2]*x.h[i])

  dat <- rweiSSVM(1000, beta = beta, alpha = resAL$par[6], mu = mu, kappa = resAL$par[5], lambda = resAL$par[7])
  ALdy.h[i,] <- dat$x
  ALdth.h[i,] <- dat$theta

}

hist(theta.test, prob = TRUE, ylim = c(0, 1.25),
     main = "", ylab = "Circular", xlab = "$\\theta_{test}$")
lines(density(res_CLPN$theta_pred[,19000:20000]), lty = 1)
lines(density(res_CLGPN$theta_pred[,19000:20000]), lty = 2)
lines(density(res_CLGPNM$theta_pred[,19000:20000]), lty = 3)
lines(density(ALdth), lty = 4)

hist(theta.hold, prob = TRUE, ylim = c(0, 1.25),
     main = "", ylab = "Circular", xlab = "$\\theta_{holdout}$")
lines(density(res_CLPN$theta_pred.hold[,19000:20000]), lty = 1)
lines(density(res_CLGPN$theta_pred.hold[,19000:20000]), lty = 2)
lines(density(res_CLGPNM$theta_pred.hold[,19000:20000]), lty = 3)
lines(density(ALdth.h), lty = 4)

hist(y.test, prob = TRUE, ylim = c(0, 3),
     main = "", ylab = "Linear", xlab = "$y_{test}$")
lines(density(res_CLPN$y_pred[,5000:20000]), lty = 1)
lines(density(res_CLGPN$y_pred[,5000:20000]), lty = 2)
lines(density(res_CLGPNM$y_pred[,5000:20000]), lty = 3)
lines(density(ALdy), lty = 4)

hist(y.hold, prob = TRUE, ylim = c(0, 4),
     main = "", ylab = "Linear", xlab = "$y_{holdout}$")
lines(density(res_CLPN$y_pred.hold[,5000:20000]), lty = 1)
lines(density(res_CLGPN$y_pred.hold[,5000:20000]), lty = 2)
lines(density(res_CLGPNM$y_pred.hold[,5000:20000]), lty = 3)
lines(density(ALdy.h), lty = 4)

par(mar = c(2,2,1,1))

plot(1, type = "n", axes=FALSE, xlab="", ylab="")
legend(x = "top", inset = 0, legend = c("CL-PN", "CL-GPN", "Abe-Ley", "GPN-SSN"), lty = c(1,2,4,3), horiz = TRUE)

dev.off()

tools::texi2dvi("preddist.tex",pdf=TRUE)

```

\begin{figure}
\centering
\includegraphics[width = 0.8\textwidth]{preddist.pdf}
\caption{Histograms of $\boldsymbol{y}$ and $\boldsymbol{\theta}$ values of the test and holdout set of the teacher data plotted together with (posterior predictive) density estimates for the modified CL-PN, CL-GPN, Abe-Ley and joint GPN-SSN models.}
\label{preddist}
\end{figure}

\section{Discussion}\label{Discussion}

\begin{itemize}
\item Comment on difficulty of interpreting parameters in GPN model.
\end{itemize}

\newpage
\section*{References}
<div id="refs"></div>

\newpage
\begin{appendices}
\section{Appendix}\label{Appendix}

In this appendix we outline the MCMC procedures to fit the cylindrical regression models from Section \ref{Models}. R-code for the MCMC sampler and the analysis of the teacher data can be found here: \url{https://github.com/joliencremers/CylindricalComparisonCircumplex}.

\subsection{Bayesian Model and MCMC procedure for the modified CL-PN model}\label{A1}

We use the following algorithm to obtain posterior estimates from the model:

\begin{enumerate}

\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_i, \dots, \theta_n$, the linear outcome $\boldsymbol{y} = y_i, \dots, y_n$ where $n$ is the sample size, and the design matrices $\boldsymbol{Z}^k$ and $\boldsymbol{X}$ for the two components of the circular and the linear outcome respectively in a training (90\%) and holdout (10\%) set.

\item Define the prior parameters for the training set. In this paper we use:

\begin{itemize}
\item Prior for $\boldsymbol{\gamma}$: $N_q(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_{0})$, with  $\boldsymbol{\mu}_{0} = (0,0,0,0)^t$ and  $\boldsymbol{\Lambda}_{0} = 10^{-4}\boldsymbol{I}_4$.
\item Prior for $\sigma^2$: $IG(\alpha_{0}, \beta_{0})$, an inverse-gamma prior with $\alpha_{0} = 0.001$ and  $\beta_{0} = 0.001$.
\item Prior for $\boldsymbol{\beta^{k}}$: $N_2(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_{0})$, with $\boldsymbol{\mu}_{0} = (0,0)^t$ and  $\boldsymbol{\Lambda}_{0} = 10^{-4}\boldsymbol{I}_2$ for $k \in I,II$.
\end{itemize}

\item Set starting values $\boldsymbol{\gamma} = (0,0,0,0)^t$, $\sigma^2 = 1$ and $\boldsymbol{\beta^{k}} = (0,0)$ for $k \in I,II$. Also set starting values $r_i = 1$ in the training and holdout set.

\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:

$$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos \theta_i \\  r_i\sin \theta_i\end{bmatrix}$$

\item Sample $\boldsymbol{\gamma}$, $\sigma^2$ and $\boldsymbol{\beta^{k}}$ for $k \in I,II$ for the training dataset from their conditional posteriors:

\begin{itemize}
\item Posterior for $\boldsymbol{\gamma}$: $N_q(\boldsymbol{\mu}_n, \sigma^2\boldsymbol{\Lambda}^{-1}_n)$, with $\boldsymbol{\mu}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)^{-1}(\boldsymbol{\Lambda}_0\boldsymbol{\mu}_0 + \boldsymbol{X}^t\boldsymbol{y})$ and $\boldsymbol{\Lambda}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)$.
\item Posterior for $\sigma^2$: $IG(\alpha_{n}, \beta_{n})$, an inverse-gamma posterior with $\alpha_{n} = \alpha_0 + n/2$ and $\beta_{n} = \beta_0 + \frac{1}{2}(\boldsymbol{y}^t\boldsymbol{y} + \boldsymbol{\mu}_{0}^t\boldsymbol{\Lambda}_0\boldsymbol{\mu}_{0} + \boldsymbol{\mu}_{n}^t\boldsymbol{\Lambda}_n\boldsymbol{\mu}_{n})$.
\item Posterior for $\boldsymbol{\beta^{k}}$: $N_2(\boldsymbol{\mu}^k_n, \boldsymbol{\Lambda}^{k}_n)$, with $\boldsymbol{\mu}^k_n = ((\boldsymbol{Z}^k)^t\boldsymbol{Z}^k + \boldsymbol{\Lambda}^k_0)^{-1}(\boldsymbol{\Lambda}^k_0\boldsymbol{\mu}^k_0 + (\boldsymbol{Z}^k)^t\boldsymbol{s}^k)$ and $\boldsymbol{\Lambda}^{k}_n = ((\boldsymbol{Z}^k)^t\boldsymbol{Z}^k + \boldsymbol{\Lambda}^k_0)$.
\end{itemize}

\item Sample new $r_i$ for the training and holdout dataset from the following posterior:

$$f(r_i \mid \theta_i, \boldsymbol{\mu}_i) \propto r_i \exp{(-\frac{1}{2}(r_i)^2 + b_ir_i)}$$ 
where $b_i = \begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}^t\boldsymbol{\mu}_i$, $\boldsymbol{\mu}_i = \boldsymbol{z_i}\boldsymbol{B}$ and $\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$.

We can sample from this posterior using a slice sampling technique (Cremers et al., 2018):

\begin{itemize}
\item In a slice sampler the joint density for an auxiliary variable $v_{i}$ with $r_{i}$ is:

$$p(r_{i}, v_{i}\mid \theta_{i}, \boldsymbol{\mu}_{i}=\boldsymbol{z}_{i}\boldsymbol{B}) \propto r_{i} \textbf{I}\left(0 < v_i < \exp\left\{ -\frac{1}{2}(r_{i} - b_{i})^2\right\}\right)\textbf{I}(r_i > 0).$$

\noindent The full conditional for $v_{i}$, $p(v_{i} \mid r_{i},\boldsymbol{\mu}_{i}, \theta_{i})$, is:

$$U\left(0, \exp\left\{-\frac{1}{2}(r_{i} -  b _{i})^2\right\}\right)$$

and the full conditional for $r_i$, $p(r_{i} \mid v_{i},\boldsymbol{\mu}_{i}, \theta_{i})$, is proportional to:
$$r_{i} \textbf{I}\left(b_{i} + \max\left\{-b_{i}, -\sqrt{-2\ln v_{i}}\right\} < r_{i} < b_{i} + \sqrt{-2\ln v_{i}}\right).$$

\noindent We thus sample $v_{i}$ from the uniform distribution specified above. Independently we sample a value $m$ from $U(0,1)$. We obtain a new value for $r_{i}$ by computing $ r_{i} = \sqrt{(r_{i_{2}}^{2}-r_{i_{1}}^{2})m + r_{i_{1}}^{2}}$ where $r_{i_{1}}=b_{i} +\max\left\{-b_{i}, -\sqrt{-2\ln v_{i}}\right\}$ and $ r_{i_{2}}= b_{i} + \sqrt{-2\ln v_{i}}$.
\end{itemize}

\item Compute the PLSL for the circular and linear outcome on the holdout set using the estimates of $\boldsymbol{\gamma}$, $\sigma^2$ and $\boldsymbol{\beta^{k}}$ for $k \in I,II$ for the training dataset.

\item Repeat steps 4 to 8 until the sampled parameter estimates have converged.

\end{enumerate}

\newpage
\subsection{Bayesian Model and MCMC procedure for the modified CL-GPN mode}\label{A2}

We use the following algorithm to obtain posterior estimates from the model:

\begin{enumerate}

\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_i, \dots, \theta_n$, the linear outcome $\boldsymbol{y} = y_i, \dots, y_n$ where $n$ is the sample size, and the design matrices $\boldsymbol{Z}^k$ and $\boldsymbol{X}$ for the two components of the circular and the linear outcome respectively in a training (90\%) and holdout (10\%) set.

\item Define the prior parameters for the training set. In this paper we use:

\begin{itemize}
\item Prior for $\boldsymbol{\gamma}$: $N_q(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_{0})$, with $\boldsymbol{\mu}_{0} = (0,0,0,0)^t$ and $\boldsymbol{\Lambda}_{0} = 10^{-4}\boldsymbol{I}_4$.
\item Prior for $\sigma^2$: $IG(\alpha_{0}, \beta_{0})$, an inverse-gamma prior with $\alpha_{0} = 0.001$ and $\beta_{0} = 0.001$.
\item Prior for $\boldsymbol{\beta}_{j}$: $N_2(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_0)$, with $\boldsymbol{\mu}_{0} = (0,0)^t$ and  $\boldsymbol{\Sigma}_{0} = 10^{5}\boldsymbol{I}_2$ for $j \in 1, \dots, p$ where $p$ is the number of covariates in $\boldsymbol{Z}$.
\item Prior for $\rho$: $N(\mu_0, \sigma^2)$, with $\mu_0 = 0$ and $\sigma^2 = 10^{4}$.
\item Prior for $\tau$: $IG(\alpha_{0}, \beta_{0})$, an inverse gamma prior with $\alpha_{0} = 0.01$ and $\beta_{0} = 0.01$.
\end{itemize}

\item Set starting values $\boldsymbol{\gamma} = (0,0,0,0)^t$, $\sigma^2 = 1$, $\boldsymbol{\beta}_j = (0,0)^t$, $\rho = 0$, $\tau = 1$ and $\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1 \end{bmatrix}$. Also set starting values $r_i = 1$ in the training and holdout set.

\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:

$$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos \theta_i \\  r_i\sin \theta_i\end{bmatrix}$$

\item Sample $\boldsymbol{\gamma}$, $\sigma^2$, $\boldsymbol{\beta}_j$, $\rho$ and $\tau$ for the training dataset from their conditional posteriors:

\begin{itemize}
\item Posterior for $\boldsymbol{\gamma}$: $N_q(\boldsymbol{\mu}_n, \sigma^2\boldsymbol{\Lambda}^{-1}_n)$, with $\boldsymbol{\mu}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)^{-1}(\boldsymbol{\Lambda}_0\boldsymbol{\mu}_0 + \boldsymbol{X}^t\boldsymbol{y})$ and $\boldsymbol{\Lambda}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)$.
\item Posterior for $\sigma^2$: $IG(\alpha_{n}, \beta_{n})$, an inverse-gamma posterior where $\alpha_{n} = \alpha_0 + n/2$ and $\beta_{n} = \beta_0 + 0.5(\boldsymbol{y}^t\boldsymbol{y} + \boldsymbol{\mu}_{0}^t\boldsymbol{\Lambda}_0\boldsymbol{\mu}_{0} + \boldsymbol{\mu}_{n}^t\boldsymbol{\Lambda}_n\boldsymbol{\mu}_{n})$.
\item Posterior for $\boldsymbol{\beta}_j$: $N_2(\boldsymbol{\mu}_{j_{n}}, \boldsymbol{\Sigma}_{j_{n}})$, with $\boldsymbol{\mu}_{j_{n}} = \boldsymbol{\Sigma}_{j_{n}}\boldsymbol{\Sigma}^{-1}\Bigg(-\sum_{i=1}^{n}z_{i,j-1}\sum_{l\neq j}z_{i,l-1}\boldsymbol{\beta}_l + \sum_{i=1}^{n}z_{i,j-1}r_i\begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}\Bigg)$ and  $\boldsymbol{\Sigma}_{j_{n}} = \Big(\sum_{i=1}^{n}z_{i,j-1}^2\boldsymbol{\Sigma}^{-1}+\boldsymbol{\Lambda}_0\Big)^{-1}$ for $j \in 1, \dots, p$ where p is the number of covariates in $\boldsymbol{Z}$.
\item Posterior for $\rho$: $N(\mu_n, \sigma^2_n)$, with $\mu_n = \frac{\tau^{-2} \sum_{i=1}^{n}(s^{I}_{i} - \mu_i^{I})(s^{II}_{i} - \mu_i^{II}) + \mu_0\sigma_0^{-2}}{\tau^{-2}\sum_{i=1}^{n}(s^{II}_{i} - \mu^{II})^2 + \sigma_0^{-2}}$ and $\sigma_n^2 = \frac{1}{\tau^{-2}\sum_{i=1}^{n}(s^{II}_{i} - \mu^{II})^2 + \sigma_0^{-2}}$ where $\mu_i^{I} = \boldsymbol{z}_i\boldsymbol{\beta}^{I}$ and $\mu_i^{II} = \boldsymbol{z}_i\boldsymbol{\beta}^{II}$.
\item Posterior for $\tau$: $IG(\alpha_n, \beta_n)$, an inverse-gamma posterior with $\alpha_n = \frac{n}{2} + \alpha_0$ and $\beta_n = \sum\limits_{i = 1}^{n}(s^{I}_{i} - \{\mu_i^{II} + \rho(s^{II}_{i} - \mu^{II})\})^2 + \beta_0$
\end{itemize}

\item Sample new $r_i$ for the training and holdout dataset from the following posterior:

$$f(r_i \mid \theta_i, \boldsymbol{\mu}_i) \propto r_i \exp{\left\{-0.5A_i\bigg(r_i-\frac{B_i}{A_i}\bigg)^2\right\}}$$ 
where $B_i = \begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}^t\boldsymbol{\Sigma}^{-1}\boldsymbol{\mu}_i$, $\boldsymbol{\mu}_i = \boldsymbol{z_i}\boldsymbol{B}$, $\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$ and $A_i = \begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}^t\boldsymbol{\Sigma}^{-1}\begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}$.

We can sample from this posterior using a slice sampling technique (Hernandez-Stumpfhauser et.al. 2018):

\begin{itemize}
\item In a slice sampler the joint density for an auxiliary variable $v_{i}$ with $r_{i}$ is:

$$p(r_{i}, v_{i}\mid \theta_{i}, \boldsymbol{\mu}_{i}=\boldsymbol{z}_{i}\boldsymbol{B}^{t}) \propto r_{i} \textbf{I}\bigg(0 < v_i < \exp\left\{ -.5 A_i(r_{i} - \frac{B_i}{A_i})^2\right\}\bigg)\textbf{I}(r_i > 0)$$

\item The full conditional for $v_{i}$, $p(v_{i} \mid r_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$, is:

$$U\Bigg(0, \exp\left\{-.5A_i\bigg(r_i -  \frac{B_{i}}{A_i}\bigg)^2\right\}\Bigg)$$
and the full conditional for $r_i$, $p(r_{i} \mid v_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$, is proportional to:
$$r_{i} \textbf{I}\left(\frac{B_i}{A_i} + \max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\} < r_{i} < \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}\right)$$
\item We thus sample $v_{i}$ from the uniform distribution specified above. Independently we sample a value $m$ from $U(0,1)$. We obtain a new value for $r_{i}$ by computing $r_{i} = \sqrt{(r_{i_{2}}^{2}-r_{i_{1}}^{2})m + r_{i_{1}}^{2}}$ where $r_{i_{1}}=\frac{B_i}{A_i} +\max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\}$ and $ r_{i_{2}}= \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}$.

\end{itemize}

\item Compute the PLSL for the circular and linear outcome on the holdoutset using the estimates of $\boldsymbol{\gamma}$, $\sigma^2$, $\boldsymbol{\beta^{k}}$, $\rho$ and $\tau$ for the training dataset. Use the density $f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma})$ for the circular outcome.

\item Repeat steps 4 to 8 until the sampled parameter estimates have converged.

\end{enumerate}

\newpage
\subsection{Bayesian Model and MCMC procedure multivariate GPN model}\label{A3}

\begin{enumerate}
\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_i, \dots, \theta_n$, the linear outcome $\boldsymbol{y} = y_i, \dots, y_n$ where $n$ is the sample size, and the design matrix $\boldsymbol{X}$ in a training (90\%) and holdout (10\%) set. Note that in this paper we have only one circular outcome and one linear outcome and the MCMC procedure outlined here is specified for this situation. It can however be generalized to a situation with multiple circular and linear outcomes without too much effort.

\item Define the prior parameters for the training set. Note that in this paper we have only one circular outcome and one linear outcome, so $p = 1$ and $q = 1$. In this paper we use the following priors:

\begin{itemize}
\item Prior for $\boldsymbol{\Sigma}$: $IW(\boldsymbol{\Psi}_0, \nu_0)$, an inverse-Wishart with $\boldsymbol{\Psi}_0 = 10^{-4}\boldsymbol{I}_{2p + q}$ and $\nu_0 = 1$.   
\item Prior for $\boldsymbol{\beta}$: $MN(\boldsymbol{\beta}_0, \boldsymbol{\Sigma}_0  \otimes \boldsymbol{\kappa}_0)$, where $\boldsymbol{\beta}$ is a vectorized $\boldsymbol{B}$, the matrix with regression coefficients, $\boldsymbol{\beta}_0 = \boldsymbol{0}_{k(2p + q)}$, $\boldsymbol{B}_0 = \boldsymbol{0}_{k \times (2p + q)}$ and $\boldsymbol{\kappa}_0 = 10^{-4}\boldsymbol{I}_k$.
\item Prior for $\lambda$: $N(\gamma_0, \omega_0)$, with $\gamma_0 = 0$ and $\omega_0 = 10000$.
\end{itemize}

\item Set starting values $\boldsymbol{\beta} = (0,0,0,0,0,0)^t$, $\boldsymbol{\Sigma} = \boldsymbol{I}_3$ and $\lambda = 0$. Also set starting values $r_i = 1$ and $d_i = 1$ in the training and holdout set.

\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:

$$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos \theta_i \\  r_i\sin \theta_i\end{bmatrix}$$
\item Compute the latent outcomes $\tilde{y}_i$ underlying the linear outcome for the holdout and training dataset as follows:

$$\tilde{y}_i = \lambda d_i $$

\item Compute $\boldsymbol{\eta}_i$ defined as follows for each individual i:

$$\boldsymbol{\eta}_i = (\boldsymbol{s}_i,y_i)^t - (\boldsymbol{0}_{2p}, \lambda d_i)$$

\item Sample $\boldsymbol{B}$, $\boldsymbol{\Sigma}$ and $\lambda$ for the training dataset from their conditional posteriors:

\begin{itemize}
\item Posterior for $\boldsymbol{B}$: $MN(\boldsymbol{B}_n, \boldsymbol{\kappa}_n, \boldsymbol{\Sigma}_n)$, with $\boldsymbol{B}_n = \boldsymbol{\kappa}_n^{-1}\boldsymbol{X}^t\boldsymbol{\eta} + \boldsymbol{\kappa}_0\boldsymbol{B}_0$ and $\boldsymbol{\kappa}_n = \boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\kappa}_0$.
\item Posterior for $\Sigma$: $IW(\boldsymbol{\Psi}_n, \nu_n)$, an inverse-Wishart with $\boldsymbol{\Psi}_n = \boldsymbol{\Psi}_0 + (\boldsymbol{\eta} - \boldsymbol{X}^t\boldsymbol{B})^t(\boldsymbol{\eta} - \boldsymbol{X}^t\boldsymbol{B}) + (\boldsymbol{B} - \boldsymbol{B}_0)^t\boldsymbol{\kappa}_0(\boldsymbol{B} - \boldsymbol{B}_0)$ and $\nu_n = \nu_0 + n$
\item Posterior for $\lambda$: $N(\gamma_n, \omega_n)$, with $\omega_n = \big(\sum_{i = 1}^{n}d_i^2\boldsymbol{\Sigma}_{y|s}^{-1} + \omega_0^{-1}\big)^{-1}$ and $\gamma_n = \omega_n \big(\sum_{i = 1}^{n}d_i\boldsymbol{\Sigma}_{y|s}^{-1}(y_i - \boldsymbol{\mu}_{y_i|s_i}) + \omega_0^{-1}\gamma_0 \big)$
\end{itemize}

\item Sample new $d_i$ for the training and holdout dataset from the following posterior:

$$f(d_i) \propto \phi_q(y_i|\boldsymbol{\mu}_{y_i|s_i} + \lambda d_i, \boldsymbol{\Sigma}_{y|s})\phi_q(d_i|0, 1),$$

where $\boldsymbol{\mu}_{y_i|s_i} = \boldsymbol{x}_i\boldsymbol{B}_{y_i|s_i}$. We can see $d_i$ as a positive regressor with $\lambda$ as covariate and $\phi_q(d_i|0, 1)$ as prior (Mastrantonio, 2018). The full conditional is then a q-dimensional truncated normal with support $\mathbb{R}^{+}$ as follows:

$$N_q(\boldsymbol{M}_{d_i}, \boldsymbol{V}_q),$$ \\where $\boldsymbol{V}_q = \big(\lambda^2\boldsymbol{\Sigma}_{y|s}^{-1} + 1\big)$ and $\boldsymbol{M}_{d_i} = \boldsymbol{V}_d\lambda\boldsymbol{\Sigma}_{y|s}^{-1}\big(y_i - \boldsymbol{\mu}_{y_i, s_i}\big)$.

\item Sample new $r_i$ for the training and holdout dataset from the following posterior:

$$f(r_i \mid \theta_i, \boldsymbol{\mu}_i) \propto r_i \exp{\left\{-0.5A_i\bigg(r_i-\frac{B_i}{A_i}\bigg)^2\right\}}$$ 
where $B_i = \begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}^t\boldsymbol{\Sigma}_{s_i|y_i}^{-1}\boldsymbol{\mu}_{s_i|y_i}$, $\boldsymbol{\mu}_{s_i|y_i} = \boldsymbol{x}_i\boldsymbol{B}_{s_i|y_i}$ and $A_i = \begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}^t\boldsymbol{\Sigma}_{s_i| y_i}^{-1}\begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}$. The parameters $\boldsymbol{\mu}_{s_i|y_i}$ and $\boldsymbol{\Sigma}_{s_i| y_i}$ are the conditional mean and covariance matrix of $\boldsymbol{s}_i$ assuming that $(\boldsymbol{s}_i, y_i)^t \sim N_{2p+q}(\boldsymbol{\mu} + (\boldsymbol{0}_{2p}, \lambda d_i)^t, \boldsymbol{\Sigma})$.

Because in this paper $\boldsymbol{\theta}$ originates from a bivariate variable that is known we can simply define the $r_i$ as the euclidean norm of the bivariate datapoints. However, for didactic purposes continue with the explanation of the sampling procedure. We
can sample from posterior for $r_i$ using a slice sampling technique (Hernandez-Stumpfhauser et.al. 2018):

\begin{itemize}
\item In a slice sampler the joint density for an auxiliary variable $v_{i}$ with $r_{i}$ is:

$$p(r_{i}, v_{i}\mid \theta_{i}, \boldsymbol{\mu}_{i}=\boldsymbol{x}_{i}\boldsymbol{B}^{t}) \propto r_{i} \textbf{I}\bigg(0 < v_i < \exp\left\{ -.5 A_i(r_{i} - \frac{B_i}{A_i})^2\right\}\bigg)\textbf{I}(r_i > 0)$$

\item The full conditionals for $v_{i}$, $p(v_{i} \mid r_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$, is:

$$U\Bigg(0, \exp\left\{-.5A_i\bigg(r_i -  \frac{B_{i}}{A_i}\bigg)^2\right\}\Bigg)$$
and the full conditional for $r_i$, $p(r_{i} \mid v_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$ , is proportional to :
$$r_{i} \textbf{I}\left(\frac{B_i}{A_i} + \max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\} < r_{i} < \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}\right)$$
\item We thus sample $v_{i}$ from the uniform distribution specified above. Independently we sample a value $m$ from $U(0,1)$. We obtain a new value for $r_{i}$ by computing $r_{i} = \sqrt{(r_{i_{2}}^{2}-r_{i_{1}}^{2})m + r_{i_{1}}^{2}}$ where $r_{i_{1}}=\frac{B_i}{A_i} +\max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\}$ and $ r_{i_{2}}= \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}$.

\end{itemize}

\item Compute the PLSL for the circular and linear outcome on the holdoutset using the estimates of $\boldsymbol{B}$, $\boldsymbol{\Sigma}$ and $\lambda$ for the training dataset.

\item Repeat steps 4 to 10 until the sampled parameter estimates have converged.

\item In the MCMC sampler we have estimated an unconstrained $\boldsymbol{\Sigma}$. However, for identification of the model we need to apply the following constraint to both $\boldsymbol{\Sigma}$ and $\boldsymbol{\mu}$:

$$\boldsymbol{C} = \begin{bmatrix} \boldsymbol{C}_w & \boldsymbol{0}_{2p \times q} \\ \boldsymbol{0}_{2p \times q}^t & \boldsymbol{I}_q \end{bmatrix}$$
where $\boldsymbol{C}_w$ is a $2p \times 2p$ diagonal matrix with every $(2(j-1) + k)^{th}$ entry $> 0$ where $k \in 1, 2$ and $k = 1, \dots, p$ (Mastrantonio, 2018). The estimates $\boldsymbol{\Sigma}$ and $\boldsymbol{\mu}$ can then be related to their constrained versions, $\tilde{\boldsymbol{\Sigma}}$ and $\tilde{\boldsymbol{\mu}}$ as follows:

$$\boldsymbol{\mu} = \boldsymbol{C}\tilde{\boldsymbol{\mu}}$$
$$\boldsymbol{\Sigma} = \boldsymbol{C}\tilde{\boldsymbol{\Sigma}}\boldsymbol{C}.$$

\end{enumerate}

\end{appendices}