### 2018-07-02 Manuscript/ManuscriptMarkdownDifferentSE.tex

\documentclass[11pt,]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={A Comparison of Regression models for Cylindrical data in Psychology.},
            pdfkeywords={bla vla},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{A Comparison of Regression models for Cylindrical data in Psychology.}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Jolien Cremers\footnote{Corresponding author:
  \href{mailto:j.cremers@uu.nl}{\nolinkurl{j.cremers@uu.nl}}} \(^1\),
Helena J.M. Pennings\(^{2,3}\), Christophe Ley \(^{4}\)\\
\(^1\)Department of Methodology and Statistics, Utrecht University\\
\(^2\)TNO\\
\(^3\)Department of Education, Utrecht University\\
\(^4\)Department of Applied Mathematics, Computer Science and
Statistics, Ghent University}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage[table]{xcolor}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}

\usepackage{multirow}
\usepackage{appendix}
\usepackage{color}
\usepackage{hyperref}
\usepackage{subcaption}

\begin{document}
\maketitle
\begin{abstract}
Cylindrical dats is multivariate data which consists of a directional,
in this paper circular, component and a linear component. Examples of
cylindrical data in psychology include human navigation (direction and
distance of movement), eye-tracking research (direction and length of
saccades) and data from an interpersonal circumplex (type and strength
of interpersonal behavior). In this paper we adapt four models for
cylindrical data to include a regression of the circular and linear
component onto a set of covariates. Subsequently, we illustrate how to
fit these models and interpret their results on a dataset on the
interpersonal behavior of teachers.
\end{abstract}

\section{Introduction}\label{Introduction}

Cylindrical data are data that consist of a linear variable and a
directional variable. In this paper, the directional variable is
circular meaning that it consists of a single angle instead of a set of
angles. A circular variable is different from a linear variable in the
sense that it is measured on a different scale. Figure \ref{circline}
shows the difference between a circular scale (right) and a linear scale
(left). The most important difference is that on a circular scale the
datapoints 0\(^\circ\) and 360\(^\circ\) are connected and in fact
represent the same number while on a linear scale the two ends,
\(-\infty\) and \(\infty\) are not connected. This difference requires
us to use different statistical methods for circular variables (see e.g.
Fisher (1995) for an introduction to circular data and Mardia \& Jupp
(2000), Jammalamadaka \& Sengupta (2001) and Ley \& Verdebout (2017) for
a more elaborate overview).

Cylindrical data occur in several fields of research, such as for
instance in meteorology (García-Portugués, Crujeiras, \&
González-Manteiga, 2013), ecology (García-Portugués, Barros, Crujeiras,
González-Manteiga, \& Pereira, 2014) or marine research (Lagona, Picone,
Maruotti, \& Cosoli, 2015). Several types of data in psychology are also
of a cylindrical nature. For example, in research on human navigation in
the field of cognitive psychology both distance, a linear variable, and
direction, a circular variable, of movement are of interest (Chrastil \&
Warren, 2017). In eye-tracking research, which can be used for
investigating various cognitive processes (e.g. those involved in
reading a text), we can also speak of cylindrical data. When measuring
eye-movements we speak of the eye-movements themselves, the saccades,
and fixations, the periods of time between movements when the eyes are
looking at one point. Of the saccades both the direction, a circular
variable, and the duration, a linear variable, are of interest (for a
review of eye-tracking research see Rayner (2009)). Data from circumplex
measurement instruments, e.g.~the interpersonal circumplex as used in
personality psychology, are also of a cylindrical nature (see Section
\ref{Example} for a more detailed explanation).

In this paper we will discuss how a correct statistical treatment of
such cylindrical data can lead to new insights. In particular we will
show how cylindrical models pave the way for circular-linear and
linear-circular regression. We will do this for a motivating example,
the teacher data, from the field of educational psychology . In this
example, apart from modelling the relation between the linear and
circular component of a cylindrical variable we would also like to
predict the two components from a set of covariates in a regression
model. The teacher data will be further introduced in Sections
\ref{Example} and \ref{DataAnalysis}.

\begin{figure}
\centering
\includegraphics[width = 0.8\textwidth]{Plots/circline.pdf}
\caption{The difference between a linear scale (left) and a circular scale (right).}
\label{circline}
\end{figure}

As is the case for circular data, the analysis of cylindrical data
requires special methods. Several methods have been put forward to model
the relation between the linear and circular component of a cylindrical
variable. Some of these are based on regressing the linear component
onto the circular component using the following type of relation:
\[y = \beta_0 + \beta_1*\cos(\theta) + 
\beta_2*\sin(\theta)+ \epsilon,\] where \(y\) is the linear component
and \(\theta\) the circular component (Johnson \& Wehrly, 1978; Mardia
\& Sutton, 1978; Mastrantonio, Maruotti, \& Jona-Lasinio, 2015). Others
model the relation in a different way, e.g.~by specifying a multivariate
model for several linear and circular variables and modelling their
covariance matrix (Mastrantonio, 2018) or by proposing a joint
cylindrical distribution. For example, Abe \& Ley (2017) introduce a
cylindrical distribution based on a Weibull distribution for the linear
component and a sine-skewed von Mises distribution for the circular
component and link these through their shape and concentration
parameters respectively. However, none of the methods that have been
proposed thus far include additional covariates onto which both the
circular and linear component are regressed.

Our aim in this paper is to fill this gap in the literature by adapting
four existing cylindrical methods in such a way that they include a
regression of both the linear and circular component of a cylindrical
variable onto a set of covariates. First however, we will introduce the
teacher data in Section \ref{Example}. In Section \ref{Models} we
introduce the four modified models for cylindrical data that we use to
analyze the data from the motivating example. We also choose a model
selection criterion to compare the models. The teacher data will be
analysed in Section \ref{DataAnalysis}. The paper will be concluded with
a discussion in Section \ref{Discussion}. The Appendix contains
technical details.

\section{Teacher data}\label{Example}

The motivating example for this article comes from the field of
educational psychology and was collected for the studies on classroom
climate of Pennings et al. (2017), Claessens (2016) and Want (2015). An
indicator of classroom climate is the students' perceptions of their
teachers' interpersonal behavior. These interpersonal perceptions, both
in educational psychology as well as in other areas of psychology, can
be measured using circumplex measurement instruments (see Horowitz \&
Strack (2011) for an overview of many such instruments).

The Questionnaire on Teacher Interaction (QTI) (Wubbels, Brekelmans,
Brok, \& Tartwijk, 2006) is one such circumplex measurement instrument.
It is used to study student perceptions of their teachers' interpersonal
behavior and contains items that load on two dimensions: Agency and
Communion. Agency refers to the degree of power or control a teacher
exerts in interaction with his/her students. Communion refers to the
degree of friendliness or affiliation a teacher conveys in interaction
with his/her students. The loadings on the two dimensions of the QTI can
be placed in a two-dimensional space formed by Agency (vertical) and
Communion (horizontal), see Figure \ref{QTI}. Different parts of this
space are characterized by different teacher behavior, e.g. `helpful' or
`uncertain'. We call the two-dimensional space the interpersonal circle
(IPC). The idea is that the IPC is ``a continuous order with no
beginning or end'' (Gurtman, 2009, p. 2). We call such ordering a
circumplex ordering and the IPC is therefore often called the
interpersonal circumplex. The ordering also implies that scores on the
IPC could be viewed as a circular variable.

\begin{figure}
\centering
\includegraphics[width = 0.8\textwidth]{Plots/IPC-T.png}
\caption{The interpersonal circle for teachers (IPC-T). The words presented in
the circumference of the circle are anchor words to describe the type of
behavior located in each part of the IPC.}
\label{QTI}
\end{figure}

Cremers, Mainhard, \& Klugkist (2018a) show how data from the IPC can be
considered circular data and analyzed as such using a regression model.
The two-dimension scores Agency and Communion can be converted to a
circular score using the two-argument arctangent function in
\eqref{PredVal}, where \(A\) represents a score on the Agency dimension
and \(C\) represents a score on the Communion dimension. The resulting
circular variable \(\theta\) can then be modeled. However, when
two-dimensional data are converted to the circle we lose some
information, the length of the two-dimensional vector \((A, C)\),
\emph{i.e.}, its euclidean norm \(\mid\mid (A, C) \mid\mid\). This
length represents the strength of the type of interpersonal behavior a
teacher shows towards his/her students. In a cylindrical model we are
able to incorporate this information, and model a circular variable
\(\theta\) together with a linear variable corresponding to
\(\mid\mid (A, C) \mid\mid\). This leads to an improved analysis of data
from the IPC. In the next section we introduce several models that can
be used for more accurate and informative regression analysis on the
teacher data. Descriptives for the teacher data are given in Section
\ref{DataAnalysis}.

\begin{equation}\label{PredVal}
\theta          = \text{atan2}\left(A, \: C\right)  =
\left|{\begin{array}{lcl}
                                                                       \arctan\left(\frac{A}{C}\right) & \text{if}  \quad&C > 0 \\
\arctan\left(\frac{A}{C}\right) + \pi & \text{if}  \quad& C  <  0  \:\: \&\:\: A \geq 0\\
 \arctan\left(\frac{A}{C}\right) - \pi & \text{if}  \quad&C  <  0 \:\:  \&\:\:A  < 0\\
 +\frac{\pi}{2} & \text{if}  \quad& C  =  0  \:\: \&\:\:A > 0\\
 -\frac{\pi}{2} & \text{if}  \quad& C =  0  \:\: \&\:\:A < 0\\
 \text{undefined} & \text{if} \quad& C =  0   \:\: \&\:\:A = 0.
 \end{array}}
\right.
\end{equation}

\section{Four cylindrical regression models}\label{Models}

In this section we introduce four cylindrical regression models to
contain predictors for the linear and circular outcomes, \(\Theta\) and
\(Y\). We extend the models from Mastrantonio (2018) and Abe \& Ley
(2017). Additionally, we introduce two models where the relation between
\(\Theta\) and \(Y\) is modelled as follows (following Mastrantonio et
al. (2015)):

\begin{equation}\label{circlinlink}
y = \gamma_0 + \gamma_{cos}*\cos(\theta)*r + \gamma_{sin}*\sin(\theta)*r + \gamma_1*x_1 + \dots + \gamma_q*x_q +  \epsilon,
\end{equation}

where \(r\) will be introduced in Section \ref{CL-(G)PN}, the error term
\(\epsilon \sim N(0, \sigma)\),
\(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are
the intercept and regression coefficients and \(x_1, \dots, x_q\) are
the \(q\) covariate values. In this model \(Y\) follows a normal
distribution and \(\Theta\) a projected normal (PN) or general projected
normal (GPN) distribution on the circle.

\subsection{The modified CL-PN and modified CL-GPN  models}\label{CL-(G)PN}

In both of these models the relation between \(\Theta \in [0, 2\pi)\)
and \(Y\in (-\infty, + \infty)\) is specified as in \eqref{circlinlink}
with the following conditional distribution:

\begin{equation}\label{ycondtheta}
f(y \mid \theta) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp\left[\frac{c^2 + (y - (\gamma_0 + \gamma_1x_1 + \dots + \gamma_qx_q))^{2} - 2c(y - (\gamma_0 + \gamma_1x_1 + \dots + \gamma_qx_q))}{2\sigma^2}\right],
\end{equation}

where
\(c = \begin{bmatrix} r \cos \theta \\ r\sin \theta \end{bmatrix}^t \begin{bmatrix} \gamma_{cos} \\ \gamma_{sin} \end{bmatrix}\),
\(r \geq 0\),
\(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are
the intercept and regression coefficients and \(\sigma^2 \geq 0\) is the
error variance. The linear outcome thus has a normal distribution,
conditional on \(\Theta\).

For the circular outcome we assume either a projected normal (PN) or a
general projected normal (GPN) distribution. These distributions arise
from the radial projection of a distribution defined on the plane onto
the circle. The relation between a bivariate variate \(\boldsymbol{S}\)
in the plane and the circular outcome \(\Theta\) is defined as follows:

\begin{equation}\label{projection}
\boldsymbol{S} = \begin{bmatrix} S^{I} \\ S^{II} \end{bmatrix} = R\boldsymbol{u} = \begin{bmatrix} R \cos \Theta \\  R\sin \Theta \end{bmatrix},
\end{equation}

where \(R = \mid\mid S \mid\mid\), the euclidean norm of the bivariate
vector \(\boldsymbol{S}\). In the PN distribution we assume
\(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{I})\) and in the
GPN we assume
\(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{\Sigma})\) where
\(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1 \end{bmatrix}\),
\(\rho \in (-\infty, +\infty)\) and \(\tau^2 \geq 0\) (as in
Hernandez-Stumpfhauser, Breidt, \& Woerd (2016)).

Following Nuñez-Antonio, Gutiérrez-Peña, \& Escarela (2011), the joint
density of \(\Theta\) and \(R\) for the PN distribution in a regression
set-up equals:

\begin{equation}\label{pnreg}
f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = [2\pi]^{-1} \exp\left[ \frac{-r^2 - \boldsymbol{\mu}^2\boldsymbol{\mu} + 2r\boldsymbol{u}^t\boldsymbol{\mu}}{2}\right],
\end{equation}

In a regression setup the outcome \(\theta_i\) for each individual
\(i = 1, \dots, n\), where \(n\) is the sample size, is generated
independently from \eqref{pnreg}. The mean vector
\(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is defined as
\(\boldsymbol{\mu}_i = \boldsymbol{z}_i\boldsymbol{B}\). The vector
\(\boldsymbol{z}_i\) is a a vector of \(p\) covariate values and
\(\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})\)
contain the regression coefficients. Note however that the dimensions of
\(\boldsymbol{\beta}^{I}\) and \(\boldsymbol{\beta }^{I}\) need not
necessarily be the same and we are thus allowed to have a different set
of predictor variables and vectors \(\boldsymbol{z}_i^I\) and
\(\boldsymbol{z}_i^{II}\) for the two components of
\(\boldsymbol{\mu}_i\).

Following Wang \& Gelfand (2013) and Hernandez-Stumpfhauser et al.
(2016) the joint density of \(r\) and \(\theta\) for the GPN
distribution equals:

\begin{equation}\label{gpnreg}
f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma}) = r(2\pi\tau)^{-1} \exp\left[ -0.5 \sigma^2(r\boldsymbol{u}-\boldsymbol{\mu})^{t}\boldsymbol{\Sigma}^{-1}(r\boldsymbol{u}-\boldsymbol{\mu})\right],
\end{equation}

where
\(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1 \end{bmatrix}\),
\(\boldsymbol{u}= \begin{bmatrix} \cos \theta \\ \sin \theta \end{bmatrix}\).
In a regression setup the outcome \(\theta_i\) for each individual is
generated independently from \eqref{gpnreg}. The mean vector
\(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is defined as
\(\boldsymbol{\mu}_i = \boldsymbol{z}_i(\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})\).
The vector \(\boldsymbol{z}_i\) is a a vector of \(p\) covariate values
and each \(\boldsymbol{\beta}^{k}\) is a vector with two regression
coefficients, one for each of the two components of
\(\boldsymbol{\mu}_i\). Note that for the CL-GPN model we do need to
have the same predictors for both components of \(\boldsymbol{\mu}_i\).

Both cylindrical models introduced here are estimated using MCMC methods
based on Nuñez-Antonio et al. (2011), Wang \& Gelfand (2013) and
Hernandez-Stumpfhauser et al. (2016) for the regression of the circular
outcome. A detailed description of the Bayesian estimation and MCMC
samplers can be found in Appendices \ref{A1} and \ref{A2}.

\subsection{The modified Abe-Ley model}\label{WeiSSVM}

This model is an extension of the cylindrical model introduced by Abe \&
Ley (2017) to the regression context. The joint density of \(\Theta\)
and \(Y\), in this model defined on the positive real half-line
\([0, + \infty)\), is:

\begin{equation}\label{WeiSSVMdensity}
f(\theta, y) = \frac{\alpha(\beta)^\alpha}{2\pi\cosh(\kappa)}
                 (1 +\lambda\sin(\theta - \mu))
                 y^{\alpha-1}
                 \exp[-((\beta y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu)))],
\end{equation}

In a regression setup the outcome vector \((\theta_i, y_i)^t\) for each
individual is generated independently from \eqref{WeiSSVMdensity}. The
parameters \(\alpha > 0\) and
\(\beta_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\nu}) > 0\) are linear
shape and scale parameters,
\(\mu_i = \eta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}) \in [0, 2\pi)\),
\(\kappa > 0\) and \(\lambda \in [-1, 1]\) are circular location,
concentration and skewness parameters. The parameter
\(\boldsymbol{\nu}\) is a vector of \(q\) regression coefficients
\(\nu_j \in (-\infty, +\infty)\) for the prediction of \(y\) where
\(j = 0, \dots, q\) and \(\nu_0\) is the intercept. The parameter
\(\eta_0 \in [0, 2\pi)\) is the intercept and \(\boldsymbol{\eta}\) is a
vector of \(p\) regression coefficients
\(\eta_j \in (-\infty, +\infty)\) for the prediction of \(\theta\) where
\(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a vector of
predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\) is
a vector of predictor values for the prediction of \(\theta\).

As in Abe \& Ley (2017), the conditional distribution of \(y\) given
\(\theta\) is a Weibull distribution with scale \(\alpha\) and shape
\(\beta(1-\tanh(\kappa)\cos(\theta - \mu))^{1-\alpha}\) and the
conditional distribution of \(\theta\) given \(y\) is a sine skewed von
Mises distribution with location parameter \(\mu\) and concentration
parameter \((\beta y)^\alpha\tanh(\kappa)\).

The log-likelihood for this model equals:

\begin{align}\label{WeiSSVMLikelihood}
l(\alpha, \boldsymbol{\nu}, \lambda, \kappa, \boldsymbol{\eta}) 
   &= n[\ln(\alpha) - \ln(2\pi\cosh(\kappa))] + \alpha \sum^{n}_{i = 1} \ln(\exp(\boldsymbol{x}_i^t\boldsymbol{\nu})) \nonumber\\
   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - \eta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}))) 
   +(\alpha-1)\sum^{n}_{i = 1} \ln(y_i) \nonumber\\
   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\nu})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - \eta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta})))
\end{align}

We can use numerical optimization (Nelder-Mead) to find solutions for
the maximum likelihood (ML) estimates for the parameters of the model.

\subsection{Modified joint projected and skew normal (GPN-SSN)}\label{CL-GPN_multivariate}

This model is an extension of the cylindrical model introduced by
Mastrantonio (2018) to the regression context. It contains \(p\)
circular outcomes and \(q\) linear outcomes. The circular outcomes
\(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots, \boldsymbol{\Theta}_p)\)
are modeled together by a multivariate GPN distribution and the linear
outcomes
\(\boldsymbol{Y} = (\boldsymbol{Y}_1, \dots, \boldsymbol{Y}_q)\) are
modeled together by a multivariate skew normal distribution (Sahu, Dey,
\& Branco, 2003). Because the GPN distribution is modelled using a
so-called augmented representation (as in \eqref{projection} and
\eqref{gpnreg}) it is convenient to use a similar tactic for modelling
the multivariate skew normal distribution. Following Mastrantonio (2018)
the linear outcomes are represented as:

\[\boldsymbol{Y} = \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D} + \boldsymbol{H},\]
where where \(\boldsymbol{\mu}_y\) is a mean vector for the linear
outcome \(\boldsymbol{Y}\),
\(\boldsymbol{\Lambda} = \text{diag}(\boldsymbol{\lambda})\) is a
\(q \times q\) diagonal matrix with diagonal elements
\(\boldsymbol{\lambda} = \lambda_1, \dots, \lambda_q\) (skewness
parameters),
\(\boldsymbol{D} \sim HN_q(\boldsymbol{0}_q, \boldsymbol{I}_q)\), a
q-dimensional half normal distribution (Olmos, Varela, Gómez, \&
Bolfarine, 2012), and
\(\boldsymbol{H} \sim N_q(\boldsymbol{0}_q, \boldsymbol{\Sigma}_y)\).
This means that conditional on the auxiliary data \(\boldsymbol{D}\),
\(\boldsymbol{Y}\) is normally distributed with mean
\(\boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D}\) and
covariance matrix \(\boldsymbol{\Sigma}_y\). The joint density for
\((\boldsymbol{Y}, \boldsymbol{D})^t\) is defined as follows:

\begin{equation}\label{YDjoint}
f(\boldsymbol{y}, \boldsymbol{d}) = 2^q\phi_q(\boldsymbol{y} \mid \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{d}, \boldsymbol{\Sigma}_y) \phi_1(\boldsymbol{d} \mid \boldsymbol{0}_q, \boldsymbol{I}_q).
\end{equation}

As in Mastrantonio (2018) dependence between the linear and circular
outcome is created by modelling the augmented representations of
\(\boldsymbol{\Theta}\), and \(\boldsymbol{Y}\) together in a
\(2 \times p + q\) dimensional normal distribution. The joint density of
the model is then represented by:

\begin{equation}\label{YDThetarjoint}
f(\boldsymbol{\theta}, \boldsymbol{r}, \boldsymbol{y}, \boldsymbol{d}) = 2^q\phi_{2p+q}((\boldsymbol{s}, \boldsymbol{y})^t \mid \boldsymbol{\mu} + (\boldsymbol{0}_{2p}, diag(\boldsymbol{\lambda})\boldsymbol{d})^t, \boldsymbol{\Sigma}) \phi_q(\boldsymbol{d} \mid \boldsymbol{0}_q, \boldsymbol{I}_q) \prod_{j = 1}^{p}r_j,
\end{equation}

where the mean vector
\(\boldsymbol{\mu} = (\boldsymbol{\mu}_s, \boldsymbol{\mu}_y)^t\) and
\(\boldsymbol{\Sigma} = \left ( \begin{matrix} \boldsymbol{\Sigma}_s & \boldsymbol{\Sigma}_{sy} \\ \boldsymbol{\Sigma}_{sy^t} & \boldsymbol{\Sigma}_y \\ \end{matrix} \right )\).
The matrix \(\boldsymbol{\Sigma}_s\) is the covariance matrix for the
variances of and covariances between the augmented representations of
the circular outcome and the matrix \(\boldsymbol{\Sigma}_{sy}\)
contains covariances between the augmented representations of the
circular outcome and the linear outcome.

In our regression extension we have \(i = 1, \dots, n\) observations of
\(p\) circular outcomes, \(q\) linear outcomes and \(k\) covariates. The
mean in the density in \eqref{YDThetarjoint} then becomes
\(\boldsymbol{\mu}_i = \boldsymbol{x}_i\boldsymbol{B}\) where
\(\boldsymbol{B}\) is a \(k \times (2 \times p + q)\) matrix with
regression coeffients. We estimate the model using MCMC methods. A
detailed description of these methods is given in Appendix \ref{A3}.

\subsection{Model fit criterion}\label{Modelfit}

For the four cylindrical models we focus on their out-of-sample
predictive performance to determine the fit of the model. To do so we
split our data in a training and holdout set (10 \(\%\) of the sample).
A proper criterion to compare out-of-sample predictive performance is
the the Predictive Log Scoring Loss (PLSL) (Gneiting \& Raftery, 2007).
The lower the value of this criterion, the better the predictive
performance of the model. Using ML estimates this criterion can be
computed as follows:

\begin{equation}\label{PLSLML}
PLSL = -2 \sum_{i = 1}^{M}\log l(x_i \mid \hat{\boldsymbol{\vartheta}}),
\end{equation}

where \(l\) is the model likelihood, \(M\) is the sample size of the
holdout set, \(x_i\) is the \(i^{th}\) datapoint from the holdout set
and \(\hat{\boldsymbol{\vartheta}}\) are the ML estimates of the model
parameters. Using posterior samples the criterion is similar to the log
pointwise predictive density (lppd) as outlined in Gelman et al. (2014)
and can be computed as:

\begin{equation}\label{PLSLBayes}
PLSL = -2 \frac{1}{B} \sum_{j = 1}^{B}\sum_{i = 1}^{M} \log l(x_i \mid \boldsymbol{\vartheta}^{(j)}),
\end{equation}

where \(B\) is the amount of posterior samples and
\(\boldsymbol{\vartheta}^{(j)}\) are the posterior estimates of the
model parameters for the \(j^{th}\) iteration. Note that although we fit
the CL-PN, CL-GPN and joint GPN-SSN models using Bayesian statistics, we
do not take prior information into account when assessing model fit.
According to Gelman et al. (2014) \textcolor{red}{(p. ???)} this is not
necessary since we are assessing the fit of a model to data, the holdout
set, only. They argue that the prior in such case is only of interest
for estimating the parameters of the model but not for determining the
predictive accuracy. For each of the cylindrical models we can then
compute a PLSL for the circular and linear outcome by using the
conditional log-likelihoods of the respective outcome.

\section{Data Analysis}\label{DataAnalysis}

In this section we analyze the teacher data with the help of the four
cylindrical models from Section \ref{Models}. First however, we give a
more detailed description of the dataset.

\subsection{Data description}\label{DataDescriptives}

The teacher data was collected between 2010 and 2015 and contains
several repeated measures on the IPC of 161 teachers. Measurements were
obtained using the QTI and taken in different years and classes. For
this paper we only consider one measurement, the first occasion (2010)
and largest class if data for multiple classes were available. In
addition to the score on the IPC, the circular outcome, and the strength
of the score on the IPC, the linear outcome, a teachers' self-efficacy
(\verb|SE|) concerning classroom management ia used as covariate in the
analysis. After the removal of missings we have a sample of 148 teachers
of which 10\% (n = 15) is used as a holdout set for the computation of
the PLSL. Table \ref{Tableteacherdescriptives} shows descriptives for
the dataset. Figure \ref{dataplot} is a scatterplot showing the relation
between the linear and circular outcome of the teacher data.

\begin{figure}
\centering
\includegraphics[width = \textwidth]{Plots/dataplot.pdf}
\caption{Plot showing the relation between the linear and circular component of the teacher data.}
\label{dataplot}
\end{figure}

\begin{table}
\centering
\caption{Descriptives for the teacher dataset} 
\begin{tabular}{lrrrl}
  \noalign{\smallskip}\hline\noalign{\smallskip}
Variable & mean/$\bar{\theta}$ & sd/$\hat{\rho}$ & Range & Type \\ \hline\noalign{\smallskip}
IPC &33.22$^\circ$& 0.76 & - & Circular\\
strength IPC & 0.43 & 0.15 & 0.08 - 0.80 & Linear\\
SE & 5.04 & 1.00 & 1.5 - 7.0 & Linear\\
   \hline
\end{tabular}
\label{Tableteacherdescriptives}
\caption*{Note that $\hat{\rho}$ is an sample estimate for the circular concentration where a value of 0 means that the data is not concentrated at all, i.e. spread over the entire circle, and a value of 1 means that all data is concentrated at a single point on the circle. }
\end{table}

\subsection{Models}\label{DataModels}

The regression equations for the linear and the circular outcome of the
four cylindrical models fit to the teacher dataset are as follows:

\begin{itemize}
\item For the modified CL-PN and CL-GPN models:

$\hat{\boldsymbol{\mu}}_{i} = \begin{pmatrix}
  \mu_{i}^{I}  \\
\mu_{i}^{II}
 \end{pmatrix}=\begin{pmatrix}
  \beta_0^{I} + \beta_1^{I}\text{SE}_i  \\
  \beta_0^{II} + \beta_1^{II}\text{SE}_i
 \end{pmatrix},$

$\hat{y}_i = \gamma_0 + \gamma_{cos}\cos\theta_ir_i + \gamma_{sin}\sin\theta_ir_i + \gamma_1\text{SE}_i.$

\item For the modified Abe-Ley model:

$\hat{\mu}_{i} = \eta_0 + 2 * \tan^{-1}(\eta_1\text{SE}_i),$

$\hat{\beta}_{i} = \exp(\nu_0 + \nu_1\text{SE}_i).$

\item For the modified joint projected and skew normal model:

$\hat{\boldsymbol{\mu}}_{i} = \boldsymbol{\beta}_0 + \boldsymbol{\beta}_1\text{SE}_i$, where $\boldsymbol{\mu}_i = (\boldsymbol{\mu}_{s_i}, \boldsymbol{\mu}_{y_i})^t$, $\boldsymbol{\beta}_0 = (\beta_{0_s^{I}}, \beta_{0_s^{II}},\beta_{0_y})$ and $\boldsymbol{\beta}_1 = (\beta_{1_s^{I}}, \beta_{1_s^{II}},\beta_{1_y})$.
\end{itemize}

We use the loglikelihoods of the following conditional densities for the
computation of the PLSL:

\begin{itemize}
\item for the modified CL-PN model:

$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{pnreg}.

\item for the modified CL-GPN model:

$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{gpnreg}.

\item for the modified Abe-Ley model:

$y_i \mid \boldsymbol{\theta}_i, \beta_i, \mu_i, \kappa, \alpha \sim W\left(\beta_i(1-\tanh(\kappa)\cos(\theta_i - \mu_i))^{1/\alpha}, \alpha\right)$, a Weibull distribution.

$\theta_i \mid y_i, \beta_i, \mu_i, \kappa, \alpha \lambda \sim SSVM\left(\mu_i, (\beta_iy_i)^{\alpha}(\tanh{\kappa})\right)$, a sine-skewed von Mises distribution.

\item for the modified joint projected and skew normal model:

$y_i \mid \boldsymbol{\mu}_i, \boldsymbol{\Sigma}, \boldsymbol{\theta}_i, r_i \sim SSN(\mu_{i_y} + \lambda d_i + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_w^{-1}(\boldsymbol{s}_i - \boldsymbol{\mu}_{i_s}), \boldsymbol{\Sigma}_y + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_s\boldsymbol{\Sigma}_{sy}),$

$\theta_i \mid \boldsymbol{\mu}_i, \boldsymbol{\Sigma}, y_i, d_i \sim GPN(\boldsymbol{\mu}_{i_s} + \boldsymbol{\Sigma}_{sy}\boldsymbol{\Sigma}_y^{-1} (y_i - \mu_{i_y} - \lambda d_i), \boldsymbol{\Sigma}_s + \boldsymbol{\Sigma}_{sy}\boldsymbol{\Sigma}_y^{-1}\boldsymbol{\Sigma}_{sy}^t)$

where $SSN$ is the skew normal distribution.

\end{itemize}

\subsection{Results \& Analysis}\label{DataResults}

Before analysis there were a couple of settings that we had to specify
for the cylindrical models. Starting values for the Abe-Ley model were
the following
\(\eta_0 = 0.9, \eta_1 = 0.9, \nu_0 = 0.9, \nu_1 = 0.9, \kappa = 0.9, \alpha = 0.9, \lambda = 0\).
The initial amount of iterations for the three MCMC samplers was set to
2000. After we checked convergence via traceplots we concluded that some
of the parameters of the joint GPN-SSN model did not converge. Therefore
we set the amount of iterations of the MCMC models to 20,000 and
subtracted a burn-in of 5000. With this specification the MCMC chains
for all parameters converge. Note that we choose the same amount of
iterations for all three Bayesian cylindrical models to make their
comparison via the PLSL as fair as possible. Lastly, the predictor
\verb|SE| was centered before inclusion in the analysis.

\begin{table}

\caption{\label{tab:estCLGPN}Results for the modified CL-PN and CL-GPN model}
\centering
\begin{tabular}[t]{lrrrrrr}
\toprule
\multicolumn{1}{c}{Parameter} & \multicolumn{3}{c}{CL-PN} & \multicolumn{3}{c}{CL-GPN} \\
\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-4} \cmidrule(l{2pt}r{2pt}){5-7}
  & Mode & HPD LB & HPD UB & Mode & HPD LB & HPD UB\\
\midrule
$\beta_0^{I}$ & 1.75 & 1.46 & 2.00 & 2.39 & 1.87 & 3.00\\
$\beta_1^{I}$ & 0.62 & 0.37 & 0.84 & 0.74 & 0.36 & 1.19\\
$\beta_0^{II}$ & 1.12 & 0.89 & 1.33 & 1.40 & 1.12 & 1.73\\
$\beta_1^{II}$ & 0.54 & 0.36 & 0.76 & 0.65 & 0.43 & 0.90\\
$\gamma_0$ & 0.37 & 0.31 & 0.43 & 0.36 & 0.30 & 0.41\\
\addlinespace
$\gamma_{cos}$ & 0.04 & 0.02 & 0.07 & 0.03 & 0.01 & 0.05\\
$\gamma_{sin}$ & -0.02 & -0.05 & 0.02 & -0.01 & -0.04 & 0.02\\
$\gamma_1$ & 0.03 & 0.00 & 0.07 & 0.03 & 0.00 & 0.06\\
$\sigma$ & 0.14 & 0.12 & 0.16 & 0.13 & 0.12 & 0.16\\
$\sum_{1,1}$ & NA & NA & NA & 3.02 & 1.87 & 4.97\\
\addlinespace
$\sum_{1,2}$ & NA & NA & NA & 0.59 & 0.20 & 0.88\\
$\sum_{2,2}$ & NA & NA & NA & 1.00 & 1.00 & 1.00\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption{\label{tab:estSL}Results for the modified Abe-Ley model}
\centering
\begin{tabular}[t]{lr}
\toprule
\multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{ML-estimate} \\
\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-2}
$\eta_0$ & 0.36\\
$\eta_1$ & -0.02\\
$\nu_0$ & 1.19\\
$\nu_1$ & 0.01\\
$\alpha$ & 3.82\\
\addlinespace
$\kappa$ & 1.58\\
$\lambda$ & 0.68\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}

\caption{\label{tab:estCLMGPN}Results for the modified joint projected and skew normal model}
\centering
\begin{tabular}[t]{lrrrrrr}
\toprule
\multicolumn{1}{c}{Parameter} & \multicolumn{3}{c}{Unconstrained} & \multicolumn{3}{c}{Constrained} \\
\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-4} \cmidrule(l{2pt}r{2pt}){5-7}
  & Mode & HPD LB & HPD UB & Mode & HPD LB & HPD UB\\
\midrule
$\beta_{0_s^{I}}$ & 0.31 & 0.26 & 0.34 & 2.07 & 1.77 & 2.50\\
$\beta_{0_s^{II}}$ & 0.19 & 0.16 & 0.21 & 1.28 & 1.06 & 1.53\\
$\beta_{0_y}$ & 0.33 & 0.30 & 0.36 & 0.33 & 0.30 & 0.36\\
$\beta_{1_s^{I}}$ & 0.08 & 0.04 & 0.12 & 0.54 & 0.28 & 0.82\\
$\beta_{1_s^{II}}$ & 0.07 & 0.04 & 0.09 & 0.46 & 0.29 & 0.65\\
\addlinespace
$\beta_{1_y}$ & 0.08 & 0.05 & 0.12 & 0.08 & 0.05 & 0.12\\
$\sum_{s_{1,1}}$ & 0.05 & 0.04 & 0.06 & 2.30 & 1.66 & 3.34\\
$\sum_{s_{2,2}}$ & 0.02 & 0.02 & 0.03 & 1.00 & 1.00 & 1.00\\
$\sum_{y_{3,3}}$ & 0.03 & 0.02 & 0.04 & 0.03 & 0.02 & 0.04\\
$\sum_{s_{1,2}}$ & 0.00 & 0.00 & 0.01 & 0.09 & -0.20 & 0.35\\
\addlinespace
$\sum_{sy_{1,3}}$ & 0.03 & 0.03 & 0.05 & 0.23 & 0.18 & 0.32\\
$\sum_{sy_{2,3}}$ & 0.01 & 0.01 & 0.02 & 0.09 & 0.05 & 0.12\\
$\lambda$ & 0.15 & 0.13 & 0.17 & 0.15 & 0.13 & 0.17\\
\bottomrule
\end{tabular}
\end{table}

Tables 2, 3 and 4 show results from the four models that were fit to the
teacher dataset. For the models in which Bayesian estimation was used we
show both the estimated posterior mode and the 95\% highest posterior
density (HPD) interval for each parameter. Before we turn to an
evaluation of the fit of the models, we discuss results of and the
interpretation of parameters from the four models. We will cover
parameters concerning the circular outcome, linear outcome and the
association between them separately.

\subsubsection{Circular component}

Firstly, we can compare the estimated mean of the circular outcome
between the four models. In Table 5 we see the means of the posterior
predictive distributions for the circular outcome of the CL-PN, CL-GPN
and joint GPN-SSN model. For the CL-PN model we can actually also
compute this mean from the estimates in Table 2 as follows:
\(\mu_{circ} = atan2(\beta_0^{II}, \beta_0^{I}) = 32.62^\circ\). The
estimates for the CL-PN, CL-GPN and joint GPN-SSN model are about equal
and correspond quite well to the actual data mean of 33.22\(^\circ\) in
Table \ref{Tableteacherdescriptives}. The estimate from the Abe-Ley
model which is 0.36 radians or 20.62\(^\circ\) is different. This
difference could be caused by the fact that the densities for the
circular ouctome, projected normal or sine-skewed von Mises, differ
between the models.

\begin{table}

\caption{\label{tab:means}Posterior estimates for the circular mean in the CL-PN, CL-GPN and joint GPN-SSN models}
\centering
\begin{tabular}[t]{lr}
\toprule
  & Mean (degrees)\\
\midrule
Cl-PN & 31.79\\
CL-GPN & 33.38\\
GPN-SSN & 34.47\\
\bottomrule
\multicolumn{2}{l}{\textsuperscript{a} Note that these means}\\
\multicolumn{2}{l}{are based on the}\\
\multicolumn{2}{l}{posterior predictive}\\
\multicolumn{2}{l}{distribution for the}\\
\multicolumn{2}{l}{intercepts following}\\
\multicolumn{2}{l}{(Wang \& Gelfand,}\\
\multicolumn{2}{l}{2013).}\\
\end{tabular}
\end{table}

We can also compare the effect of self-efficacy (\verb|SE|) on the
circular outcome. For the CL-PN model and Abe-Ley models we can get
estimates of a circular regression coefficient of this effect. For the
Abe-Ley model this estimate is the parameter \(\eta_1 = -0.02\). This
means that for each unit increase in self-efficacy, at the inflection
point of the circular regression line, the score of the teacher on the
IPC decreases with \(0.02*(180/\pi) = 1.15^\circ\).
\textcolor{red}{CHRISTOPHE: KLOPT DIT, OF MOET IK $\lambda$ HIER NOG MEENEMEN; NEGEER IK NU HET SINE-SKEWED GEDEELTE}
The inflection point is the point at which the regression line starts
flattening off, i.e.~the steepness of the slope decreases. For the CL-PN
model we can use methods from Cremers, Mulder, \& Klugkist (2018b). The
estimated posterior mode of \(b_c\), a parameter that is comparable to
\(\eta_1\) in the Abe-Ley model, equals \(1.53\) and its 95\% HPD
interval is \((-17.37; 21.26)\). This means that for each unit increase
in self-efficacy, at the inflection point of the circular regression
line, the score of the teacher on the IPC increases with
\(1.53*(180/\pi) = 87.66^\circ\). The values of \(b_c\) and \(\eta_1\)
are thus quite different even though they should represent the same
effect. However, the slope at the inflection point that \(b_c\) and
\(\eta_1\) describe is not necessarily representative of the effect of
\verb|SE| in the data range. In Figure \ref{regline} we see that the
inflection point (square) of the predicted circular regression line for
the CL-PN model lies just ouside of the data range and that for the
Abe-Ley model even further (outside the x-axis range). The slope of the
regression lines in the data range is much more alike even though one is
positive and the other negative. The slope at the mean self-efficicay,
which has a value of 0 because it was centered, \(SAM\) is estimated at
\(0.07 (-0.03; 0.17)\). The \(SAM\) can be interpreted as the amount of
change on the circle (in radians) for a unit increase at the mean of the
predictor variable. However, for the CL-PN model the 95\% HPD interval
of \(SAM\) for the effect of \verb|SE| includes 0 meaning that the value
is not different from 0. THis explains in part how the slopes of the two
regression lines are predicted to be different. For the Abe-Ley model
standard errors of the parameters are not known so we cannot formally
test whether \(\eta_1\) differs from 0. Neither is there a known way of
computing the effect at predictor values other than the one of the
inflection point.

For the CL-GPN and GPN-SSN we cannot compute circular regression
coefficients. Instead, we will compute posterior predictive
distributions for the predicted circular outcome of individuals scoring
the minimum, maximum and average self-efficacy. The modes and 95\% HPD
intervals of these posterior predictive distributions in the CL-GPN
model are
\(\hat{\theta}_{SE_{min}} = 211.31^\circ (144.41^\circ \: 43.43^\circ)\),
\(\hat{\theta}_{SE_{mean}} = 25.82^\circ (335.55^\circ \: 145.11^\circ)\)
and \(\hat{\theta}_{SE_{max}} =30.96^\circ (6.72^\circ \: 74.35^\circ)\)
respectively. For the GPN-SSN model the modes and 95\% HPD intervals of
the posterior predictive distributions are
\(\hat{\theta}_{SE_{min}} = 352.06^\circ (121.82^\circ \: 77.75^\circ)\),
\(\hat{\theta}_{SE_{mean}} = 22.74^\circ (336.71^\circ \: 132.25^\circ)\)
and
\(\hat{\theta}_{SE_{max}} = 30.87^\circ (358.80^\circ \: 82.13^\circ)\)
respectively. For both models we can thus conclude that as the
self-efficacy increases, the score of the teacher on the IPC moves
counterlockwise. We can however not reach any conclusions about the size
of this effect or whether it is different from zero.

\begin{figure}
\centering
\includegraphics[width = \textwidth]{Plots/reglinediffSE.pdf}
\caption{Plot showing circular regresion lines for the effect of self-efficacy as predicted by the Abe-Ley (solid line) and CL-PN model (dashed line). The black square indicated the inflection point of the circular regression line for the CL-PN model.}
\label{regline}
\end{figure}

\subsubsection{Linear component}

For the linear outcome, we compare the regression coefficients of
\verb|SE| for the linear outcome, the strength of the interpersonal
behavior of a teacher. For the CL-PN, CL-GPN and joint GPN-SSN model
this are the parameters \(\gamma_1\) and \(\beta_{1_y}\) respectively.
Note that in the CL-PN and CL-GPN model the effect of \verb|SE| is the
effect controlled for \(\cos\theta\) and \(\sin\theta\) similar to a
`normal' multiple regression model. For the GPN-SSN model the effect of
\verb|SE| is also controlled for the circular outcome. However in this
model the associaton is modelled in a way that is similar to modelling
the association between several outcome variables in a multivariate
regression or ANOVA model. Additionally, the distribution we assume for
the linear outcome is different. For the CL-PN and CL-GPN models we
assume a normal distribution while for the GPN-SSN model we assume a
skewed normal distribution. These differences will lead to differences
in the estimated effect of \verb|SE| on the linear outcome. The
interpretation of the regression coefficients is however similar for all
three models and is like the interpretation of a regression coefficient
in any other standard linear regression model.

For all three models there is an effect of \verb|SE| on the strength of
the interpersonal behavior of a teacher, the HPD interval does not
include 0: \(\gamma_1 = 0.03 \: (0.00; 0.07)\) (CL-PN),
\(\gamma_1 = 0.03 \: (0.00; 0.06)\) (CL-GPN) and
\(\beta_{1_y} = 0.08 \: (0.05; 0.12)\) (GPN-SSN). This means that
teachers with a higher self-efficacy for classroom management are
stronger in their interpersonal behavior than teachers with a lower
self-efficacy. In particular, for each unit increase in self-efficacy
the strength of their interpersonal behavior is 0.03 or 0.08 (depending
on the model) higher. This is an increase of 4.17 or 11.11 \%
considering the range of the strength on the IPC in the data.

For the Abe-Ley model we make use of the fact that the conditional
distribution for the linear outcome is Weibull. This means that we can
use methods from survival analsis in medical statistics to interpret the
effect of self-efficacy. In survival analysis they make use of a
`survival' function in which the time is plotted against the probability
of survival of subjects suffering from a specific medical condition. In
our data however we plot the strength on the IPC (instead of time)
against the probability of a teacher having such a strength. This
probability can be computed using the so-called `survival-function'
which in our case is equal to \(\exp(-\alpha y_i^{\beta})\) where
\(\beta = \exp(\nu_0 + \nu_1SE_i)\). In Figure \ref{reglineweib} we plot
the survival function for the minimum, mean and maximum value of
self-efficacy. From this Figure we conclude that the stronger
interpersonal behaviors are less probable. We also see that the
self-efficacy positively influences the strength on the IPC, the
probability of having a stronger interpersonal behavior increases with
increasing self-efficacy. This result is equivalent to the effect of
self-efficacy found using the CL-PN, CL-GPN and GPN-SSN models.

\begin{figure}
\centering
\includegraphics[width = \textwidth]{Plots/survivaldiffSE.pdf}
\caption{Plot showing the probability of having a particular strength of interersonal behavior (survival plot) for the minimum, mean and maximum self-efficacy in the data.}
\label{reglineweib}
\end{figure}

\subsubsection{Association between the linear and circular component}

For all four models we can also investigate the association between the
linear and circular component. In the CL-PN and CL-GPN model we do this
using the parameters \(\gamma_{\cos}\) and \(\gamma_{sin}\), the
coefficients of the regression of the linear outcome onto the sine and
cosine of the circular outcome. In the joint SSN-GPN model we look at
the covariances between the linear outcome and the sine and cosine of
the circular outcome \(\sum_{sy_{2,3}}\) and \(\sum_{sy_{1,3}}\). For
the Abe-Ley model we can use the model parameters in Table 3 to compute
a circular-linear correlation (Abe \& Ley, 2017).

In both the CL-PN and CL-GPN model there is an effect of the cosine of
the circular outcome onto the linear outcome (HPD interval does not
include 0). In the teacher data the sine and cosine component have a
substantive meaning. In this case the Communion component of the IPC
positively effects the strength of a teachers' type of interpersonal
behavior. Thus the teachers showing interpersonal behavior types with
higher communion scores (e.g. `helpful' and `understanding' in Figure
\ref{QTI}) are stronger in their behavior. In the joint SSN-GPN model we
reach a sligtly different conclusion. In this model both covariances,
\(\sum_{sy_{2,3}} = 0.09\) and \(\sum_{sy_{1,3}} = 0.23\), are different
from 0, but the one between the cosine of the circular outcome and the
linear outcome is larger. This means that both the Communion and Agency
component of the IPC positively effect the strength of a teachers' type
of interpersonal behavior but the effect of the Communon component is
larger. Thus, teachers scoring high on Agency but slightly higher on
Communion (the `helpful' category) are strongest in their behavior.
Following Abe \& Ley (2017) the estimated correlation between the
circular and linear outcome is equal to
\textcolor{red}{CHRISTOPHE: KAN JIJ DEZE BEREKENEN, IN R KAN IK GEEN MANIER
VINDEN OM LEGENDRE FUNCTIES MET NIET INTEGER DEGREE TE EVALUEREN EN ALS IK
http://functions.wolfram.com/webMathematica/FunctionEvaluation.jsp?name=LegendreP2General
GEBRUIKT KRIJG IK IMAGINAIRE GETALLEN (voor m = 1 en m = 2), IK WEET NIET OF DAT
DE BEDOELING IS..}. This correlation is not affected by \(\mu\) or
\(\beta\) meaning that the covariate, self-efficacy, can not influence
the correlation between the type and strength of interpersonal behavior.

\subsubsection{Model fit}

Instead of interpreting specific model parameters we can also look at
the overall fit of the models to the data. In Section \ref{Modelfit} we
introduced the criterion, PLSL, that we use to assess model fit in this
paper. This criterion is focused on the out-of-sample predictive
performance. Table 6 shows the values of this criterion for the linear
and circular outcome of the four different models fit to the teacher
data.

The CL-PN and CL-GPN model have the best out-of-sample predictive
performance for the linear outcome. Note that they have roughly the same
performance since they model the linear outcome in the same way apart
from the value of \(r\) in \eqref{circlinlink} which is a parameter
computed in the estimation of the parameters for the circular variable.
This similarity in fit is also shown in Figure \ref{preddist} which
shows histograms of the test and holdout data together with the
(posterior) predicitive distributions for all four cylindrical models.
The posterior predictive distributions for the CL-PN and CL-GPN model
are almost the same for the linear outcome.

The CL-GPN model has the best out-of-sample predicitve performance for
the circular outcome. The Abe-Ley model has the second best performance.
This means that for the circular variable a slightly skewed distribution
(see Figure \ref{preddist}) fits best. The GPN-SSN model fits worst even
though it assumes the same distribution for the circular outcome as the
CL-GPN model. This could be due to the fact that the relation between
the circular and linear outcome is modeled differently in these two
models. In the CL-GPN model the linear outcome depends on the circular
one following the relation in \eqref{circlinlink} but the circular
outcome is not dependent on the linear one. In the GPN-SSN model both
outcomes are modelled with a joint variance-covariance matrix governing
their dependence. In Figure \ref{preddist} we see however that the
actual predictive densities do not differ that much between the four
models. It is possible that the large difference in PLSL are caused by
the fact that they are computed for a relatively small holdout dataset
(n = 15) \textcolor{red}{CHRISTOPHE: IS DIT LOGISCH?}.

\begin{table}

\caption{\label{tab:ModelFit}PLSL criteria for the circular and linear outcome in the four cylindrical models}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
  & CL-PN & CL-GPN & Abe-Ley & Joint GPN-SSN\\
\midrule
circular & 86.49 & 46.83 & 59.34 & 111.63\\
linear & -20.85 & -20.94 & 51.46 & 15.48\\
\bottomrule
\end{tabular}
\end{table}

\begin{figure}
\centering
\includegraphics[width = 0.8\textwidth]{Plots/preddistdiffSE.pdf}
\caption{Histograms of linear and circular values of the test and holdout set of the teacher data plotted together with (posterior) predictive density estimates for the modified CL-PN, CL-GPN, Abe-Ley and joint GPN-SSN models based on $n \times 1000$ samples from the (posterior) predictive distribution.}
\label{preddist}
\end{figure}

\section{Discussion}\label{Discussion}

In this paper we have modified four models for cylindrical data such
that they include a regression of both the linear and circular
components onto a set of covariates. Subsequently we have used these
four methods to analyze a dataset on the interpersonal behavior of
teachers. Here we will comment on the differences between these models,
the results from the analysis of the teacher data and how cylindrical
models improve the analysis of this type of data; data from an
interpersonal circumplex.

In terms of interpretability, the CL-PN and Abe-Ley model score best. As
shown in the previous section it is relatively easy to interpret the
parameters of the linear and the circular component for these models. In
the CL-GPN and joint GPN-SSN model the interpretation of the parameters
of the circular component it not easy, if at all possible. This is
caused by the fact that in addition to the mean vector the covariance
matrix of the GPN distribution affects the location of the circular
data, making it difficult to compute regression coefficients on the
circle. Wang \& Gelfand (2013) state that we can use Monte Carlo
integration to compute a circular mean and variance for the GPN
distribution. This might also be a solution to computing coefficients on
the circle and could be applied to the methods of Cremers et al. (2018b)
such that they can be used for circular regression coefficients in GPN
models as well. \textcolor{red}{HIER OOK
NOG EEN OPMERKING OVER HOE DE ASSOCIATION TUSSEN CIRCULAR AND LINEAR COMPONENT
WORDT GEMODELLEERD}.

In terms of flexibility the joint GPN-SSN scores best. In this model
multiple linear and circular outcomes can be included and we can thus
apply it to multivariate cylindrical data. In addition for both the
linear anc circular components this model and the CL-GPN and CL-PN
models are extendable to a mixed-effects model and can thus also be fit
to longitudinal data. Extension to mixed-effects, longitudinal (see
Nuñez-Antonio \& Gutiérrez-Peña (2014) and Hernandez-Stumpfhauser et al.
(2016) for hierarchical/mixed-effects models for the PN and GPN
distributions respectively). For the Abe-Ley model this may also be
possible and has been done for the conditional distribution of its
linear component, the Weibull distribution, but has not been done in
previous literature for the conditional distribution of its circular
component (sine-skewed von
Mises)\textcolor{red}{CHRISTOPHE: KLOPT DIT?}. In addition, the joint
GPN-SSN model allows for non-symmetrical shapes of the distributions of
both the linear (sine-skewed normal) and the circular (general projected
normal) components. The Abe-Ley model also allows for this and the
CL-GPN model allows for a non-symmetrical shape of the circular
component.

To investigate model fit for the teacher data we asessed the fit to the
linear and circular component separately using the PLSL, a criterion
that is designed to assess the out-of-sample preditive performance. The
CL-PN and CL-GPN model have the best fit for the linear component and
the CL-GPN model has the best fit for the circular component.
Differences in fit, in addition to being a result of the different
distributions that were used to model the linear and circular component
of the data, may also be caused by the way in which the relation between
the linear and circular component is modeled. Whereas in both the
Abe-Ley and joint GPN-SSN model the distribution of the linear component
is conditional on the circular component and vice versa, the
distribution of the circular component in the CL-PN and CL-GPN model is
independent of the linear component. In these models the linear
component are related through a regression structure where the circular
component serves as a predictor and the linear component as the outcome.
We also investigated the (posterior) predictive distributions of each
model and its fit to the training data, the part of the teacher data the
model was fit on, and the holdout data, the part of the teacher data
that we used to assess the out-of-sample predictive performance. From
these plots we saw that whereas the PLSL criteria seem to indicate a
substantive difference in fit, the (posterior) predicitve distributions
seem to be quite similar (especially for the circular component). This
discrepancy might be caused by the fact that the holdout dataset, for
which the PLSL criteria were computed, is quite small and thus that
small deviations in fit could lead to large deviations in the criterion
value.

The four cylindrical models that were modified to the regression context
in this paper are not the only cylindrical distributions available from
the literature. Several other cylindrical distributions, amongst which
are Kato \& Shimizu (2008), Sugasawa (2015) and Fernández-Durán (2007)
have been introduced (for more references we refer to chapter 2 of Ley
\& Verdebout (2017)). In the present research we have decided not to
include these models for reasons of space, complexity of the models and
ease of implementing a regression structure. In further research however
it would be interesting to investigate other types of cylindrical
distributions as well.

Even though some of the models have downsides regarding interpretation,
e.g.~the Abe-Ley model does not provide standard errors and parameters
from the models with a GPN distribution for the circular component are
hard to interpret, cylindrical models in general offer new insights into
data of a cylindrical nature in psychology. Concerning the particular
example in this paper, the teacher data that were measured using an
interpersonal circumplex, we were able to analyze all information in the
data simultaneously. Before, the two components from the interpersonal
circumplex were analyzed separately. With such an approach we get
information about the strength of a teachers' score on the two axes of
interpersonal behavior, in our case Agency and Communion. This however
does not allow us to distinguish between different types of behavior
that are specific combinations of Agency and Communion. A solution to
that would be to treat circumplex data as circular (Cremers et al.,
2018a). This solution however, does not retain the information about the
strength of the specific type of interpersonal behavior a teacher
displays. This is possible using the cylindrical models in the present
paper, we can simultaneosly model the information about the type (a
circular variable) and strength (a linear variable) of interpersonal
behavior that a teacher displays and the association between them.

\newpage

\section*{References}

\hypertarget{refs}{}
\hypertarget{ref-abe2017tractable}{}
Abe, T., \& Ley, C. (2017). A tractable, parsimonious and flexible model
for cylindrical data, with applications. \emph{Econometrics and
Statistics}, \emph{4}, 91--104.
doi:\href{https://doi.org/10.1016/j.ecosta.2016.04.001}{10.1016/j.ecosta.2016.04.001}

\hypertarget{ref-chrastil2017rotational}{}
Chrastil, E. R., \& Warren, W. H. (2017). Rotational error in path
integration: Encoding and execution errors in angle reproduction.
\emph{Experimental Brain Research}, \emph{235}(6), 1885--1897.
doi:\href{https://doi.org/10.1007/s00221-017-4910-y}{10.1007/s00221-017-4910-y}

\hypertarget{ref-Claessens2016side}{}
Claessens, L. C. (2016). \emph{Be on my side i'll be on your side :
Teachers' perceptions of teacher--student relationships} (PhD thesis).

\hypertarget{ref-Cremers2018Assessing}{}
Cremers, J., Mainhard, M. T., \& Klugkist, I. (2018a). Assessing a
bayesian embedding approach to circular regression models.
\emph{Methodology}, \emph{14}(2), 69--81.

\hypertarget{ref-CremersMulderKlugkist2017}{}
Cremers, J., Mulder, K. T., \& Klugkist, I. (2018b). Circular
interpretation of regression coefficients. \emph{British Journal of
Mathematical and Statistical Psychology}, \emph{71}(1), 75--95.
doi:\href{https://doi.org/10.1111/bmsp.12108}{10.1111/bmsp.12108}

\hypertarget{ref-fernandez2007models}{}
Fernández-Durán, J. (2007). Models for circular--linear and
circular--circular data constructed from circular distributions based on
nonnegative trigonometric sums. \emph{Biometrics}, \emph{63}(2),
579--585.
doi:\href{https://doi.org/10.1111/j.1541-0420.2006.00716.x}{10.1111/j.1541-0420.2006.00716.x}

\hypertarget{ref-fisher1995statistical}{}
Fisher, N. I. (1995). \emph{Statistical analysis of circular data}.
Cambridge: Cambridge University Press.

\hypertarget{ref-garcia2014test}{}
García-Portugués, E., Barros, A. M., Crujeiras, R. M.,
González-Manteiga, W., \& Pereira, J. (2014). A test for
directional-linear independence, with applications to wildfire
orientation and size. \emph{Stochastic Environmental Research and Risk
Assessment}, \emph{28}(5), 1261--1275.
doi:\href{https://doi.org/10.1007/s00477-013-0819-6}{10.1007/s00477-013-0819-6}

\hypertarget{ref-garcia2013exploring}{}
García-Portugués, E., Crujeiras, R. M., \& González-Manteiga, W. (2013).
Exploring wind direction and so2 concentration by circular--linear
density estimation. \emph{Stochastic Environmental Research and Risk
Assessment}, \emph{27}(5), 1055--1067.
doi:\href{https://doi.org/10.1007/s00477-012-0642-5}{10.1007/s00477-012-0642-5}

\hypertarget{ref-BDA}{}
Gelman, A., Carlin, J., Stern, H., Dunson, D., Vehtari, A., \& Rubin, D.
(2014). \emph{Bayesian data analysis} (3rd ed.). Boca Raton, FL: Chapman
\& Hall/CRC.

\hypertarget{ref-gneiting2007strictly}{}
Gneiting, T., \& Raftery, A. E. (2007). Strictly proper scoring rules,
prediction, and estimation. \emph{Journal of the American Statistical
Association}, \emph{102}(477), 359--378.
doi:\href{https://doi.org/10.1198/016214506000001437}{10.1198/016214506000001437}

\hypertarget{ref-gurtman2009exploring}{}
Gurtman, M. B. (2009). Exploring personality with the interpersonal
circumplex. \emph{Social and Personality Psychology Compass},
\emph{3}(4), 601--619.
doi:\href{https://doi.org/10.1111/j.1751-9004.2009.00172.x}{10.1111/j.1751-9004.2009.00172.x}

\hypertarget{ref-hernandez2016general}{}
Hernandez-Stumpfhauser, D., Breidt, F. J., \& Woerd, M. J. van der.
(2016). The general projected normal distribution of arbitrary
dimension: Modeling and bayesian inference. \emph{Bayesian Analysis},
\emph{12}(1), 113--133.
doi:\href{https://doi.org/10.1214/15-BA989}{10.1214/15-BA989}

\hypertarget{ref-horowitz2010handbook}{}
Horowitz, L. M., \& Strack, S. (2011). \emph{Handbook of interpersonal
psychology: Theory, research, assessment, and therapeutic
interventions}. Hoboken, NJ: John Wiley \& Sons.

\hypertarget{ref-jammalamadaka2001topics}{}
Jammalamadaka, S. R., \& Sengupta, A. (2001). \emph{Topics in circular
statistics} (Vol. 5). World Scientific.

\hypertarget{ref-johnson1978some}{}
Johnson, R. A., \& Wehrly, T. E. (1978). Some angular-linear
distributions and related regression models. \emph{Journal of the
American Statistical Association}, \emph{73}(363), 602--606.

\hypertarget{ref-kato2008dependent}{}
Kato, S., \& Shimizu, K. (2008). Dependent models for observations which
include angular ones. \emph{Journal of Statistical Planning and
Inference}, \emph{138}(11), 3538--3549.
doi:\href{https://doi.org/10.1016/j.jspi.2006.12.009}{10.1016/j.jspi.2006.12.009}

\hypertarget{ref-lagona2015hidden}{}
Lagona, F., Picone, M., Maruotti, A., \& Cosoli, S. (2015). A hidden
markov approach to the analysis of space--time environmental data with
linear and circular components. \emph{Stochastic Environmental Research
and Risk Assessment}, \emph{29}(2), 397--409.
doi:\href{https://doi.org/10.1007/s00477-014-0919-y}{10.1007/s00477-014-0919-y}

\hypertarget{ref-ley2017modern}{}
Ley, C., \& Verdebout, T. (2017). \emph{Modern directional statistics}.
CRC Press.

\hypertarget{ref-mardia2000directional}{}
Mardia, K. V., \& Jupp, P. E. (2000). \emph{Directional statistics}
(Vol. 494). Chichester, England: Wiley.

\hypertarget{ref-mardia1978model}{}
Mardia, K. V., \& Sutton, T. W. (1978). A model for cylindrical
variables with applications. \emph{Journal of the Royal Statistical
Society. Series B (Methodological)}, \emph{40}(2), 229--233.

\hypertarget{ref-mastrantonio2018joint}{}
Mastrantonio, G. (2018). The joint projected normal and skew-normal: A
distribution for poly-cylindrical data. \emph{Journal of Multivariate
Analysis}, \emph{165}, 14--26.
doi:\href{https://doi.org/10.1016/j.jmva.2017.11.006}{10.1016/j.jmva.2017.11.006}

\hypertarget{ref-mastrantonio2015bayesian}{}
Mastrantonio, G., Maruotti, A., \& Jona-Lasinio, G. (2015). Bayesian
hidden markov modelling using circular-linear general projected normal
distribution. \emph{Environmetrics}, \emph{26}(2), 145--158.
doi:\href{https://doi.org/10.1002/env.2326}{10.1002/env.2326}

\hypertarget{ref-nunez2014bayesian}{}
Nuñez-Antonio, G., \& Gutiérrez-Peña, E. (2014). A Bayesian model for
longitudinal circular data based on the projected normal distribution.
\emph{Computational Statistics \& Data Analysis}, \emph{71}, 506--519.
doi:\href{https://doi.org/10.1016/j.csda.2012.07.025}{10.1016/j.csda.2012.07.025}

\hypertarget{ref-nunez2011bayesian}{}
Nuñez-Antonio, G., Gutiérrez-Peña, E., \& Escarela, G. (2011). A
Bayesian regression model for circular data based on the projected
normal distribution. \emph{Statistical Modelling}, \emph{11}(3),
185--201.
doi:\href{https://doi.org/10.1177/1471082X1001100301}{10.1177/1471082X1001100301}

\hypertarget{ref-olmos2012extension}{}
Olmos, N. M., Varela, H., Gómez, H. W., \& Bolfarine, H. (2012). An
extension of the half-normal distribution. \emph{Statistical Papers},
\emph{53}(4), 875--886.
doi:\href{https://doi.org/10.1007/s00362-011-0391-4}{10.1007/s00362-011-0391-4}

\hypertarget{ref-pennings2017interpersonal}{}
Pennings, H. J. M., Brekelmans, M., Sadler, P., Claessens, L. C., Want,
A. C. van der, \& Tartwijk, J. van. (2017). Interpersonal adaptation in
teacher-student interaction. \emph{Learning and Instruction}, \emph{55},
41--57.
doi:\href{https://doi.org/10.1016/j.learninstruc.2017.09.005}{10.1016/j.learninstruc.2017.09.005}

\hypertarget{ref-rayner200935th}{}
Rayner, K. (2009). The 35th sir frederick bartlett lecture: Eye
movements and attention in reading, scene perception, and visual search.
\emph{Quarterly Journal of Experimental Psychology}, \emph{62}(8),
1457--1506.
doi:\href{https://doi.org/10.1080/17470210902816461}{10.1080/17470210902816461}

\hypertarget{ref-sahu2003new}{}
Sahu, S. K., Dey, D. K., \& Branco, M. D. (2003). A new class of
multivariate skew distributions with applications to bayesian regression
models. \emph{Canadian Journal of Statistics}, \emph{31}(2), 129--150.
doi:\href{https://doi.org/10.2307/3316064}{10.2307/3316064}

\hypertarget{ref-sugasawa2015flexible}{}
Sugasawa, S., S. (2015). \emph{A flexible family of distributions on the
cylinder}. Retrieved from
\href{arXiv:\%201501.06332v2}{arXiv: 1501.06332v2}

\hypertarget{ref-wang2012directional}{}
Wang, F., \& Gelfand, A. E. (2013). Directional data analysis under the
general projected normal distribution. \emph{Statistical Methodology},
\emph{10}(1), 113--127.
doi:\href{https://doi.org/10.1016/j.stamet.2012.07.005}{10.1016/j.stamet.2012.07.005}

\hypertarget{ref-vanderWant2015role}{}
Want, A. C. van der. (2015). \emph{Teachers' interpersonal role
identity.} (PhD thesis).

\hypertarget{ref-wubbels2006interpersonal}{}
Wubbels, T., Brekelmans, M., Brok, P. den, \& Tartwijk, J. van. (2006).
An interpersonal perspective on classroom management in secondary
classrooms in the netherlands. In C. Evertson \& C. S. Weinstein (Eds.),
\emph{Handbook of classroom management: Research, practice, and
contemporary issues} (pp. 1161--1191). Malwah, NJ: Lawrence Erlbaum
Associates.

\newpage

\begin{appendices}
\section{Appendix}\label{Appendix}

In this appendix we outline the MCMC procedures to fit the cylindrical regression models from Section \ref{Models}. R-code for the MCMC sampler and the analysis of the teacher data can be found here: \url{https://github.com/joliencremers/CylindricalComparisonCircumplex}.

\subsection{Bayesian Model and MCMC procedure for the modified CL-PN model}\label{A1}

We use the following algorithm to obtain posterior estimates from the model:

\begin{enumerate}

\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_i, \dots, \theta_n$, the linear outcome $\boldsymbol{y} = y_i, \dots, y_n$ where $n$ is the sample size, and the design matrices $\boldsymbol{Z}^k$ and $\boldsymbol{X}$ for the two components of the circular and the linear outcome respectively in a training (90\%) and holdout (10\%) set.

\item Define the prior parameters for the training set. In this paper we use:

\begin{itemize}
\item Prior for $\boldsymbol{\gamma}$: $N_q(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_{0})$, with  $\boldsymbol{\mu}_{0} = (0,0,0,0)^t$ and  $\boldsymbol{\Lambda}_{0} = 10^{-4}\boldsymbol{I}_4$.
\item Prior for $\sigma^2$: $IG(\alpha_{0}, \beta_{0})$, an inverse-gamma prior with $\alpha_{0} = 0.001$ and  $\beta_{0} = 0.001$.
\item Prior for $\boldsymbol{\beta^{k}}$: $N_2(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_{0})$, with $\boldsymbol{\mu}_{0} = (0,0)^t$ and  $\boldsymbol{\Lambda}_{0} = 10^{-4}\boldsymbol{I}_2$ for $k \in I,II$.
\end{itemize}

\item Set starting values $\boldsymbol{\gamma} = (0,0,0,0)^t$, $\sigma^2 = 1$ and $\boldsymbol{\beta^{k}} = (0,0)$ for $k \in I,II$. Also set starting values $r_i = 1$ in the training and holdout set.

\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:

$$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos \theta_i \\  r_i\sin \theta_i\end{bmatrix}$$

\item Sample $\boldsymbol{\gamma}$, $\sigma^2$ and $\boldsymbol{\beta^{k}}$ for $k \in I,II$ for the training dataset from their conditional posteriors:

\begin{itemize}
\item Posterior for $\boldsymbol{\gamma}$: $N_q(\boldsymbol{\mu}_n, \sigma^2\boldsymbol{\Lambda}^{-1}_n)$, with $\boldsymbol{\mu}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)^{-1}(\boldsymbol{\Lambda}_0\boldsymbol{\mu}_0 + \boldsymbol{X}^t\boldsymbol{y})$ and $\boldsymbol{\Lambda}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)$.
\item Posterior for $\sigma^2$: $IG(\alpha_{n}, \beta_{n})$, an inverse-gamma posterior with $\alpha_{n} = \alpha_0 + n/2$ and $\beta_{n} = \beta_0 + \frac{1}{2}(\boldsymbol{y}^t\boldsymbol{y} + \boldsymbol{\mu}_{0}^t\boldsymbol{\Lambda}_0\boldsymbol{\mu}_{0} + \boldsymbol{\mu}_{n}^t\boldsymbol{\Lambda}_n\boldsymbol{\mu}_{n})$.
\item Posterior for $\boldsymbol{\beta^{k}}$: $N_2(\boldsymbol{\mu}^k_n, \boldsymbol{\Lambda}^{k}_n)$, with $\boldsymbol{\mu}^k_n = ((\boldsymbol{Z}^k)^t\boldsymbol{Z}^k + \boldsymbol{\Lambda}^k_0)^{-1}(\boldsymbol{\Lambda}^k_0\boldsymbol{\mu}^k_0 + (\boldsymbol{Z}^k)^t\boldsymbol{s}^k)$ and $\boldsymbol{\Lambda}^{k}_n = ((\boldsymbol{Z}^k)^t\boldsymbol{Z}^k + \boldsymbol{\Lambda}^k_0)$.
\end{itemize}

\item Sample new $r_i$ for the training and holdout dataset from the following posterior:

$$f(r_i \mid \theta_i, \boldsymbol{\mu}_i) \propto r_i \exp{(-\frac{1}{2}(r_i)^2 + b_ir_i)}$$ 
where $b_i = \begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}^t\boldsymbol{\mu}_i$, $\boldsymbol{\mu}_i = \boldsymbol{z_i}\boldsymbol{B}$ and $\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$.

We can sample from this posterior using a slice sampling technique (Cremers et al., 2018):

\begin{itemize}
\item In a slice sampler the joint density for an auxiliary variable $v_{i}$ with $r_{i}$ is:

$$p(r_{i}, v_{i}\mid \theta_{i}, \boldsymbol{\mu}_{i}=\boldsymbol{z}_{i}\boldsymbol{B}) \propto r_{i} \textbf{I}\left(0 < v_i < \exp\left\{ -\frac{1}{2}(r_{i} - b_{i})^2\right\}\right)\textbf{I}(r_i > 0).$$

\noindent The full conditional for $v_{i}$, $p(v_{i} \mid r_{i},\boldsymbol{\mu}_{i}, \theta_{i})$, is:

$$U\left(0, \exp\left\{-\frac{1}{2}(r_{i} -  b _{i})^2\right\}\right)$$

and the full conditional for $r_i$, $p(r_{i} \mid v_{i},\boldsymbol{\mu}_{i}, \theta_{i})$, is proportional to:
$$r_{i} \textbf{I}\left(b_{i} + \max\left\{-b_{i}, -\sqrt{-2\ln v_{i}}\right\} < r_{i} < b_{i} + \sqrt{-2\ln v_{i}}\right).$$

\noindent We thus sample $v_{i}$ from the uniform distribution specified above. Independently we sample a value $m$ from $U(0,1)$. We obtain a new value for $r_{i}$ by computing $ r_{i} = \sqrt{(r_{i_{2}}^{2}-r_{i_{1}}^{2})m + r_{i_{1}}^{2}}$ where $r_{i_{1}}=b_{i} +\max\left\{-b_{i}, -\sqrt{-2\ln v_{i}}\right\}$ and $ r_{i_{2}}= b_{i} + \sqrt{-2\ln v_{i}}$.
\end{itemize}

\item Compute the PLSL for the circular and linear outcome on the holdout set using the estimates of $\boldsymbol{\gamma}$, $\sigma^2$ and $\boldsymbol{\beta^{k}}$ for $k \in I,II$ for the training dataset.

\item Repeat steps 4 to 8 until the sampled parameter estimates have converged.

\end{enumerate}

\newpage
\subsection{Bayesian Model and MCMC procedure for the modified CL-GPN mode}\label{A2}

We use the following algorithm to obtain posterior estimates from the model:

\begin{enumerate}

\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_i, \dots, \theta_n$, the linear outcome $\boldsymbol{y} = y_i, \dots, y_n$ where $n$ is the sample size, and the design matrices $\boldsymbol{Z}^k$ and $\boldsymbol{X}$ for the two components of the circular and the linear outcome respectively in a training (90\%) and holdout (10\%) set.

\item Define the prior parameters for the training set. In this paper we use:

\begin{itemize}
\item Prior for $\boldsymbol{\gamma}$: $N_q(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_{0})$, with $\boldsymbol{\mu}_{0} = (0,0,0,0)^t$ and $\boldsymbol{\Lambda}_{0} = 10^{-4}\boldsymbol{I}_4$.
\item Prior for $\sigma^2$: $IG(\alpha_{0}, \beta_{0})$, an inverse-gamma prior with $\alpha_{0} = 0.001$ and $\beta_{0} = 0.001$.
\item Prior for $\boldsymbol{\beta}_{j}$: $N_2(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_0)$, with $\boldsymbol{\mu}_{0} = (0,0)^t$ and  $\boldsymbol{\Sigma}_{0} = 10^{5}\boldsymbol{I}_2$ for $j \in 1, \dots, p$ where $p$ is the number of covariates in $\boldsymbol{Z}$.
\item Prior for $\rho$: $N(\mu_0, \sigma^2)$, with $\mu_0 = 0$ and $\sigma^2 = 10^{4}$.
\item Prior for $\tau$: $IG(\alpha_{0}, \beta_{0})$, an inverse gamma prior with $\alpha_{0} = 0.01$ and $\beta_{0} = 0.01$.
\end{itemize}

\item Set starting values $\boldsymbol{\gamma} = (0,0,0,0)^t$, $\sigma^2 = 1$, $\boldsymbol{\beta}_j = (0,0)^t$, $\rho = 0$, $\tau = 1$ and $\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1 \end{bmatrix}$. Also set starting values $r_i = 1$ in the training and holdout set.

\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:

$$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos \theta_i \\  r_i\sin \theta_i\end{bmatrix}$$

\item Sample $\boldsymbol{\gamma}$, $\sigma^2$, $\boldsymbol{\beta}_j$, $\rho$ and $\tau$ for the training dataset from their conditional posteriors:

\begin{itemize}
\item Posterior for $\boldsymbol{\gamma}$: $N_q(\boldsymbol{\mu}_n, \sigma^2\boldsymbol{\Lambda}^{-1}_n)$, with $\boldsymbol{\mu}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)^{-1}(\boldsymbol{\Lambda}_0\boldsymbol{\mu}_0 + \boldsymbol{X}^t\boldsymbol{y})$ and $\boldsymbol{\Lambda}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)$.
\item Posterior for $\sigma^2$: $IG(\alpha_{n}, \beta_{n})$, an inverse-gamma posterior where $\alpha_{n} = \alpha_0 + n/2$ and $\beta_{n} = \beta_0 + 0.5(\boldsymbol{y}^t\boldsymbol{y} + \boldsymbol{\mu}_{0}^t\boldsymbol{\Lambda}_0\boldsymbol{\mu}_{0} + \boldsymbol{\mu}_{n}^t\boldsymbol{\Lambda}_n\boldsymbol{\mu}_{n})$.
\item Posterior for $\boldsymbol{\beta}_j$: $N_2(\boldsymbol{\mu}_{j_{n}}, \boldsymbol{\Sigma}_{j_{n}})$, with $\boldsymbol{\mu}_{j_{n}} = \boldsymbol{\Sigma}_{j_{n}}\boldsymbol{\Sigma}^{-1}\Bigg(-\sum_{i=1}^{n}z_{i,j-1}\sum_{l\neq j}z_{i,l-1}\boldsymbol{\beta}_l + \sum_{i=1}^{n}z_{i,j-1}r_i\begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}\Bigg)$ and  $\boldsymbol{\Sigma}_{j_{n}} = \Big(\sum_{i=1}^{n}z_{i,j-1}^2\boldsymbol{\Sigma}^{-1}+\boldsymbol{\Lambda}_0\Big)^{-1}$ for $j \in 1, \dots, p$ where p is the number of covariates in $\boldsymbol{Z}$.
\item Posterior for $\rho$: $N(\mu_n, \sigma^2_n)$, with $\mu_n = \frac{\tau^{-2} \sum_{i=1}^{n}(s^{I}_{i} - \mu_i^{I})(s^{II}_{i} - \mu_i^{II}) + \mu_0\sigma_0^{-2}}{\tau^{-2}\sum_{i=1}^{n}(s^{II}_{i} - \mu^{II})^2 + \sigma_0^{-2}}$ and $\sigma_n^2 = \frac{1}{\tau^{-2}\sum_{i=1}^{n}(s^{II}_{i} - \mu^{II})^2 + \sigma_0^{-2}}$ where $\mu_i^{I} = \boldsymbol{z}_i\boldsymbol{\beta}^{I}$ and $\mu_i^{II} = \boldsymbol{z}_i\boldsymbol{\beta}^{II}$.
\item Posterior for $\tau$: $IG(\alpha_n, \beta_n)$, an inverse-gamma posterior with $\alpha_n = \frac{n}{2} + \alpha_0$ and $\beta_n = \sum\limits_{i = 1}^{n}(s^{I}_{i} - \{\mu_i^{II} + \rho(s^{II}_{i} - \mu^{II})\})^2 + \beta_0$
\end{itemize}

\item Sample new $r_i$ for the training and holdout dataset from the following posterior:

$$f(r_i \mid \theta_i, \boldsymbol{\mu}_i) \propto r_i \exp{\left\{-0.5A_i\bigg(r_i-\frac{B_i}{A_i}\bigg)^2\right\}}$$ 
where $B_i = \begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}^t\boldsymbol{\Sigma}^{-1}\boldsymbol{\mu}_i$, $\boldsymbol{\mu}_i = \boldsymbol{z_i}\boldsymbol{B}$, $\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$ and $A_i = \begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}^t\boldsymbol{\Sigma}^{-1}\begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}$.

We can sample from this posterior using a slice sampling technique (Hernandez-Stumpfhauser et.al. 2018):

\begin{itemize}
\item In a slice sampler the joint density for an auxiliary variable $v_{i}$ with $r_{i}$ is:

$$p(r_{i}, v_{i}\mid \theta_{i}, \boldsymbol{\mu}_{i}=\boldsymbol{z}_{i}\boldsymbol{B}^{t}) \propto r_{i} \textbf{I}\bigg(0 < v_i < \exp\left\{ -.5 A_i(r_{i} - \frac{B_i}{A_i})^2\right\}\bigg)\textbf{I}(r_i > 0)$$

\item The full conditional for $v_{i}$, $p(v_{i} \mid r_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$, is:

$$U\Bigg(0, \exp\left\{-.5A_i\bigg(r_i -  \frac{B_{i}}{A_i}\bigg)^2\right\}\Bigg)$$
and the full conditional for $r_i$, $p(r_{i} \mid v_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$, is proportional to:
$$r_{i} \textbf{I}\left(\frac{B_i}{A_i} + \max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\} < r_{i} < \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}\right)$$
\item We thus sample $v_{i}$ from the uniform distribution specified above. Independently we sample a value $m$ from $U(0,1)$. We obtain a new value for $r_{i}$ by computing $r_{i} = \sqrt{(r_{i_{2}}^{2}-r_{i_{1}}^{2})m + r_{i_{1}}^{2}}$ where $r_{i_{1}}=\frac{B_i}{A_i} +\max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\}$ and $ r_{i_{2}}= \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}$.

\end{itemize}

\item Compute the PLSL for the circular and linear outcome on the holdoutset using the estimates of $\boldsymbol{\gamma}$, $\sigma^2$, $\boldsymbol{\beta^{k}}$, $\rho$ and $\tau$ for the training dataset. Use the density $f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma})$ for the circular outcome.

\item Repeat steps 4 to 8 until the sampled parameter estimates have converged.

\end{enumerate}

\newpage
\subsection{Bayesian Model and MCMC procedure multivariate GPN model}\label{A3}

\begin{enumerate}
\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_i, \dots, \theta_n$, the linear outcome $\boldsymbol{y} = y_i, \dots, y_n$ where $n$ is the sample size, and the design matrix $\boldsymbol{X}$ in a training (90\%) and holdout (10\%) set. Note that in this paper we have only one circular outcome and one linear outcome and the MCMC procedure outlined here is specified for this situation. It can however be generalized to a situation with multiple circular and linear outcomes without too much effort.

\item Define the prior parameters for the training set. Note that in this paper we have only one circular outcome and one linear outcome, so $p = 1$ and $q = 1$. In this paper we use the following priors:

\begin{itemize}
\item Prior for $\boldsymbol{\Sigma}$: $IW(\boldsymbol{\Psi}_0, \nu_0)$, an inverse-Wishart with $\boldsymbol{\Psi}_0 = 10^{-4}\boldsymbol{I}_{2p + q}$ and $\nu_0 = 1$.   
\item Prior for $\boldsymbol{\beta}$: $MN(\boldsymbol{\beta}_0, \boldsymbol{\Sigma}_0  \otimes \boldsymbol{\kappa}_0)$, where $\boldsymbol{\beta}$ is a vectorized $\boldsymbol{B}$, the matrix with regression coefficients, $\boldsymbol{\beta}_0 = \boldsymbol{0}_{k(2p + q)}$, $\boldsymbol{B}_0 = \boldsymbol{0}_{k \times (2p + q)}$ and $\boldsymbol{\kappa}_0 = 10^{-4}\boldsymbol{I}_k$.
\item Prior for $\lambda$: $N(\gamma_0, \omega_0)$, with $\gamma_0 = 0$ and $\omega_0 = 10000$.
\end{itemize}

\item Set starting values $\boldsymbol{\beta} = (0,0,0,0,0,0)^t$, $\boldsymbol{\Sigma} = \boldsymbol{I}_3$ and $\lambda = 0$. Also set starting values $r_i = 1$ and $d_i = 1$ in the training and holdout set.

\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:

$$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos \theta_i \\  r_i\sin \theta_i\end{bmatrix}$$
\item Compute the latent outcomes $\tilde{y}_i$ underlying the linear outcome for the holdout and training dataset as follows:

$$\tilde{y}_i = \lambda d_i $$

\item Compute $\boldsymbol{\eta}_i$ defined as follows for each individual i:

$$\boldsymbol{\eta}_i = (\boldsymbol{s}_i,y_i)^t - (\boldsymbol{0}_{2p}, \lambda d_i)$$

\item Sample $\boldsymbol{B}$, $\boldsymbol{\Sigma}$ and $\lambda$ for the training dataset from their conditional posteriors:

\begin{itemize}
\item Posterior for $\boldsymbol{B}$: $MN(\boldsymbol{B}_n, \boldsymbol{\kappa}_n, \boldsymbol{\Sigma}_n)$, with $\boldsymbol{B}_n = \boldsymbol{\kappa}_n^{-1}\boldsymbol{X}^t\boldsymbol{\eta} + \boldsymbol{\kappa}_0\boldsymbol{B}_0$ and $\boldsymbol{\kappa}_n = \boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\kappa}_0$.
\item Posterior for $\Sigma$: $IW(\boldsymbol{\Psi}_n, \nu_n)$, an inverse-Wishart with $\boldsymbol{\Psi}_n = \boldsymbol{\Psi}_0 + (\boldsymbol{\eta} - \boldsymbol{X}^t\boldsymbol{B})^t(\boldsymbol{\eta} - \boldsymbol{X}^t\boldsymbol{B}) + (\boldsymbol{B} - \boldsymbol{B}_0)^t\boldsymbol{\kappa}_0(\boldsymbol{B} - \boldsymbol{B}_0)$ and $\nu_n = \nu_0 + n$
\item Posterior for $\lambda$: $N(\gamma_n, \omega_n)$, with $\omega_n = \big(\sum_{i = 1}^{n}d_i^2\boldsymbol{\Sigma}_{y|s}^{-1} + \omega_0^{-1}\big)^{-1}$ and $\gamma_n = \omega_n \big(\sum_{i = 1}^{n}d_i\boldsymbol{\Sigma}_{y|s}^{-1}(y_i - \boldsymbol{\mu}_{y_i|s_i}) + \omega_0^{-1}\gamma_0 \big)$
\end{itemize}

\item Sample new $d_i$ for the training and holdout dataset from the following posterior:

$$f(d_i) \propto \phi_q(y_i|\boldsymbol{\mu}_{y_i|s_i} + \lambda d_i, \boldsymbol{\Sigma}_{y|s})\phi_q(d_i|0, 1),$$

where $\boldsymbol{\mu}_{y_i|s_i} = \boldsymbol{x}_i\boldsymbol{B}_{y_i|s_i}$. We can see $d_i$ as a positive regressor with $\lambda$ as covariate and $\phi_q(d_i|0, 1)$ as prior (Mastrantonio, 2018). The full conditional is then a q-dimensional truncated normal with support $\mathbb{R}^{+}$ as follows:

$$N_q(\boldsymbol{M}_{d_i}, \boldsymbol{V}_q),$$ \\where $\boldsymbol{V}_q = \big(\lambda^2\boldsymbol{\Sigma}_{y|s}^{-1} + 1\big)$ and $\boldsymbol{M}_{d_i} = \boldsymbol{V}_d\lambda\boldsymbol{\Sigma}_{y|s}^{-1}\big(y_i - \boldsymbol{\mu}_{y_i, s_i}\big)$.

\item Sample new $r_i$ for the training and holdout dataset from the following posterior:

$$f(r_i \mid \theta_i, \boldsymbol{\mu}_i) \propto r_i \exp{\left\{-0.5A_i\bigg(r_i-\frac{B_i}{A_i}\bigg)^2\right\}}$$ 
where $B_i = \begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}^t\boldsymbol{\Sigma}_{s_i|y_i}^{-1}\boldsymbol{\mu}_{s_i|y_i}$, $\boldsymbol{\mu}_{s_i|y_i} = \boldsymbol{x}_i\boldsymbol{B}_{s_i|y_i}$ and $A_i = \begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}^t\boldsymbol{\Sigma}_{s_i| y_i}^{-1}\begin{bmatrix} \cos \theta_i \\ \sin \theta_i\end{bmatrix}$. The parameters $\boldsymbol{\mu}_{s_i|y_i}$ and $\boldsymbol{\Sigma}_{s_i| y_i}$ are the conditional mean and covariance matrix of $\boldsymbol{s}_i$ assuming that $(\boldsymbol{s}_i, y_i)^t \sim N_{2p+q}(\boldsymbol{\mu} + (\boldsymbol{0}_{2p}, \lambda d_i)^t, \boldsymbol{\Sigma})$.

Because in this paper $\boldsymbol{\theta}$ originates from a bivariate variable that is known we can simply define the $r_i$ as the euclidean norm of the bivariate datapoints. However, for didactic purposes continue with the explanation of the sampling procedure. We
can sample from posterior for $r_i$ using a slice sampling technique (Hernandez-Stumpfhauser et.al. 2018):

\begin{itemize}
\item In a slice sampler the joint density for an auxiliary variable $v_{i}$ with $r_{i}$ is:

$$p(r_{i}, v_{i}\mid \theta_{i}, \boldsymbol{\mu}_{i}=\boldsymbol{x}_{i}\boldsymbol{B}^{t}) \propto r_{i} \textbf{I}\bigg(0 < v_i < \exp\left\{ -.5 A_i(r_{i} - \frac{B_i}{A_i})^2\right\}\bigg)\textbf{I}(r_i > 0)$$

\item The full conditionals for $v_{i}$, $p(v_{i} \mid r_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$, is:

$$U\Bigg(0, \exp\left\{-.5A_i\bigg(r_i -  \frac{B_{i}}{A_i}\bigg)^2\right\}\Bigg)$$
and the full conditional for $r_i$, $p(r_{i} \mid v_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$ , is proportional to :
$$r_{i} \textbf{I}\left(\frac{B_i}{A_i} + \max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\} < r_{i} < \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}\right)$$
\item We thus sample $v_{i}$ from the uniform distribution specified above. Independently we sample a value $m$ from $U(0,1)$. We obtain a new value for $r_{i}$ by computing $r_{i} = \sqrt{(r_{i_{2}}^{2}-r_{i_{1}}^{2})m + r_{i_{1}}^{2}}$ where $r_{i_{1}}=\frac{B_i}{A_i} +\max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\}$ and $ r_{i_{2}}= \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}$.

\end{itemize}

\item Compute the PLSL for the circular and linear outcome on the holdoutset using the estimates of $\boldsymbol{B}$, $\boldsymbol{\Sigma}$ and $\lambda$ for the training dataset.

\item Repeat steps 4 to 10 until the sampled parameter estimates have converged.

\item In the MCMC sampler we have estimated an unconstrained $\boldsymbol{\Sigma}$. However, for identification of the model we need to apply the following constraint to both $\boldsymbol{\Sigma}$ and $\boldsymbol{\mu}$:

$$\boldsymbol{C} = \begin{bmatrix} \boldsymbol{C}_w & \boldsymbol{0}_{2p \times q} \\ \boldsymbol{0}_{2p \times q}^t & \boldsymbol{I}_q \end{bmatrix}$$
where $\boldsymbol{C}_w$ is a $2p \times 2p$ diagonal matrix with every $(2(j-1) + k)^{th}$ entry $> 0$ where $k \in 1, 2$ and $k = 1, \dots, p$ (Mastrantonio, 2018). The estimates $\boldsymbol{\Sigma}$ and $\boldsymbol{\mu}$ can then be related to their constrained versions, $\tilde{\boldsymbol{\Sigma}}$ and $\tilde{\boldsymbol{\mu}}$ as follows:

$$\boldsymbol{\mu} = \boldsymbol{C}\tilde{\boldsymbol{\mu}}$$
$$\boldsymbol{\Sigma} = \boldsymbol{C}\tilde{\boldsymbol{\Sigma}}\boldsymbol{C}.$$

\end{enumerate}

\end{appendices}

\end{document}