### 2018-07-02 Manuscript/Plots/survivaldiffSE.tex

% Created by tikzDevice version 0.11 on 2018-07-02 13:49:30
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,252.94);
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (408.42,203.75);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50,198.47) --
	( 79.14,198.43) --
	( 95.77,198.15) --
	(112.40,197.32) --
	(129.03,195.59) --
	(145.66,192.64) --
	(162.29,188.17) --
	(178.92,181.98) --
	(195.55,173.98) --
	(212.18,164.25) --
	(228.81,153.04) --
	(245.44,140.78) --
	(262.07,128.04) --
	(278.70,115.47) --
	(295.33,103.70) --
	(311.96, 93.25) --
	(328.59, 84.49) --
	(345.22, 77.56) --
	(361.85, 72.41) --
	(378.49, 68.82) --
	(395.12, 66.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50, 61.20) -- (395.12, 61.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50, 61.20) -- ( 62.50, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (129.03, 61.20) -- (129.03, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (195.55, 61.20) -- (195.55, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.07, 61.20) -- (262.07, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (328.59, 61.20) -- (328.59, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.12, 61.20) -- (395.12, 55.20);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 62.50, 39.60) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (129.03, 39.60) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (195.55, 39.60) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (262.07, 39.60) {0.6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (328.59, 39.60) {0.8};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (395.12, 39.60) {1.0};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 63.51) -- ( 49.20,198.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 63.51) -- ( 43.20, 63.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 90.50) -- ( 43.20, 90.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,117.49) -- ( 43.20,117.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,144.48) -- ( 43.20,144.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,171.47) -- ( 43.20,171.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,198.47) -- ( 43.20,198.47);

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80, 63.51) {0.0};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80, 90.50) {0.2};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,117.49) {0.4};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,144.48) {0.6};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,171.47) {0.8};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,198.47) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (228.81, 15.60) {strength IPC};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 10.80,132.47) {P(strength IPC)};
\end{scope}
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (408.42,203.75);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 62.50,198.47) --
	( 79.14,198.44) --
	( 95.77,198.21) --
	(112.40,197.48) --
	(129.03,195.93) --
	(145.66,193.23) --
	(162.29,189.08) --
	(178.92,183.22) --
	(195.55,175.55) --
	(212.18,166.08) --
	(228.81,155.03) --
	(245.44,142.81) --
	(262.07,129.97) --
	(278.70,117.17) --
	(295.33,105.08) --
	(311.96, 94.28) --
	(328.59, 85.19) --
	(345.22, 77.97) --
	(361.85, 72.61) --
	(378.49, 68.89) --
	(395.12, 66.48);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 1pt off 3pt ,line join=round,line cap=round] ( 62.50,198.47) --
	( 79.14,198.44) --
	( 95.77,198.23) --
	(112.40,197.56) --
	(129.03,196.11) --
	(145.66,193.55) --
	(162.29,189.57) --
	(178.92,183.91) --
	(195.55,176.42) --
	(212.18,167.10) --
	(228.81,156.16) --
	(245.44,143.97) --
	(262.07,131.08) --
	(278.70,118.15) --
	(295.33,105.89) --
	(311.96, 94.89) --
	(328.59, 85.60) --
	(345.22, 78.22) --
	(361.85, 72.73) --
	(378.49, 68.93) --
	(395.12, 66.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.33,186.47) -- (322.33,186.47);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (304.33,174.47) -- (322.33,174.47);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 1pt off 3pt ,line join=round,line cap=round] (304.33,162.47) -- (322.33,162.47);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (331.33,183.02) {min(SE)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (331.33,171.02) {mean(SE)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (331.33,159.02) {max(SE)};
\end{scope}
\end{tikzpicture}

\end{document}