### 2018-09-24 Manuscript/Plots/survivaldiffSE.tex

% Created by tikzDevice version 0.11 on 2018-09-18 14:43:22
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (433.62,252.94);
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (408.42,203.75);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50,198.47) --
	( 79.14,198.38) --
	( 95.77,197.84) --
	(112.40,196.49) --
	(129.03,194.01) --
	(145.66,190.14) --
	(162.29,184.70) --
	(178.92,177.62) --
	(195.55,168.93) --
	(212.18,158.83) --
	(228.81,147.63) --
	(245.44,135.76) --
	(262.07,123.72) --
	(278.70,112.05) --
	(295.33,101.23) --
	(311.96, 91.68) --
	(328.59, 83.63) --
	(345.22, 77.20) --
	(361.85, 72.33) --
	(378.49, 68.84) --
	(395.12, 66.48);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50, 61.20) -- (395.12, 61.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 62.50, 61.20) -- ( 62.50, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (129.03, 61.20) -- (129.03, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (195.55, 61.20) -- (195.55, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (262.07, 61.20) -- (262.07, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (328.59, 61.20) -- (328.59, 55.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (395.12, 61.20) -- (395.12, 55.20);

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 62.50, 39.60) {0.0};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (129.03, 39.60) {0.2};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (195.55, 39.60) {0.4};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (262.07, 39.60) {0.6};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (328.59, 39.60) {0.8};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (395.12, 39.60) {1.0};

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 62.99) -- ( 49.20,198.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 62.99) -- ( 43.20, 62.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20, 90.08) -- ( 43.20, 90.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,117.18) -- ( 43.20,117.18);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,144.27) -- ( 43.20,144.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,171.37) -- ( 43.20,171.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.20,198.47) -- ( 43.20,198.47);

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80, 62.99) {0.0};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80, 90.08) {0.2};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,117.18) {0.4};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,144.27) {0.6};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,171.37) {0.8};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 34.80,198.47) {1.0};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (433.62,252.94);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at (228.81, 15.60) {strength IPC};

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  1.00] at ( 10.80,132.47) {P(strength IPC)};
\end{scope}
\begin{scope}
\path[clip] ( 49.20, 61.20) rectangle (408.42,203.75);
\definecolor{drawColor}{RGB}{0,0,0}

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] ( 62.50,198.47) --
	( 79.14,196.04) --
	( 95.77,190.46) --
	(112.40,182.63) --
	(129.03,173.19) --
	(145.66,162.73) --
	(162.29,151.74) --
	(178.92,140.68) --
	(195.55,129.91) --
	(212.18,119.75) --
	(228.81,110.39) --
	(245.44,101.97) --
	(262.07, 94.57) --
	(278.70, 88.20) --
	(295.33, 82.82) --
	(311.96, 78.36) --
	(328.59, 74.74) --
	(345.22, 71.84) --
	(361.85, 69.57) --
	(378.49, 67.81) --
	(395.12, 66.48);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 1pt off 3pt ,line join=round,line cap=round] ( 62.50,198.47) --
	( 79.14,198.08) --
	( 95.77,196.53) --
	(112.40,193.46) --
	(129.03,188.72) --
	(145.66,182.27) --
	(162.29,174.23) --
	(178.92,164.82) --
	(195.55,154.36) --
	(212.18,143.26) --
	(228.81,131.93) --
	(245.44,120.83) --
	(262.07,110.32) --
	(278.70,100.74) --
	(295.33, 92.29) --
	(311.96, 85.12) --
	(328.59, 79.23) --
	(345.22, 74.56) --
	(361.85, 70.98) --
	(378.49, 68.35) --
	(395.12, 66.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (304.33,186.47) -- (322.33,186.47);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 4pt off 4pt ,line join=round,line cap=round] (304.33,174.47) -- (322.33,174.47);

\path[draw=drawColor,line width= 0.4pt,dash pattern=on 1pt off 3pt ,line join=round,line cap=round] (304.33,162.47) -- (322.33,162.47);

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (331.33,183.02) {min(SE)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (331.33,171.02) {median(SE)};

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  1.00] at (331.33,159.02) {max(SE)};
\end{scope}
\end{tikzpicture}

\end{document}