### 2018-09-24 Manuscript/ManuscriptMarkdownCV.Rmd
Manuscript update

---
title: "Regression models for Cylindrical data in Psychology"
author: #|
 # | Jolien Cremers^[Corresponding author: j.cremers@uu.nl] $^1$, Helena J.M. Pennings$^{2,3}$,  Christophe Ley $^{4}$
 # | $^1$Department of Methodology and Statistics, Utrecht University
 # | $^2$TNO
  #| $^3$Department of Education, Utrecht University
  #| $^4$Department of Applied Mathematics, Computer Science and Statistics, Ghent University

bibliography: CircularData.bib
csl: apa.csl
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: yes
  html_document: default
  word_document: default
fontsize: 12pt
geometry: margin=1in
header-includes:
  - \usepackage{multirow}
  - \usepackage{appendix}
  - \usepackage{color}
  - \usepackage{hyperref}
  - \usepackage{subcaption}
  - \setlength\parindent{24pt}
  - \usepackage[nodisplayskipstretch]{setspace}\doublespacing
  #- \usepackage{endfloat}
  - \DeclareRobustCommand{\VANDER}[3]{#2} % set up for citation (in .tex %  place \DeclareRobustCommand{\VANDER}[3]{#3} before bibliography)

keywords: bla vla
documentclass: article
pandoc_args: --natbib
abstract: |
  Cylindrical data are multivariate data which consist of a directional,
  in this paper circular, and a linear component. Examples of cylindrical
  data in psychology include human navigation (direction and distance of 
  movement), eye-tracking research (direction and length of saccades) and
  data from an interpersonal circumplex (type and strength of
  interpersonal behavior). In this paper we adapt four models for
  cylindrical data to include a regression of the circular and linear
  component onto a set of covariates. Subsequently, we illustrate how to
  fit these models and interpret their results on a dataset on the
  interpersonal behavior of teachers.
---
```{r inlude = FALSE, echo = FALSE}

knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

```{r, echo = FALSE}
library(circular)
library(bpnreg)
library(haven)
library(tikzDevice)
library(plotrix)
library(tidyr)
library(dplyr) #sample_n()
library(MASS) #mvrnorm()
library(kableExtra) #latex tables
library(plotrix)
library(caret) #cross-validation
```

\section{Introduction}\label{Introduction}

Cylindrical data are data that consist of a linear variable and a directional
variable. In this paper, the directional variable is circular, meaning that it
consists of a single angle instead of a set of angles. A circular variable is
different from a linear variable in the sense that it is measured on a different
scale. Figure \ref{circline} shows the difference between a circular scale
(right) and a linear scale (left). The most important difference is that on a
circular scale the datapoints 0\(^\circ\) and 360\(^\circ\) are connected and in
fact represent the same number while on a linear scale the two ends, \(-\infty\)
and \(\infty\), are not connected and consequently the values 0\(^\circ\) and
360\(^\circ\) are located on different places on the scale. This difference
requires us to use special statistical methods for circular variables (see e.g.
@fisher1995statistical for an introduction to circular data and
@mardia2000directional, @jammalamadaka2001topics and @ley2017modern for a more
  elaborate overview). For instance, a notion as simple as the sample average
  needs to be re-defined. As is the case for circular data, the analysis of
  cylindrical data also requires special methods.\newline
\indent Cylindrical data occur in several fields of research, such as for
instance in meteorology [@garcia2013exploring], ecology [@garcia2014test] or
marine research [@lagona2015hidden]. However, also in psychology several types
of cylindrical data can be found. For example, in research on human navigation
in the field of cognitive psychology both the distance, a linear variable, and
the direction, a circular variable, of movement are of interest
[@chrastil2017rotational]. In eye-tracking research, saccade data are an example
of cylindrical data, because both the direction (\emph{i.e.}, the circular variable)
and the duration (\emph{i.e.}, the linear variable) of the saccades are of interest
(for a review of eye-tracking research see  @rayner200935th).\newline
\indent The type of data that is used in the present study are also
psychological, namely data from circumplex measurement instruments. For
instance, data from the interpersonal circumplex as used in personality
psychology are by definition of a cylindrical nature (see Section \ref{Example}
for a more detailed explanation).\newline
\indent In the literature, several methods have been put forward to model the
relation between the linear and circular component of a cylindrical variable.
Some of these are based on regressing the linear component onto the circular
component using the following type of relation: \[y = \beta_0 +
\beta_1*\cos(\theta) + \beta_2*\sin(\theta)+ \epsilon,\] where \(y\) is the
linear component and \(\theta\) the circular component [@mardia1978model;
@johnson1978some; @mastrantonio2015bayesian]. Others model the relation in a
different way, e.g. by specifying a multivariate model for several linear and
circular variables and modelling their covariance matrix
[@mastrantonio2018joint] or by proposing a joint cylindrical distribution. For
example, @abe2017tractable introduce a cylindrical distribution based on a
Weibull distribution for the linear component and a sine-skewed von Mises
distribution for the circular component and link these through their respective
shape and concentration parameters. However, none of the methods that have been
proposed thus far include additional covariates onto which both the circular and
linear component are regressed.\newline
\indent Our aim in this paper is to fill this gap in the literature by adapting
four existing cylindrical models in such a way that they include a regression of
both the linear and circular component of a cylindrical variable onto a set of
covariates. From now on we will therefore refer to the components of the
cylindrical variable as outcome components. Additionally, we will show how a
correct statistical treatment of such cylindrical data can lead to new insights.
We will do this for the teacher data, a dataset from the field of educational
psychology. In the teacher data, apart from modelling the relation between the
linear and circular component of a cylindrical variable we would also like to
predict the two components from a set of covariates in a regression model.

The paper is organized as follows: Section \ref{Example} describes the teacher
data, while Section \ref{Models} presents the four cylindrical models and our
associated adaptations to new regression models. In that same section we also
discuss the model fit criterion that we will use in Section \ref{DataAnalysis}
for the comparison of the four models. A detailed data analysis with interesting
new insights is also provided in Section \ref{DataAnalysis}. We conclude the
paper with a discussion in Section \ref{Discussion}, and the Supplementary Material collects
the technical details of the MCMC procedures.

\begin{figure}
\centering
\includegraphics[width = 0.8\textwidth]{Plots/circline.pdf}
\caption{The difference between a linear scale (left) and a circular scale (right).}
\label{circline}
\end{figure}

\section{Teacher data}\label{Example}

\begin{figure}
\centering
\includegraphics[width = 0.8\textwidth]{Plots/IPC-T.png}
\caption{The interpersonal circle for teachers (IPC-T). The words presented in
the circumference of the circle are anchor words to describe the type of
behavior located in each part of the IPC.}
\label{QTI}
\end{figure}

The motivating example for this article comes from the field of educational
psychology and was collected for the studies on classroom climate of
@vanderWant2015role, @Claessens2016side and @pennings2018interpersonal. An
indicator of the quality of the classroom climate is the students' perception of
their teachers' interpersonal behavior. These interpersonal perceptions, both in
educational psychology as well as in other areas of psychology, can be measured
using circumplex measurement instruments (see @horowitz2010handbook for an
overview of many such instruments).\newline
\indent The circumplex data used in this paper are measured using the
Questionnaire on Teacher Interaction (QTI) [@wubbels2006interpersonal] which is
one such circumplex measurement instrument. The QTI is designed to measure
student perceptions of their teachers' interpersonal behavior and contains items
that load on two interpersonal dimensions: Agency and Communion. Agency refers
to the degree of power or control a teacher exerts in interaction with his/her
students. Communion refers to the degree of friendliness or affiliation a
teacher conveys in interaction with his/her students. The loadings on the two
dimensions of the QTI can be placed in a two-dimensional space formed by Agency
(vertical) and Communion (horizontal), see Figure \ref{QTI}. Different parts of
this space are characterized by different teacher behavior, e.g. 'helpful' or
'uncertain'. This two-dimensional space is called the interpersonal
circle/circumplex (IPC). The IPC is ``a continuous order with no beginning or
end'' [@gurtman2009exploring, p. 2]. We call such ordering a circumplex ordering
and the IPC is therefore often called the interpersonal circumplex. The ordering
also implies that scores on the IPC could be viewed as a circular
variable.\newline
\indent @Cremers2018Assessing explain the circular nature of the IPC data and
analyze them as such using a circular regression model. The two-dimension scores
Agency and Communion can be converted to a circular score using the two-argument
arctangent function in \eqref{PredVal}, where \(A\) represents a score on the
Agency dimension and \(C\) represents a score on the Communion dimension
\begin{equation}\label{PredVal}
\theta          = \text{atan2}\left(A, \: C\right)  =
\left\{{\begin{array}{lcl}
                                                                       \arctan\left(\frac{A}{C}\right) & \text{if}  \quad&C > 0 \\
\arctan\left(\frac{A}{C}\right) + \pi & \text{if}  \quad& C  <  0  \:\: \&\:\: A \geq 0\\
 \arctan\left(\frac{A}{C}\right) - \pi & \text{if}  \quad&C  <  0 \:\:  \&\:\:A  < 0\\
 +\frac{\pi}{2} & \text{if}  \quad& C  =  0  \:\: \&\:\:A > 0\\
 -\frac{\pi}{2} & \text{if}  \quad& C =  0  \:\: \&\:\:A < 0\\
 \text{undefined} & \text{if} \quad& C =  0   \:\: \&\:\:A = 0.
 \end{array}}
\right.
\end{equation}
The resulting circular variable \(\theta\) can then be modelled and takes values
in the interval $[0, 2\pi)$. However, when two-dimensional data are converted
to the circle we lose some information, namely the length of the two-dimensional
vector \((A, C)^t\), \emph{i.e.}, its Euclidean norm \(\mid\mid (A, C)^t \mid\mid\).
This length represents the strength of the type of interpersonal behavior a
teacher shows towards his/her students and can be considered as the linear
variable in a cylindrical model, allowing us to model a circular variable
\(\theta\) together with the linear variable corresponding to \(\mid\mid (A, C)^t
\mid\mid\). This leads to an improved analysis of interpersonal circumplex data
as we take all information into account. In the next section we introduce
several models that can be used for a more accurate and informative regression
analysis on the teacher data. First however we will provide descriptives for our
data set.

\subsection{Data description}\label{DataDescriptives}

```{r loaddata, return = FALSE, echo = FALSE}
Dat <- read_spss("Data Heleen/mergeddata.sav")

#only select the first measurement occasion
Dat <- subset(Dat, Time == 0)

#sort on class size
Dat <- Dat[order(-Dat$nklas), ]

#Now remove duplicates 
#(because dataframe is sorted on class size the smallest classes are removed)

duplicates <- duplicated(Dat[,c("docnr", "Time")])
Dat <- Dat[!duplicates,]

#center and remove missings for self-efficacy
Dat <- Dat %>% drop_na(Efficacy_CM)
Dat$SEc <- Dat$Efficacy_CM- mean(Dat$Efficacy_CM)
Dat$theta <- atan2(Dat$lAgency, Dat$lcommunion)
Dat$y <- sqrt(Dat$lAgency^2 + Dat$lcommunion^2)

set.seed(30)

#Create holdout and training samples for cross-validation

vec <-  seq(1:nrow(Dat))
flds <- createFolds(vec, k = 10, list = TRUE, returnTrain = FALSE)

for(i in 1:10) {

 #Create name for holdout data 
 nam.h <- paste("hold.", i, sep = "")

 #Get holdout dataset for fold i
 assign(nam.h, Dat[flds[[i]],])

 #Create name for training data
 nam.tr <- paste("train.", i, sep = "")

 #Get training dataset for fold i
 assign(nam.tr, Dat[-flds[[i]],])

}

```

```{r dataplot, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}

tikz("Plots/dataplot.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")

plot(Dat$y, Dat$theta*(180/pi), 
     ylab = "IPC location", xlab = "IPC strength",
     bty = "n", ylim = c(-180, 180))

dev.off()
tools::texi2dvi("Plots/dataplot.tex", pdf=TRUE)

```

```{r descripitves, echo = FALSE, results = FALSE}
#sample size
nrow(Dat)

#summary statistics
summary(Dat$Efficacy_CM)
summary(Dat$y)
summary(as.circular(Dat$theta%%(2*pi)))

sd(Dat$Efficacy_CM)
sd(Dat$y)

```

The teacher data was collected between 2010 and 2015 and contains several
repeated measures on the IPC of 161 teachers. Measurements were obtained using
the QTI and taken in different years and classes. For this paper we only
consider one measurement, the first occasion (2010) and largest class if data
for multiple classes were available. In addition to the score on the IPC, the
circular outcome, and the strength of the score on the IPC, the linear outcome,
a teachers' self-efficacy (\verb|SE|) concerning classroom management is used as
covariate in the analysis. After listwise deletion of missings ($3$ in total,
only for the self-efficacy) we have a sample of 148 teachers. Table
\ref{Tableteacherdescriptives} shows descriptives for the dataset. Here
$\hat{\rho}$ is a sample estimate for the circular concentration where a value
of 0 means that the data is not concentrated at all, \emph{i.e.} spread over the
entire circle, and a value of 1 means that all data is concentrated at a single
point on the circle. Figure \ref{dataplot} is a scatterplot showing the relation
between the linear and circular outcome of the teacher data.

\begin{figure}
\centering
\includegraphics[width = \textwidth]{Plots/dataplot.pdf}
\caption{Plot showing the relation between the linear and circular outcome component (in degrees) of the teacher data.}
\label{dataplot}
\end{figure}

\begin{table}[h]
\centering
\caption{Descriptives for the teacher dataset.} 
\begin{tabular}{lrrrl}
  \noalign{\smallskip}\hline\noalign{\smallskip}
Variable & mean/$\bar{\theta}$ & sd/$\hat{\rho}$ & Range & Type \\ \hline\noalign{\smallskip}
IPC &33.22$^\circ$& 0.76 & - & Circular\\
strength IPC & 0.43 & 0.15 & 0.08 - 0.80 & Linear\\
SE & 5.04 & 1.00 & 1.5 - 7.0 & Linear\\
   \hline
\end{tabular}
\label{Tableteacherdescriptives}
\end{table}

\section{Four cylindrical regression models}\label{Models}

In this section we present four cylindrical models and adapt them such
that they contain predictors for the linear and circular outcomes, \(Y\)
and \(\Theta\). The first two models are based on a construction by
@mastrantonio2015bayesian, while the other models are extensions of the
models from @abe2017tractable and @mastrantonio2018joint.

\subsection{The modified CL-PN and modified CL-GPN  models}\label{CL-(G)PN}

Following @mastrantonio2015bayesian we consider in this section two models where
the relation between \(\Theta \in [0, 2\pi)\) and \(Y\in (-\infty, + \infty)\)
and $q$ covariates is specified as
\begin{equation}\label{circlinlink}
Y = \gamma_0 + \gamma_{cos}*\cos(\Theta)*R + \gamma_{sin}*\sin(\Theta)*R + \gamma_1*x_1 + \dots + \gamma_q*x_q +  \epsilon,
\end{equation}
\noindent where the random variable \(R\geq0\) will be introduced below, the
error term \(\epsilon \sim N(0, \sigma^2)\) with variance \(\sigma^2>0\),
\(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are the
intercept and regression coefficients and \(x_1, \dots, x_q\) are the \(q\)
covariates. In both of these models the
conditional distribution of \(Y\) given \(\Theta=\theta\) and \(R = r\) is given
by
\begin{equation}\label{ycondtheta}
f(y \mid \theta, r) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp\left[-\frac{(y - (\gamma_0 + \gamma_1x_1 + \dots + \gamma_qx_q+c))^{2}}{2\sigma^2}\right],\nonumber
\end{equation}
\noindent where \(c = \begin{bmatrix} r \cos(\theta) \\ r\sin(\theta)
\end{bmatrix}^t \begin{bmatrix} \gamma_{cos} \\ \gamma_{sin} \end{bmatrix}\),
\(r \geq 0\). The linear outcome thus has a normal distribution conditional on
\(\Theta\) and \(R\) and contains already linear covariates \(x_1, \dots, x_q\)
in its location part. For the teacher dataset, the regression equation for the
linear outcome in the CL-PN and CL-GPN model is the following: \[\hat{y}_i =
\gamma_0 + \gamma_{cos}\cos(\theta_i)r_i + \gamma_{sin}\sin(\theta_i)r_i +
\gamma_1\text{SE}_i,\]
\noindent where $\text{SE}_i$ is the self-efficacy score of one individual $i = 1, \dots, n$ where $n$ is the sample size.\newline
\indent For the circular outcome we assume either a projected normal (PN) or a general
projected normal (GPN) distribution. These distributions arise from the radial
projection of a distribution defined on the plane onto the circle. The relation
between a bivariate vector \(\boldsymbol{S}\) in the plane and the circular
outcome \(\Theta\) is defined as follows
\begin{equation}\label{projection}
\boldsymbol{S} = \begin{bmatrix} S^{I} \\ S^{II} \end{bmatrix} = R\boldsymbol{u} = \begin{bmatrix} R \cos (\Theta) \\  R\sin (\Theta) \end{bmatrix},
\end{equation}
\noindent where \(R = \mid\mid \boldsymbol{S} \mid\mid\), the Euclidean norm of
the bivariate vector \(\boldsymbol{S}\). In the PN distribution we assume
\(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{I})\) and in the GPN we
assume \(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{\Sigma})\) where
$\boldsymbol{\mu} \in \mathbb{R}^2$, \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \rho^2 & \rho\\ \rho & 1
\end{bmatrix}\), \(\rho \in (-\infty, +\infty)\) and \(\tau^2 \geq 0\) (as in
@hernandez2016general). This leads to the
circular-linear PN (CL-PN) and circular-linear GPN (CL-GPN) distributions. We
will now detail how we modify both cylindrical distributions to also incorporate
covariates for the circular part.

\subsubsection{The modified CL-PN distribution}

Following @nunez2011bayesian, the joint density of \(\Theta\) and \(R\) for the
PN distribution equals
\begin{equation}\label{pnreg}
f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = \frac{r}{2\pi} \exp\left[- \frac{(r\boldsymbol{u} - \boldsymbol{\mu})^t(r\boldsymbol{u} - \boldsymbol{\mu})}{2}\right],
\end{equation}
\noindent where \(\boldsymbol{u}= \begin{bmatrix} \cos (\theta) \\ \sin (\theta)
\end{bmatrix}\) and $r$ is the same as in \eqref{circlinlink} and
\eqref{ycondtheta} and is defined in \eqref{projection}. In a regression setup
the outcomes \(\theta_i,r_i\) for each individual \(i = 1, \dots, n\), where
\(n\) is the sample size, are generated independently from the distribution with
density \eqref{pnreg}. The mean vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\)
is then defined as \(\boldsymbol{\mu}_i = \boldsymbol{B}^t\boldsymbol{z}_i\)
where the vector \(\boldsymbol{z}_i\) is a vector of dimension $p + 1$ that
contains the covariate values and the value 1 to estimate an intercept and
\(\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})\) contains
the regression coefficients and intercepts. Note however that the dimensions of
\(\boldsymbol{\beta}^{I}\) and \(\boldsymbol{\beta }^{II}\) need not necessarily
be the same and we are thus allowed to have a different set of predictor
variables and vectors \(\boldsymbol{z}_i^I\) and \(\boldsymbol{z}_i^{II}\) for
the two components of \(\boldsymbol{\mu}_i\). For the teacher dataset, the
regression equation for the circular outcome in the CL-PN model is
\[\hat{\boldsymbol{\mu}}_{i} = \begin{pmatrix} \mu_{i}^{I}  \\ \mu_{i}^{II}
\end{pmatrix}=\begin{pmatrix} \beta_0^{I} + \beta_1^{I}\text{SE}_i  \\
\beta_0^{II} + \beta_1^{II}\text{SE}_i \end{pmatrix}.\]

\subsubsection{The modified CL-GPN distribution}

Following @wang2012directional and @hernandez2016general the joint density of
\(R\) and \(\Theta\) for the GPN distribution equals
\begin{equation}\label{gpnreg}
f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma}) = \frac{r}{2\pi\tau} \exp\left[ -\frac{(r\boldsymbol{u}-\boldsymbol{\mu})^{t}\boldsymbol{\Sigma}^{-1}(r\boldsymbol{u}-\boldsymbol{\mu})}{2}\right],
\end{equation}
\noindent where we recall that \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 +
\rho^2 & \rho\\ \rho & 1 \end{bmatrix}\). In a regression setup the outcomes
\(\theta_i\) and \(r_i\) for each individual are generated independently from
\eqref{gpnreg}. The mean vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is
defined in the same way via covariates as for the modified CL-PN distribution.
Note in contrast with the CL-PN model where \(\boldsymbol{z}_i^I\) and
\(\boldsymbol{z}_i^{II}\) are allowed to differ we do need to have the same
predictors for both components of \(\boldsymbol{\mu}_i\) in the CL-GPN model.
This is due to the fact that the variance-covariance matrix
\(\boldsymbol{\Sigma}\) is no longer identity. For the teacher dataset, the
regression equation for the circular outcome in the CL-GPN model is the same as
in the CL-PN model.

\subsubsection{Parameter estimation}

Both cylindrical models introduced here are estimated using Markov Chain Monte
Carlo (MCMC) methods based on @nunez2011bayesian, @wang2012directional and
@hernandez2016general for the regression of the circular outcome. A detailed
description of the Bayesian estimation and MCMC samplers can be found in the
Supplementary Material.

\subsection{The modified Abe-Ley model}\label{WeiSSVM}

This model is an extension of the cylindrical model introduced in
@abe2017tractable to the regression context. The joint density of \(\Theta\) and
\(Y\), in this model defined only on the positive real half-line \([0, +
\infty)\), reads
\begin{equation}\label{WeiSSVMdensity}
f(\theta, y) = \frac{\alpha\beta^\alpha}{2\pi\cosh(\kappa)}
                 (1 +\lambda\sin(\theta - \mu))
                 y^{\alpha-1}
                 \exp[-(\beta y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu))],
\end{equation}
\noindent where \(\alpha > 0\) is a linear shape parameter, \(\kappa > 0\) and
\(\lambda \in [-1, 1]\) are circular concentration and skewness parameters with
\(\kappa\) also regulating the circular-linear dependence. Our modification
occurs at the level of the linear scale parameter \(\beta>0\) and circular
location parameter \(\mu\in [0, 2\pi)\), both of which we express in terms of
covariates: \(\beta_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\nu}) > 0\) and
\(\mu_i = \eta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta})\). The
parameter \(\boldsymbol{\nu}\) is a vector of \(q\) regression coefficients
\(\nu_j \in (-\infty, +\infty)\) for the prediction of \(y\) where \(j = 0,
\dots, q\) and \(\nu_0\) is the intercept. The parameter \(\eta_0 \in [0,
2\pi)\) is the intercept and \(\boldsymbol{\eta}\) is a vector of \(p\)
regression coefficients \(\eta_j \in (-\infty, +\infty)\) for the prediction of
\(\theta\) where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a
vector of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
is a vector of predictor values for the prediction of \(\theta\). In a
regression setup the outcome vector \((\theta_i, y_i)^t\) for each individual is
generated independently from the modified
density \eqref{WeiSSVMdensity}.\newline
\indent As in @abe2017tractable, the conditional distribution of \(Y\) given
\(\Theta=\theta\) is a Weibull distribution with shape \(\alpha\) and scale
\(\beta(1-\tanh(\kappa)\cos(\theta - \mu))^{1/\alpha}\) and the conditional
distribution of \(\Theta\) given \(Y=y\) is a sine skewed von Mises distribution
with location parameter \(\mu\) and concentration parameter \((\beta
y)^\alpha\tanh(\kappa)\). The log-likelihood for this model equals
\begin{align}\label{WeiSSVMLikelihood}
l(\alpha, \boldsymbol{\nu}, \lambda, \kappa, \boldsymbol{\eta}) 
   &= n[\ln(\alpha) - \ln(2\pi\cosh(\kappa))] + \alpha \sum^{n}_{i = 1} \boldsymbol{x}_i^t\boldsymbol{\nu} \nonumber\\
   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - \eta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}))) 
   +(\alpha-1)\sum^{n}_{i = 1} \ln(y_i) \nonumber\\
   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\nu})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - \eta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\eta}))).\nonumber
\end{align}
\noindent For the teacher data, $\boldsymbol{z} = \boldsymbol{x}$ and the regression equations for the circular and
linear outcomes in the Abe-Ley model are: \[\hat{\mu}_{i} = \eta_0 + 2 *
\tan^{-1}(\eta_1\text{SE}_i),\] and \[\hat{\beta}_{i} = \exp(\nu_0 +
\nu_1\text{SE}_i).\]
\noindent We can use numerical optimization (Nelder-Mead) to find solutions for
the maximum likelihood (ML) estimates for the parameters of the model.

\subsection{Modified joint projected and skew normal (GPN-SSN)}\label{CL-GPN_multivariate}

This model is an extension of the cylindrical model introduced by
@mastrantonio2018joint to the regression context. Both models contain \(m\) independent circular
  outcomes and \(w\) independent linear outcomes. The circular outcomes
  \(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,
  \boldsymbol{\Theta}_m)\) are modelled together by a multivariate GPN
  distribution. The joint distribution of $\boldsymbol{\Theta}$ and $\boldsymbol{R}$ can thus be modeled as the product of (\ref{gpnreg}) for each of the $m$ circular outcomes. The linear outcomes \(\boldsymbol{Y} = (\boldsymbol{Y}_1,
  \dots, \boldsymbol{Y}_w)\) are modelled together by a multivariate skew normal
  distribution [@sahu2003new]. Because the GPN distribution is modelled using a
  so-called augmented representation (as in \eqref{projection} and
  \eqref{gpnreg}) it is convenient to use a similar tactic for modelling the
  multivariate skew normal distribution. Following @mastrantonio2018joint the
  linear outcomes are represented as
\[\boldsymbol{Y} = \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{D} + \boldsymbol{H},\]
\noindent where \(\boldsymbol{\mu}_y\) is a mean vector for the linear outcome
\(\boldsymbol{Y}\), \(\boldsymbol{\Lambda} = \text{diag}(\boldsymbol{\lambda})\)
is a \(w \times w\) diagonal matrix with diagonal elements
\(\lambda_1, \dots, \lambda_w\) (skewness parameters),
\(\boldsymbol{D} \sim HN_w(\boldsymbol{0}_w, \boldsymbol{I}_w)\), a
$w$-dimensional half normal distribution [@olmos2012extension], and
\(\boldsymbol{H} \sim N_w(\boldsymbol{0}_w, \boldsymbol{\Sigma}_y)\). This means
that, conditional on the auxiliary data \(\boldsymbol{D}\), \(\boldsymbol{Y}\)
is normally distributed with mean \(\boldsymbol{\mu}_y +
\boldsymbol{\Lambda}\boldsymbol{D}\) and covariance matrix
\(\boldsymbol{\Sigma}_y\). The joint density for \((\boldsymbol{Y}^t,
\boldsymbol{D}^t)^t\) is defined as:
\begin{equation}\label{YDjoint}
f(\boldsymbol{y}, \boldsymbol{d}) = 2^w\phi_w(\boldsymbol{y} \mid \boldsymbol{\mu}_y + \boldsymbol{\Lambda}\boldsymbol{d}, \boldsymbol{\Sigma}_y) \phi_w(\boldsymbol{d} \mid \boldsymbol{0}_w, \boldsymbol{I}_w),\nonumber
\end{equation}
\noindent where
\(\phi_\ell(\cdot|\boldsymbol{\mu}_\ell,\boldsymbol{\Sigma}_\ell)\) stands for
the \(\ell\)-dimensional normal density with mean vector
\(\boldsymbol{\mu}_\ell\) and
covariance \(\boldsymbol{\Sigma}_\ell\). As in @mastrantonio2018joint dependence between the linear and circular outcome is created by
modelling the augmented representations of \(\boldsymbol{\Theta}\) and
\(\boldsymbol{Y}\) together in a \(2m + w\) dimensional normal
distribution. The joint density of the model is then represented by:
\begin{equation}\label{YDThetarjoint} 
f(\boldsymbol{\theta}, \boldsymbol{r},
\boldsymbol{y}, \boldsymbol{d}) = 2^w\phi_{2m+w}((\boldsymbol{s}^t,
\boldsymbol{y}^t)^t \mid \boldsymbol{\mu} + (\boldsymbol{0}_{2m}^t, ({\rm
diag}(\boldsymbol{\lambda})\boldsymbol{d})^t)^t, \boldsymbol{\Sigma})
\phi_w(\boldsymbol{d} \mid \boldsymbol{0}_w, \boldsymbol{I}_w) \prod_{j =
1}^{m}r_j, 
\end{equation}
\noindent where \(\boldsymbol{s} = (r_1(\cos(\theta_1), \sin(\theta_1)), \dots,
r_m(\cos(\theta_m), \sin(\theta_m)))^t\), the mean vector \(\boldsymbol{\mu} =
(\boldsymbol{\mu}_s^t, \boldsymbol{\mu}_y^t)^t\) and \(\boldsymbol{\Sigma} =
\left ( \begin{matrix} \boldsymbol{\Sigma}_s & \boldsymbol{\Sigma}_{sy} \\
\boldsymbol{\Sigma}_{sy}^t & \boldsymbol{\Sigma}_y \\ \end{matrix} \right )\).
The matrix \(\boldsymbol{\Sigma}_s\) is the covariance matrix for the variances
of and covariances between the augmented representations of the circular outcome
and the matrix \(\boldsymbol{\Sigma}_{sy}\) contains covariances between the
augmented representations of the circular outcome and the linear outcome.
\newline
\indent In our regression extension we have \(i = 1, \dots, n\) observations of
\(m\) circular outcomes, \(w\) linear outcomes and \(g\) covariates. The mean in
the density in \eqref{YDThetarjoint} then becomes \(\boldsymbol{\mu}_i =
\boldsymbol{B}^t\boldsymbol{x}_i\) where \(\boldsymbol{B}\) is a \((g + 1)
\times (2m + w)\) matrix with regression coefficients and intercepts and $\boldsymbol{x}_i$
is a $g + 1$ dimensional vector containing the value 1 to estimate an intercept
and the $g$ covariate values. \newline
\indent For the teacher data, the regression equations for
the circular and linear outcomes in the GPN-SSN model are
\[\hat{\boldsymbol{\mu}}_{i} = \boldsymbol{\beta}_0 +
\boldsymbol{\beta}_1\text{SE}_i,\] where
\(\hat{\boldsymbol{\mu}}_i = (\hat{\boldsymbol{\mu}}_{s_i}^t, \hat{\mu}_{y_i})^t\),
\(\boldsymbol{\beta}_0 = (\beta_{0_{s^{I}}}, \beta_{0_{s^{II}}},\beta_{0_y})^t\)
and
\(\boldsymbol{\beta}_1 = (\beta_{1_{s^{I}}}, \beta_{1_{s^{II}}},\beta_{1_y})^t\). Note that because $m = 1$ and $w = 1$, $\boldsymbol{\mu}_{s_i}$ is a 2 dimensional vector and $\mu_{y_i}$ is a scalar.
We estimate the model using MCMC methods. A detailed description of these
methods is given in the Supplementary Material.

\subsection{Model fit criterion}\label{Modelfit}

For the four cylindrical models we focus on their out-of-sample predictive
performance to determine the fit of the model. To do so we use k-fold
cross-validation and split our data into 10 folds. Each of these folds (10 \(\%\)
of the sample) is used once as a holdout set and 9 times as part of a training
set. The analysis will thus be performed 10 times, each time on a different
training set.\newline
\indent A proper criterion to compare out-of-sample predictive performance is
the Predictive Log Scoring Loss (PLSL) [@gneiting2007strictly]. The lower the
value of this criterion, the better the predictive performance of the model.
Using ML estimates this criterion can be computed as follows:
\begin{equation}\label{PLSLML}
PLSL = -2 \sum_{i = 1}^{M}\log l(x_i \mid \hat{\boldsymbol{\vartheta}}),\nonumber
\end{equation}
\noindent where \(l\) is the model likelihood, \(M\) is the sample size of the
holdout set, \(x_i\) is the \(i^{th}\) datapoint from the holdout set and
\(\hat{\boldsymbol{\vartheta}}\) are the ML estimates of the model parameters.
Using posterior samples the criterion is similar to the log pointwise predictive
density (lppd) [@BDA, p. 169] and can be computed as:
\begin{equation}\label{PLSLBayes}
PLSL = -2 \frac{1}{B} \sum_{j = 1}^{B}\sum_{i = 1}^{M} \log l(x_i \mid \boldsymbol{\vartheta}^{(j)}),\nonumber
\end{equation}
\noindent where \(B\) is the amount of posterior samples and
\(\boldsymbol{\vartheta}^{(j)}\) are the posterior estimates of the model
parameters for the \(j^{th}\) iteration. Because the joint density and thus also
the likelihood for the modified GPN-SSN model in (\ref{YDThetarjoint}) is not
available in closed form [@mastrantonio2018joint] we compute the PLSL for the
circular and linear outcome separately for all models. Note that although we fit
the CL-PN, CL-GPN and GPN-SSN models using Bayesian statistics, we do not take
prior information into account when assessing model fit with the PLSL. According
to @BDA this is not necessary since we are assessing the fit of a model
to data, the holdout set, only. They argue that the prior in such case is only
of interest for estimating the parameters of the model but not for determining
the predictive accuracy.\newline
\indent We use the loglikelihoods of the following conditional densities for the
computation of the PLSL in the teacher data:

\begin{itemize}
\item For the modified CL-PN model:

$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{pnreg}.

\item For the modified CL-GPN model:

$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{gpnreg}.

\item For the modified Abe-Ley model:

$y_i \mid \theta_i, \beta_i, \mu_i, \kappa, \alpha \sim W\left(\beta_i(1-\tanh(\kappa)\cos(\theta_i - \mu_i))^{1/\alpha}, \alpha\right)$, a Weibull distribution.

$\theta_i \mid y_i, \beta_i, \mu_i, \kappa, \alpha, \lambda \sim SSVM\left(\mu_i, (\beta_iy_i)^{\alpha}(\tanh{\kappa})\right)$, a sine-skewed von Mises distribution.

\item For the modified joint projected and skew normal model:

$y_i \mid \boldsymbol{\mu}_i, \boldsymbol{\Sigma}, \theta_i, r_i \sim SSN(\mu_{i_y} + \lambda d_i + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_s^{-1}(\boldsymbol{s}_i - \boldsymbol{\mu}_{i_s}), \sigma^2_y + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_s^{-1}\boldsymbol{\Sigma}_{sy}),$

$\theta_i \mid \boldsymbol{\mu}_i, \boldsymbol{\Sigma}, y_i, d_i \sim GPN(\boldsymbol{\mu}_{i_s} + \boldsymbol{\Sigma}_{sy}\sigma^{-2}_y(y_i - \mu_{i_y} - \lambda d_i), \boldsymbol{\Sigma}_s + \boldsymbol{\Sigma}_{sy}\sigma_y^{-2}\boldsymbol{\Sigma}_{sy}^t)$

where $SSN$ is the skew normal distribution.
\end{itemize}

\noindent For each of the four cylindrical models and for each of the 10
cross-validation analyses we can then compute a PLSL for the circular and linear
outcome by using the conditional log-likelihoods of the respective outcome. To
evaluate the predictive performance we average across the PLSL criteria of the
cross-validation analyses. We also assess the cross-validation variability by
means of the standard deviations of the PLSL criteria.

\section{Data Analysis}\label{DataAnalysis}

In this section we analyze the teacher data with the help of the four
cylindrical models from Section \ref{Models}. We will present the results,
posterior estimates and their interpretation, per model and finish with a
section comparing the fit of the different models.

```{r FitCLPN, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}

source("R-code/Posterior Sampling CL-PN.R")

its <- 20000
set.seed(101)

res_CLPN <- list()

for(i in 1:10){

  train <- get(paste("train.", i, sep = ""))
  hold <- get(paste("hold.", i, sep = ""))

  ZI.PN.tr  <- as.matrix(cbind(rep(1, length(train$theta)), train$SEc))
  ZII.PN.tr <- as.matrix(cbind(rep(1, length(train$theta)), train$SEc))
  X.PN.tr   <- as.matrix(cbind(rep(1, length(train$theta)),
                                 cos(train$theta),
                                 sin(train$theta),
                                 train$SEc))

  ZI.PN.h  <- as.matrix(cbind(rep(1, length(hold$theta)), hold$SEc))
  ZII.PN.h <- as.matrix(cbind(rep(1, length(hold$theta)), hold$SEc))
  X.PN.h   <- as.matrix(cbind(rep(1, length(hold$theta)),
                               cos(hold$theta),
                               sin(hold$theta),
                               hold$SEc))

  res_CLPN[[i]] <- CLPN(train$theta, train$y, X.PN.tr, ZI.PN.tr, ZII.PN.tr, its,
                        hold$theta, hold$y, X.PN.h, ZI.PN.h, ZII.PN.h)

}

```

```{r FitCLGPN, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}

source("R-code/Posterior Sampling CL-GPN.R")

its <- 20000
set.seed(101)

res_CLGPN <- list()

for(i in 1:10){

  train <- get(paste("train.", i, sep = ""))
  hold  <- get(paste("hold.", i, sep = ""))

  Z.GPN.tr <- as.matrix(cbind(rep(1, length(train$theta)), train$SEc))
  X.GPN.tr <- as.matrix(cbind(rep(1, length(train$theta)),
                              cos(train$theta),
                              sin(train$theta),
                              train$SEc))

  Z.GPN.h <- as.matrix(cbind(rep(1, length(hold$theta)), hold$SEc))
  X.GPN.h <- as.matrix(cbind(rep(1, length(hold$theta)),
                             cos(hold$theta),
                             sin(hold$theta),
                             hold$SEc))

  res_CLGPN[[i]] <- CLGPN(train$theta, train$y, X.GPN.tr, Z.GPN.tr, its, p = 2, 
                          hold$theta, hold$y, X.GPN.h, Z.GPN.h)

}

```

```{r FitCLGPNM, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}

source("R-code/Posterior Sampling Joint GPN-SSN.R")

its <- 20000
set.seed(101)

res_CLGPNM <- list()

for(i in 1:10){

  train <- get(paste("train.", i, sep = ""))
  hold  <- get(paste("hold.", i, sep = ""))

  X.MGPN.tr <- as.matrix(cbind(rep(1, length(train$theta)), train$SEc))

  X.MGPN.h <- as.matrix(cbind(rep(1, length(hold$theta)), hold$SEc))

  res_CLGPNM[[i]] <- JGPNSSN(train$theta, train$y, X.MGPN.tr, its, p = 1, q = 1, 
                             hold$theta, hold$y, X.MGPN.h)

}

```

```{r FitAbeLey, cache = TRUE, results = FALSE, echo = FALSE, eval = TRUE}

source("R-code/Abe-Ley optimization.R")

resAL <- list()

set.seed(101)

for(i in 1:10){

  train <- get(paste("train.", i, sep = ""))
  hold <- get(paste("hold.", i, sep = ""))

  Z.tr    <- as.matrix(cbind(rep(1, length(train$theta)), train$SEc))
  X.tr    <- as.matrix(cbind(rep(1, length(train$theta)), train$SEc))

  Z.h     <- as.matrix(cbind(rep(1, length(hold$theta)), hold$SEc))
  X.h     <- as.matrix(cbind(rep(1, length(hold$theta)), hold$SEc))

  theta.tr <- as.numeric(train$theta)
  y.tr     <- as.numeric(train$y)
  theta.h  <- as.numeric(hold$theta)
  y.h      <- as.numeric(hold$y)

  colnames(X.tr) <- c("ax", "bx")
  colnames(Z.tr) <- c("az", "bz")
  colnames(X.h) <- c("ax", "bx")
  colnames(Z.h) <- c("az", "bz")

  dat.t <- as.data.frame(cbind(theta.tr, y.tr, X.tr, Z.tr))
  dat.h <- as.data.frame(cbind(theta.h, y.h, X.h, Z.h))

  ui.reg = rbind(c(1, 0, 0, 0, 0, 0, 0),
                 c(-1, 0, 0, 0, 0, 0, 0),
                 c(0, 0, 0, 0, 1, 0, 0),
                 c(0, 0, 0, 0, 0, 1, 0),
                 c(0, 0, 0, 0, 0, 0, 1),
                 c(0, 0, 0, 0, 0, 0, -1))

  ci.reg = c(-pi, -pi, 0, 0, -1, -1)

  param <- c(1,1,1,1,1,1,0)*0.9

  ui.reg %*% param - ci.reg

  resAL[[i]] <- constrOptim(param, func.regII, ui = ui.reg, ci = ci.reg, method = "Nelder-Mead",
                            control = list(maxit = 1000000, fnscale = -1), data = dat.t)

  resAL[[i]]$ll.circ <- -2*func.reg.cond.circ(resAL[[i]]$par, dat.h)
  resAL[[i]]$ll.lin  <- -2*func.reg.cond.lin(resAL[[i]]$par, dat.h)

}

```

\subsection{Results \& Analysis}\label{DataResults}

In the Supplementary Material we have described the starting values for the MCMC
procedures, hence it remains to specify  the starting  values for the
maximum likelihood based Abe-Ley model: \(\eta_0 = 0.9, \eta_1 = 0.9, \nu_0 =
0.9, \nu_1 = 0.9, \kappa = 0.9, \alpha = 0.9, \lambda = 0\). The initial amount
of iterations for the three MCMC samplers was set to 2000. After convergence
checks via traceplots we concluded that some of the parameters of the GPN-SSN
model did not converge. Therefore we set the amount of iterations of the MCMC
models to 20,000 and subtracted a burn-in of 5000 to reach convergence. Note
that we choose the same amount of iterations for all three Bayesian models to
make their comparison via the PLSL as fair as possible. Lastly, the predictor
\verb|SE| was centered before inclusion in the analysis as this allows the
intercepts to bear the classical meaning of average behavior.

\subsubsection{The modified CL-PN and CL-GPN models} \label{resCL(G)PN}

```{r estCLPN, cache = TRUE, echo = FALSE}

#Get 3 dimensional arrays (iteration, parameter, fold) for each parameter type
Gamma <- simplify2array(lapply(res_CLPN, "[[", "Gamma"))
BI    <- simplify2array(lapply(res_CLPN, "[[", "BI"))
BII   <- simplify2array(lapply(res_CLPN, "[[", "BII"))
Sigma <- simplify2array(lapply(res_CLPN, "[[", "Sigma"))

#Averge over the folds and compute the posterior modes and hpd intervals
CLPN.Gamma.m   <- rowMeans(apply(Gamma[5000:20000,,], 2:3, mode_est))
CLPN.Gamma.hpd <- apply(apply(Gamma[5000:20000,,], 2:3, hpd_est), c(1,2), mean)
CLPN.Gamma.m.sd   <- apply(apply(Gamma[5000:20000,,], 2:3, mode_est),1, sd)
CLPN.Gamma.hpd.sd <- apply(apply(Gamma[5000:20000,,], 2:3, hpd_est), c(1,2), sd)

CLPN.BI.m   <- rowMeans(apply(BI[5000:20000,,], 2:3, mode_est))
CLPN.BI.hpd <- apply(apply(BI[5000:20000,,], 2:3, hpd_est), c(1,2), mean)
CLPN.BI.m.sd   <- apply(apply(BI[5000:20000,,], 2:3, mode_est), 1, sd)
CLPN.BI.hpd.sd <- apply(apply(BI[5000:20000,,], 2:3, hpd_est), c(1,2), sd)

CLPN.BII.m   <- rowMeans(apply(BII[5000:20000,,], 2:3, mode_est))
CLPN.BII.hpd <- apply(apply(BII[5000:20000,,], 2:3, hpd_est), c(1,2), mean)
CLPN.BII.m.sd   <- apply(apply(BII[5000:20000,,], 2:3, mode_est), 1, sd)
CLPN.BII.hpd.sd <- apply(apply(BII[5000:20000,,], 2:3, hpd_est), c(1,2), sd)

CLPN.Sigma.m   <- mean(apply(Sigma[5000:20000,], 2, mode_est))
CLPN.Sigma.hpd <- rowMeans(apply(Sigma[5000:20000,], 2, hpd_est))
CLPN.Sigma.m.sd  <- sd(apply(Sigma[5000:20000,], 2, mode_est))
CLPN.Sigma.hpd.sd <- apply(apply(Sigma[5000:20000,], 2, hpd_est), 1, sd)

#Compute a cross-validation variance/standard deviation

#Put estimates in correct format for tables
modes.PN <- c(CLPN.BI.m, CLPN.BII.m, CLPN.Gamma.m, CLPN.Sigma.m, NA, NA, NA)
hpd.LB.PN <- c(CLPN.BI.hpd[1,], CLPN.BII.hpd[1,], CLPN.Gamma.hpd[1,], CLPN.Sigma.hpd[1], NA, NA, NA)
hpd.UB.PN <- c(CLPN.BI.hpd[2,], CLPN.BII.hpd[2,], CLPN.Gamma.hpd[2,], CLPN.Sigma.hpd[2], NA, NA, NA)

modes.PN.sd <- c(CLPN.BI.m.sd, CLPN.BII.m.sd, CLPN.Gamma.m.sd, CLPN.Sigma.m.sd, NA, NA, NA)
hpd.LB.PN.sd <- c(CLPN.BI.hpd.sd[1,], CLPN.BII.hpd.sd[1,], CLPN.Gamma.hpd.sd[1,], CLPN.Sigma.hpd.sd[1], NA, NA, NA)
hpd.UB.PN.sd <- c(CLPN.BI.hpd.sd[2,], CLPN.BII.hpd.sd[2,], CLPN.Gamma.hpd.sd[2,], CLPN.Sigma.hpd.sd[2], NA, NA, NA)

CLPNres.tab <- cbind(modes.PN, modes.PN.sd, hpd.LB.PN, hpd.LB.PN.sd, hpd.UB.PN, hpd.UB.PN.sd)
CLPNres.tab = CLPNres.tab

CLPNres.tab.new = cbind(apply(CLPNres.tab[,1:2], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                       sprintf("%.2f", round(x[2], 2)), ")")),
                        apply(CLPNres.tab[,3:4], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (",
                                                                       sprintf("%.2f", round(x[2], 2)), ")")),
                        apply(CLPNres.tab[,5:6], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (",
                                                                       sprintf("%.2f", round(x[2], 2)), ")")))

```

```{r estCLGPN, cache = TRUE, echo = FALSE}

#Get 3 or 4 dimensional arrays (iteration, parameter, fold)/(,,itertation, fold) for each parameter type
Gamma <- simplify2array(lapply(res_CLGPN, "[[", "Gamma"))
BI    <- simplify2array(lapply(res_CLGPN, "[[", "BI"))
BII   <- simplify2array(lapply(res_CLGPN, "[[", "BII"))
Sig   <- simplify2array(lapply(res_CLGPN, "[[", "Sig"))
Sigma <- simplify2array(lapply(res_CLGPN, "[[", "Sigma"))

#Averge over the folds and compute the posterior modes and hpd intervals
CLGPN.Gamma.m   <- rowMeans(apply(Gamma[5000:20000,,], 2:3, mode_est))
CLGPN.Gamma.m.sd   <- apply(apply(Gamma[5000:20000,,], 2:3, mode_est), 1, sd)
CLGPN.Gamma.hpd <- apply(apply(Gamma[5000:20000,,], 2:3, hpd_est), c(1,2), mean)
CLGPN.Gamma.hpd.sd <- apply(apply(Gamma[5000:20000,,], 2:3, hpd_est), c(1,2), sd)

CLGPN.BI.m   <- rowMeans(apply(BI[5000:20000,,], 2:3, mode_est))
CLGPN.BI.m.sd   <- apply(apply(BI[5000:20000,,], 2:3, mode_est), 1, sd)
CLGPN.BI.hpd <- apply(apply(BI[5000:20000,,], 2:3, hpd_est), c(1,2), mean)
CLGPN.BI.hpd.sd <- apply(apply(BI[5000:20000,,], 2:3, hpd_est), c(1,2), sd)

CLGPN.BII.m   <- rowMeans(apply(BII[5000:20000,,], 2:3, mode_est))
CLGPN.BII.m.sd   <- apply(apply(BII[5000:20000,,], 2:3, mode_est), 1, sd)
CLGPN.BII.hpd <- apply(apply(BII[5000:20000,,], 2:3, hpd_est), c(1,2), mean)
CLGPN.BII.hpd.sd <- apply(apply(BII[5000:20000,,], 2:3, hpd_est), c(1,2), sd)

CLGPN.Sig.m   <- mean(apply(Sig[5000:20000,], 2, mode_est))
CLGPN.Sig.m.sd   <- sd(apply(Sig[5000:20000,], 2, mode_est))
CLGPN.Sig.hpd <- rowMeans(apply(Sig[5000:20000,], 2, hpd_est))
CLGPN.Sig.hpd.sd <- apply(apply(Sig[5000:20000,], 2, hpd_est), 1, sd)

CLGPN.Sigma11.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,1]
CLGPN.Sigma11.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,1]
CLGPN.Sigma12.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,2]
CLGPN.Sigma12.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,1]
CLGPN.Sigma22.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,2]
CLGPN.Sigma22.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,2]

CLGPN.Sigma11.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,1]
CLGPN.Sigma11.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,1]
CLGPN.Sigma12.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,2]
CLGPN.Sigma12.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,1]
CLGPN.Sigma22.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,2]
CLGPN.Sigma22.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,2]

#Compute a cross-validation variance/standard deviation

#Put estimates in correct format for tables
modes.GPN  <- c(CLGPN.BI.m, CLGPN.BII.m,
                CLGPN.Gamma.m, CLGPN.Sig.m,
                CLGPN.Sigma11.m, CLGPN.Sigma12.m, CLGPN.Sigma22.m)
hpd.LB.GPN <- c(CLGPN.BI.hpd[1,], CLGPN.BII.hpd[1,],
               CLGPN.Gamma.hpd[1,], CLGPN.Sig.hpd[1],
               CLGPN.Sigma11.hpd[1], CLGPN.Sigma12.hpd[1], CLGPN.Sigma22.hpd[1])
hpd.UB.GPN <- c(CLGPN.BI.hpd[2,], CLGPN.BII.hpd[2,], 
               CLGPN.Gamma.hpd[2,], CLGPN.Sig.hpd[2],
               CLGPN.Sigma11.hpd[2], CLGPN.Sigma12.hpd[2], CLGPN.Sigma22.hpd[2])

modes.GPN.sd  <- c(CLGPN.BI.m.sd, CLGPN.BII.m.sd,
                   CLGPN.Gamma.m.sd, CLGPN.Sig.m.sd,
                   CLGPN.Sigma11.m.sd, CLGPN.Sigma12.m.sd, CLGPN.Sigma22.m.sd)
hpd.LB.GPN.sd <- c(CLGPN.BI.hpd.sd[1,], CLGPN.BII.hpd.sd[1,],
                   CLGPN.Gamma.hpd.sd[1,], CLGPN.Sig.hpd.sd[1],
                   CLGPN.Sigma11.hpd.sd[1], CLGPN.Sigma12.hpd.sd[1], CLGPN.Sigma22.hpd.sd[1])
hpd.UB.GPN.sd <- c(CLGPN.BI.hpd.sd[2,], CLGPN.BII.hpd.sd[2,], 
                   CLGPN.Gamma.hpd.sd[2,], CLGPN.Sig.hpd.sd[2],
                   CLGPN.Sigma11.hpd.sd[2], CLGPN.Sigma12.hpd.sd[2], CLGPN.Sigma22.hpd.sd[2])

CLGPNres.tab <- cbind(modes.GPN, modes.GPN.sd, hpd.LB.GPN, hpd.LB.GPN.sd, hpd.UB.GPN, hpd.UB.GPN.sd)

CLGPNres.tab.new = cbind(apply(CLGPNres.tab[,1:2], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                         sprintf("%.2f", round(x[2], 2)), ")")),
                         apply(CLGPNres.tab[,3:4], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                         sprintf("%.2f", round(x[2], 2)), ")")),
                         apply(CLGPNres.tab[,5:6], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                         sprintf("%.2f", round(x[2], 2)), ")")))

CLPNGPNres.tab <- cbind(CLPNres.tab.new, CLGPNres.tab.new)

rownames(CLPNGPNres.tab) <- c("$\\beta_0^{I}$", "$\\beta_1^{I}$",
                              "$\\beta_0^{II}$", "$\\beta_1^{II}$",
                              "$\\gamma_0$", "$\\gamma_{cos}$",
                              "$\\gamma_{sin}$", "$\\gamma_1$",
                              "$\\sigma$", "$\\sum_{1,1}$",
                              "$\\sum_{1,2}$", "$\\sum_{2,2}$")
colnames(CLPNGPNres.tab) <- rep(c('Mode', "HPD LB", "HPD UB"), 2)

#Create table
#kable(CLPNGPNres.tab, "latex", booktabs = TRUE, escape = FALSE, 
#      caption = "Results, cross-validation mean and standard deviation, for the modified CL-PN and CL-GPN model")%>%
#add_header_above(c("Parameter" = 1, "CL-PN" = 3, "CL-GPN" = 3))

```

```{r means, cache = TRUE, echo = FALSE}

theta_pred_PN   <- simplify2array(lapply(res_CLPN, "[[", "theta_pred"))
theta_pred_GPN  <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred"))
theta_pred_GPNM <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred"))

theta_pred_PN <- array(unlist(theta_pred_PN),
                       dim = c(nrow(theta_pred_PN[[1]]), ncol(theta_pred_PN[[1]]), length(theta_pred_PN)))
theta_pred_GPN <- array(unlist(theta_pred_GPN),
                        dim = c(nrow(theta_pred_GPN[[1]]), ncol(theta_pred_GPN[[1]]), length(theta_pred_GPN)))
theta_pred_GPNM <- array(unlist(theta_pred_GPNM),
                         dim = c(nrow(theta_pred_GPNM[[1]]), ncol(theta_pred_GPNM[[1]]), length(theta_pred_GPNM)))

means <- round(rbind(mean_circ(apply(apply(theta_pred_PN[,5000:20000,], c(2,3), mean_circ), 2, mode_est_circ)),
                     mean_circ(apply(apply(theta_pred_GPN[,5000:20000,], c(2,3), mean_circ), 2, mode_est_circ)),
                     mean_circ(apply(apply(theta_pred_GPNM[,5000:20000,], c(2,3), mean_circ), 2, mode_est_circ)))*(180/pi), 2)
hpds  <- round(rbind(apply(apply(apply(theta_pred_PN[,5000:20000,], c(2,3), mean_circ), 2, hpd_est_circ), 1, mean_circ),
                     apply(apply(apply(theta_pred_GPN[,5000:20000,], c(2,3), mean_circ), 2, hpd_est_circ), 1, mean_circ),
                     apply(apply(apply(theta_pred_GPNM[,5000:20000,], c(2,3), mean_circ), 2, hpd_est_circ), 1, mean_circ))*(180/pi), 2)

means = cbind(means, hpds)

rownames(means) <- c("Cl-PN", "CL-GPN", "GPN-SSN")
colnames(means) <- c('Mode', "HPD LB", "HPD UB")
#kable(means, "latex", booktabs = T, escape = FALSE, digits = 2,
#      caption = "Posterior estimates for the circular mean in the CL-PN, CL-GPN and GPN-SSN models")%>%
#kable_styling() %>%
#add_footnote("Note that these means are based on the posterior predictive distribution for the intercepts following (Wang & Gelfand, 2013).", #notation="alphabet")

```

\begin{table}

\caption{\label{tab:estCLGPN}Results, cross-validation mean and standard deviation, for the modified CL-PN and CL-GPN models}
\centering
\begin{tabular}[t]{lllllll}
\toprule
\multicolumn{1}{c}{Parameter} & \multicolumn{3}{c}{CL-PN} & \multicolumn{3}{c}{CL-GPN} \\
\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-4} \cmidrule(l{2pt}r{2pt}){5-7}
  & Mode & HPD LB & HPD UB & Mode & HPD LB & HPD UB\\
\midrule
$\beta_0^{I}$ & 1.76 (0.09) & 1.50 (0.07) & 0.09 (0.09) & 2.42 (0.15) & 1.90 (0.10) & 3.05 (0.17)\\
$\beta_1^{I}$ & 0.65 (0.07) & 0.42 (0.06) & 0.08 (0.08) & 0.85 (0.12) & 0.45 (0.09) & 1.30 (0.15)\\
$\beta_0^{II}$ & 1.15 (0.05) & 0.92 (0.04) & 0.04 (0.04) & 1.47 (0.05) & 1.17 (0.04) & 1.79 (0.05)\\
$\beta_1^{II}$ & 0.58 (0.03) & 0.38 (0.04) & 0.04 (0.04) & 0.70 (0.08) & 0.46 (0.05) & 0.96 (0.07)\\
$\gamma_0$ & 0.38 (0.01) & 0.31 (0.01) & 0.01 (0.01) & 0.37 (0.01) & 0.31 (0.01) & 0.42 (0.01)\\
\addlinespace
$\gamma_{cos}$ & 0.04 (0.00) & 0.01 (0.00) & 0.00 (0.00) & 0.03 (0.00) & 0.01 (0.00) & 0.05 (0.00)\\
$\gamma_{sin}$ & -0.01 (0.00) & -0.04 (0.00) & 0.00 (0.00) & -0.00 (0.00) & -0.03 (0.00) & 0.03 (0.00)\\
$\gamma_1$ & 0.03 (0.01) & -0.00 (0.00) & 0.07 (0.01) & 0.03 (0.01) & -0.00 (0.00) & 0.06 (0.00)\\
$\sigma$ & 0.14 (0.00) & 0.12 (0.00) & 0.00 (0.00) & 0.14 (0.00) & 0.12 (0.00) & 0.16 (0.00)\\
$\sum_{1,1}$ & NA (NA) & NA (NA) & NA (NA) & 3.02 (0.25) & 1.83 (0.14) & 5.02 (0.41)\\
\addlinespace
$\sum_{1,2}$ & NA (NA) & NA (NA) & NA (NA) & 0.46 (0.12) & 0.12 (0.12) & 0.80 (0.10)\\
$\sum_{2,2}$ & NA (NA) & NA (NA) & NA (NA) & 1.00 (0.00) & 1.00 (0.00) & 1.00 (0.00)\\
\bottomrule
\end{tabular}
\end{table}

\begin{table}
\caption{\label{tab:means}Posterior estimates (in degrees) for the circular mean in the CL-PN, CL-GPN and GPN-SSN models}
\centering
\begin{tabular}[t]{lrrr}
\toprule
  & Mode & HPD LB & HPD UB\\
  \midrule
Cl-PN & 32.29 & 24.81 & 39.71\\
CL-GPN & 33.70 & 26.72 & 41.15\\
GPN-SSN & 35.30 & 28.31 & 43.10\\
\bottomrule
\multicolumn{4}{l}{\textsuperscript{a} Note that these means are based on}\\
\multicolumn{4}{l}{the posterior predictive}\\
\multicolumn{4}{l}{distribution for the intercepts}\\
\multicolumn{4}{l}{following (Wang \& Gelfand, 2013).}\\
\end{tabular}
\end{table}

First recall the regression equations predicting the linear outcome
\[\hat{y}_i = \gamma_0 + \gamma_{cos}\cos(\theta_i)r_i + \gamma_{sin}\sin(\theta_i)r_i + \gamma_1\text{SE}_i\]
\noindent and circular outcome
\[\hat{\boldsymbol{\mu}}_{i} = \begin{pmatrix}
  \mu_{i}^{I}  \\
\mu_{i}^{II}
 \end{pmatrix}=\begin{pmatrix}
  \beta_0^{I} + \beta_1^{I}\text{SE}_i  \\
  \beta_0^{II} + \beta_1^{II}\text{SE}_i
 \end{pmatrix}.\]
\noindent  For both models, \(\hat{y}\) is the predicted strength of
interpersonal behavior, \(\hat{\boldsymbol{\mu}}\) is the predicted mean vector
of the type of interpersonal behavior and the \(\gamma\)'s and \(\beta\)'s are
intercepts and regression coefficients.\newline \indent Table \ref{tab:estCLGPN}
shows the results for the modified CL-PN and CL-GPN models fit to the teacher
dataset. The estimates in this table are the averages and standard deviations of
the 10 cross-validation estimates. We show both the estimated posterior mode and
the 95\% highest posterior density (HPD) interval for each parameter. The
posterior estimates for the \(\beta\)'s and \(\gamma\)'s are quite similar for
both models. We also see this similarity when we look at the predicted circular
and linear means. The predicted linear mean is equal to \(\gamma_0\) which is
equal to 0.38 and 0.37 in the CL-PN and CL-GPN models, respectively. The
predicted circular means can be computed from \(\beta_0^{I}\) and
\(\beta_0^{II}\) using a double arctangent function \(\mbox{atan2}(\beta_0^{II},
\beta_0^{I})\), see \eqref{PredVal}. If we do this at each iteration of the MCMC
sampler we get the posterior distribution of the circular mean in both models.
Table \ref{tab:means} shows the posterior mode of the estimated circular means,
which are very similar (32.29\(^\circ\) and 33.54\(^\circ\)).\newline
\indent Next we  investigate the effect of self-efficacy. In the CL-PN model we
can transform the regression parameters on the two components \(I\) and \(II\)
to one circular regression coefficient \(b_c\) using the methods described by
@CremersMulderKlugkist2017. The dashed line in Figure \ref{regline} shows the
predicted regression line from the CL-PN model. Here $b_c$ is the slope of this
line at the inflection point (black square) and its posterior mode is estimated
at 1.67 (-24.66, 29.33)\footnote{Note that this is a linear approximation to the
circular regression line representing the slope at a specific point. Therefore
it is possible for the HPD interval to be wider than $2\pi$. In this case the
interval is much wider and covers 0, indicating there is no evidence to reject
the null-hypothesis of no effect.}. Even though the HPD of \(b_c\) includes 0,
indicating there is no evidence to reject the null-hypotheis of no effect, we
continue with its interpretation for eductional purposes. The interpretation of
$b_c$ is that at the inflection point an increase of 1 unit in self-efficacy
leads to an increase of \(1.67*(180/\pi) = 95.68^\circ\) in the type of
interpersonal behavior..However, as we can see in Figure \ref{regline} the
inflection point lies almost outside the range of the actual data. Instead of
looking at the inflection point we might compute the slope of the regression
line at the average self-efficacy. This parameter we call the slope at the mean
(\(SAM\)) [@CremersMulderKlugkist2017] and it is estimated at 0.02 (0.01; 0.04)
for our data. This means that at the average self-efficacy an increase of 1 unit
only leads to an increase of \(0.02*(180/\pi) = 1.15^\circ\) in the type of
interpersonal behavior. The HPD of the \(SAM\) does not include 0 which means
that the effect at the average self-efficacy can be distinguished from
0.\newline
\indent In the CL-GPN model we cannot compute circular regression coefficients
such as the \(b_c\) and \(SAM\) computed above  due to the fact that not only
the mean vector $\boldsymbol{\mu}$ but also the covariance matrix
$\boldsymbol{\Sigma}$ influences the predicted value on the circle. Instead, we
will compute posterior predictive distributions for the predicted circular
outcome of individuals scoring the minimum, maximum and median self-efficacy.
The modes and 95\% HPD intervals of these posterior predictive distributions are
\(\hat{\theta}_{SE_{min}} = 215.74^\circ (147.36^\circ, \: 44.49^\circ)\),
\(\hat{\theta}_{SE_{median}} = 25.93^\circ (337.02^\circ, \: 138.59^\circ)\),
\(\hat{\theta}_{SE_{max}} = 30.86^\circ (8.63^\circ, \: 72.19^\circ)\). Note
that we display the modes and HPD intervals for the posterior predictive
distributions on the interval $[0^\circ, 360^\circ)$. Note that  $44.49^\circ =
404.49^\circ$ due to the periodicity of a circular variable. The posterior mode estimate of
$215.74^\circ$ thus lies within its HPD interval $(147.36^\circ, \: 44.49^\circ)$.The
HPD intervals of the three posterior predictive distributions overlap. Had they
not overlapped we could have concluded that as the self-efficacy increases, the
score of the teacher on the IPC moves counterclockwise. \newline
\indent The effect of self-efficacy on the strength of interpersonal behavior is
quantified by \(\gamma_1\) in both the CL-PN and CL-GPN models. This parameter
can be interpreted in the same way as in a usual regression
model. The HPD interval of  \(\gamma_1\) includes 0 which means
that there is not enough evidence for an effect of self-efficacy on the strength
of interpersonal behavior. Note that the lower bounds of the HPD interval of
this parameter are  however very close to zero.\newline
\indent  The relation between the circular and linear outcome, that is, between
type of interpersonal behavior and strength of interpersonal behavior, is
described by the parameters $\gamma_{\cos}$ and $\gamma_{\sin}$. The HPD
interval of \(\gamma_{\cos}\) does not include 0 for both the CL-PN and CL-GPN
models, meaning that the cosine component of the type of interpersonal behavior
has an effect on the strength of interpersonal behavior. In the teacher data the
sine and cosine components have a substantive meaning. In this case the
Communion (cosine) component of the IPC positively effects the strength of a
teachers' type of interpersonal behavior, in plain words: teachers exhibiting
interpersonal behavior types with higher communion scores (e.g., 'helpful' and
'understanding' in Figure 2) are stronger in their behavior.

```{r circreg, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}

tikz("Plots/reglinediffSE.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")

a1 <- CLPN.BI.m[1]
b1 <- CLPN.BI.m[2]
a2 <- CLPN.BII.m[1]
b2 <- CLPN.BII.m[2]
ax <- -(a1*b1 + a2*b2)/(b1^2 + b2^2)
ac <- atan2(a2 + b2*ax, a1 + b1*ax)

x <- seq(-5, 5, by = 0.1)

pred.AL <- 0.37 + 2*atan(-0.01*x)
pred.PN <- atan2(a2 + b2*x, a1 + b1*x)

plot(Dat$SEc, Dat$theta*(180/pi), bty = "n", yaxt = "n",
     xlim = c(-5, 2), ylab = "$\\theta$", xlab = "self-efficacy")
axis(side = 2,  at = c(-160, -80, 0, 80, 160))
points(x, pred.AL*(180/pi), type = "l", lty = 1)
points(x, pred.PN*(180/pi), type = "l", lty = 2)
points(ax, ac*(180/pi), pch = 15)

dev.off()
tools::texi2dvi("Plots/reglinediffSE.tex", pdf=TRUE)

```

```{r reg.coef, cache = TRUE, echo = FALSE, results = FALSE}

BI    <- simplify2array(lapply(res_CLPN, "[[", "BI"))
BII   <- simplify2array(lapply(res_CLPN, "[[", "BII"))

bc <- matrix(NA, 15001, 10)
SAM <- matrix(NA, 15001, 10)

for(i in 1:10){

  a1 <- BI[5000:20000,1,i]
  b1 <- BI[5000:20000,2,i]
  a2 <- BII[5000:20000,1,i]
  b2 <- BII[5000:20000,2,i]

  ax <- -(a1*b1 + a2*b2)/(b1^2 + b2^2)
  ac <- atan2(a2 + b2*ax, a1 + b1*ax)
  bc[,i] <- -tan(atan2(a2, a1)-ac)/ax
  SAM[,i] <- bc[i]/(1 + (bc[i]*(-ax))^2)

}

mean(apply(bc, 2, mode_est))
rowMeans(apply(bc, 2, hpd_est))

mean(apply(SAM, 2, mode_est))
rowMeans(apply(SAM, 2, hpd_est))

```

\begin{figure}
\centering
\includegraphics[width = \textwidth]{Plots/reglinediffSE.pdf}
\caption{Plot showing circular regression lines for the effect of self-efficacy
as predicted by the Abe-Ley model (solid line) and CL-PN model (dashed line). The
black square indicates the inflection point of the circular regression line for
the CL-PN model.}
\label{regline}
\end{figure}

```{r ppreg, echo = FALSE, results = FALSE, eval = FALSE}
theta_pred_GPN.min <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred.min"))
theta_pred_GPN.median <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred.meadian"))
theta_pred_GPN.max <- simplify2array(lapply(res_CLGPN, "[[", "theta_pred.max"))

mean_circ(apply(theta_pred_GPN.min[5000:20000,], 2, mode_est_circ))*(180/pi)
mean_circ(apply(theta_pred_GPN.median[5000:20000,], 2, mode_est_circ))*(180/pi)
mean_circ(apply(theta_pred_GPN.max[5000:20000,], 2, mode_est_circ))*(180/pi)

apply(apply(theta_pred_GPN.min[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
apply(apply(theta_pred_GPN.median[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
apply(apply(theta_pred_GPN.max[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)

```

\subsubsection{The modified Abe-Ley model} \label{resAL}

```{r estAL, cache = TRUE, echo = FALSE}
#Get 2 dimensional arrays (parameter, fold)
resAL_arr <- simplify2array(lapply(resAL, "[[", "par"))

#Compute a cross-validation variance/standard deviation

#Put estimates in correct format for table
ALres.tab <- matrix(round(c(rowMeans(resAL_arr[1:4,]), mean(resAL_arr[6,]), 
                            mean(resAL_arr[5,]), mean(resAL_arr[7,]), 
                            apply(resAL_arr[1:4,], 1, sd), sd(resAL_arr[6,]), 
                            sd(resAL_arr[5,]), sd(resAL_arr[7,])), 2), 7, 2)

ALres.tab.new = cbind(apply(ALres.tab, 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (",
                                                             sprintf("%.2f", round(x[2], 2)), ")")))

rownames(ALres.tab.new) <- c("$\\eta_0$", "$\\eta_1$",
                             "$\\nu_0$", "$\\nu_1$",
                             "$\\alpha$", "$\\kappa$", "$\\lambda$")

#Create table
#kable(ALres.tab.new, "latex", booktabs = T, escape = FALSE, digits = 2,
#      caption = "Results, cross-validation mean and standard deviation, for the modified Abe-Ley model")%>%
#add_header_above(c("Parameter" = 1, "ML-estimate" = 1))

```

\begin{table}

\caption{\label{tab:estAL}Results, cross-validation mean and standard deviation, for the modified Abe-Ley model}
\centering
\begin{tabular}[t]{ll}
\toprule
\multicolumn{1}{c}{Parameter} & \multicolumn{1}{c}{ML-estimate} \\
\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-2}
$\eta_0$ & 0.36 (0.02)\\
$\eta_1$ & -0.03 (0.01)\\
$\nu_0$ & 1.17 (0.02)\\
$\nu_1$ & 0.04 (0.02)\\
$\alpha$ & 3.66 (0.12)\\
$\kappa$ & 1.51 (0.08)\\
$\lambda$ & 0.70 (0.05)\\
\bottomrule
\end{tabular}
\end{table}

First recall the regression equations predicting the linear outcome
\[\hat{\mu}_{i} = \eta_0 + 2 * \tan^{-1}(\eta_1\text{SE}_i)\]

\noindent and circular outcome
\[\hat{\beta}_{i} = \exp(\nu_0 + \nu_1\text{SE}_i).\]

\noindent Here \(\hat{\mu}\) is the predicted mean vector of the type
of interpersonal behavior and \(\hat{\beta}\) is the predicted scale
parameter of the strength of interpersonal behavior, while the \(\eta\)'s and
\(\nu\)'s are regression parameters.\newline
\indent Table \ref{tab:estAL} shows the results for the modified Abe-Ley model
fit to the teacher dataset. The estimates in this table are the averages and
standard deviations of the 10 cross-validation estimates. The estimate of the
circular mean at an average self-efficacy, \(\eta_0\), equals 0.36 radians or
20.62\(^\circ\). For the Abe-Ley model we can also investigate the effect of
self-efficacy on the type of interpersonal behavior. The solid line in
Figure \ref{regline} shows this effect. Unlike for the regression line of the
CL-PN model, the inflection point of the regression line of the Abe-Ley model
lies outside the x-range of the figure. \newline
\indent For the Abe-Ley model the conditional distribution for the linear
outcome is Weibull. This means that we can use methods from survival analsis to
interpret the effect of self-efficacy. In survival analysis a 'survival'
function is used in which time is plotted against the probability of survival of
subjects suffering from a specific medical condition. In our data however we
plot the strength on the IPC against the probability of a teacher having such a
strength. This probability is computed using the
'survival-function' \(\exp(-\alpha y_i^{\beta(1-\tanh(\kappa)\cos(\theta_i -
\mu_i))^{1/\alpha}})\) with \(\beta = \exp(\nu_0 + \nu_1\mbox{SE}_i)\). In Figure
\ref{reglineweib}, we plot the survival function for the minimum, median and
maximum value of self-efficacy. We conclude that stronger interpersonal
behaviors are less probable. We also see that the relation between self-efficacy and the strength on the IPC is not linear; the probability of having a stronger
interpersonal behavior is higher for both the minimum and maximum self-efficacy compared to a median self-efficacy score. Note however that the circular outcome also influences the survival function. The relation between the type of interpersonal behavior and the strength of interpersonal behavior may thus influence the shape of the survival function.\newline

```{r survivalplot, cache = TRUE, dev = "tikz",fig.ext="svg", echo = FALSE, results = FALSE}
summary(Dat$SEc)

datafig =  rbind(Dat[Dat$SEc == min(Dat$SEc),], Dat[Dat$SEc == median(Dat$SEc),], Dat[Dat$SEc == max(Dat$SEc),])

x = unique(datafig$SEc)
theta = c(datafig$theta[1], mean.circular(datafig$theta[2:(nrow(datafig)-1)]), datafig$theta[nrow(datafig)])
y <- seq(0, 1, by = 0.05)
#h <- matrix(NA, length(y), length(x))
s <- matrix(NA, length(y), length(x))
scross <- array(NA, dim = c(length(y), length(x), 10))

for(k in 1:10){

  train <- get(paste("train.", k, sep = ""))

  datafig =  rbind(train[train$SEc == min(train$SEc),],
                   train[train$SEc == median(train$SEc),],
                   train[train$SEc == max(train$SEc),])
  x = unique(datafig$SEc)
  theta = c(datafig$theta[1], mean.circular(datafig$theta[2:(nrow(datafig)-1)]), datafig$theta[nrow(datafig)])

    for(j in 1:length(x)){

      for(i in 1:length(y)){
          mu <- mean(resAL_arr[1,k]) + 2*atan(mean(resAL_arr[2,k])*x[j])
          beta <- exp(mean(resAL_arr[3,k]) + mean(resAL_arr[4,k])*x[j])*(1 - tanh(mean(resAL_arr[5,k]))*cos(theta[j] - mu))^(1/mean(resAL_arr[6,k]))
    #h[i,j] <- beta*mean(resAL_arr[6,])*y[i]^(mean(resAL_arr[6,])-1)
          s[i,j] <- exp(-mean(resAL_arr[6,k])*y[i]^beta)

    }

     scross[,,k] <- s

  }

}

s = apply(scross, c(1,2), mean)

tikz("Plots/survivaldiffSE.tex", standAlone =TRUE, height = 3.5, width = 6, pointsize = 12, engine = "pdftex")

plot(y, s[,1], type = "l", ylab = "P(strength IPC)", xlab = "strength IPC", bty = "n")
points(y, s[,2], type = "l", lty = 2)
points(y, s[,3], type = "l", lty = 3)
legend(0.7, 1, legend = c("min(SE)", "median(SE)", "max(SE)"), lty = c(1,2,3), bty = "n")

dev.off()
tools::texi2dvi("Plots/survivaldiffSE.tex", pdf=TRUE)

```

\begin{figure}
\centering
\includegraphics[width = \textwidth]{Plots/survivaldiffSE.pdf}
\caption{Plot showing the probability of having a particular strength of interpersonal behavior (survival plot) for the minimum, mean and maximum self-efficacy in the data.}
\label{reglineweib}
\end{figure}

```{r correlation, echo = FALSE, eval = FALSE, results = FALSE}

require(pracma)

alpha <- mean(resAL_arr[6,])
kappa <- mean(resAL_arr[5,])
lambda <- mean(resAL_arr[7,])

#R cannot compute legendre functions of non-integer degree
#leg.0 <- legendre(1/alpha, cosh(kappa))[1,] 
#leg.1 <- legendre(1/alpha, cosh(kappa))[2,]
#leg.2 <- legendre(1/alpha, cosh(kappa))[3,]

#so we use http://functions.wolfram.com/webMathematica/FunctionEvaluation.jsp?name=LegendreP2General

1/3.82
cosh(1.58)

```

\subsubsection{The modified GPN-SSN model} \label{resGPNSSN}

```{r estCLGPNM, cache = TRUE, echo = FALSE}
#Get 3 or 4 dimensional arrays (iteration, parameter, fold)/(,,itertation, fold) for each parameter type
lambda   <- simplify2array(lapply(res_CLGPNM, "[[", "lambda"))
Beta <- simplify2array(lapply(res_CLGPNM, "[[", "Beta"))
Betacon <- simplify2array(lapply(res_CLGPNM, "[[", "Betacon"))
Sigma <- simplify2array(lapply(res_CLGPNM, "[[", "Sigma"))
Sigmacon <- simplify2array(lapply(res_CLGPNM, "[[", "Sigmacon"))

#Averge over the folds and compute the posterior modes and hpd intervals
CLGPNM.Sigma11.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,1]
CLGPNM.Sigma11.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,1]
CLGPNM.Sigma12.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,2]
CLGPNM.Sigma12.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,1]
CLGPNM.Sigma13.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,3]
CLGPNM.Sigma13.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,3,1]
CLGPNM.Sigma22.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,2]
CLGPNM.Sigma22.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,2]
CLGPNM.Sigma23.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,3]
CLGPNM.Sigma23.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,3,2]
CLGPNM.Sigma33.m   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[3,3]
CLGPNM.Sigma33.hpd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,3,3]

CLGPNM.Sigma11.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,1]
CLGPNM.Sigma11.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,1]
CLGPNM.Sigma12.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,2]
CLGPNM.Sigma12.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,1]
CLGPNM.Sigma13.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,3]
CLGPNM.Sigma13.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,3,1]
CLGPNM.Sigma22.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,2]
CLGPNM.Sigma22.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,2]
CLGPNM.Sigma23.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,3]
CLGPNM.Sigma23.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,3,2]
CLGPNM.Sigma33.m.sd   <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[3,3]
CLGPNM.Sigma33.hpd.sd <- apply(apply(Sigma[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,3,3]

CLGPNM.Sigmacon11.m   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,1]
CLGPNM.Sigmacon11.hpd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,1]
CLGPNM.Sigmacon12.m   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,2]
CLGPNM.Sigmacon12.hpd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,1]
CLGPNM.Sigmacon13.m   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,3]
CLGPNM.Sigmacon13.hpd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,3,1]
CLGPNM.Sigmacon22.m   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,2]
CLGPNM.Sigmacon22.hpd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,2]
CLGPNM.Sigmacon23.m   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,3]
CLGPNM.Sigmacon23.hpd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,3,2]
CLGPNM.Sigmacon33.m   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[3,3]
CLGPNM.Sigmacon33.hpd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,3,3]

CLGPNM.Sigmacon11.m.sd   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,1]
CLGPNM.Sigmacon11.hpd.sd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,1]
CLGPNM.Sigmacon12.m.sd   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,2]
CLGPNM.Sigmacon12.hpd.sd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,1]
CLGPNM.Sigmacon13.m.sd   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,3]
CLGPNM.Sigmacon13.hpd.sd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,3,1]
CLGPNM.Sigmacon22.m.sd   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,2]
CLGPNM.Sigmacon22.hpd.sd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,2]
CLGPNM.Sigmacon23.m.sd   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,3]
CLGPNM.Sigmacon23.hpd.sd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,3,2]
CLGPNM.Sigmacon33.m.sd   <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[3,3]
CLGPNM.Sigmacon33.hpd.sd <- apply(apply(Sigmacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,3,3]

CLGPNM.Beta11.m   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,1]
CLGPNM.Beta11.hpd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,1]
CLGPNM.Beta21.m   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,1]
CLGPNM.Beta21.hpd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,1]
CLGPNM.Beta12.m   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,2]
CLGPNM.Beta12.hpd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,2]
CLGPNM.Beta22.m   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,2]
CLGPNM.Beta22.hpd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,2]
CLGPNM.Beta13.m   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,3]
CLGPNM.Beta13.hpd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,3]
CLGPNM.Beta23.m   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,3]
CLGPNM.Beta23.hpd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,3]

CLGPNM.Beta11.m.sd   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,1]
CLGPNM.Beta11.hpd.sd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,1]
CLGPNM.Beta21.m.sd   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,1]
CLGPNM.Beta21.hpd.sd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,1]
CLGPNM.Beta12.m.sd   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,2]
CLGPNM.Beta12.hpd.sd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,2]
CLGPNM.Beta22.m.sd   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,2]
CLGPNM.Beta22.hpd.sd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,2]
CLGPNM.Beta13.m.sd   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,3]
CLGPNM.Beta13.hpd.sd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,3]
CLGPNM.Beta23.m.sd   <- apply(apply(Beta[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,3]
CLGPNM.Beta23.hpd.sd <- apply(apply(Beta[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,3]

CLGPNM.Betacon11.m   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,1]
CLGPNM.Betacon11.hpd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,1]
CLGPNM.Betacon21.m   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,1]
CLGPNM.Betacon21.hpd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,1]
CLGPNM.Betacon12.m   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,2]
CLGPNM.Betacon12.hpd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,2]
CLGPNM.Betacon22.m   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,2]
CLGPNM.Betacon22.hpd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,2]
CLGPNM.Betacon13.m   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[1,3]
CLGPNM.Betacon13.hpd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,1,3]
CLGPNM.Betacon23.m   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, mean)[2,3]
CLGPNM.Betacon23.hpd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, mean)[,2,3]

CLGPNM.Betacon11.m.sd   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,1]
CLGPNM.Betacon11.hpd.sd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,1]
CLGPNM.Betacon21.m.sd   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,1]
CLGPNM.Betacon21.hpd.sd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,1]
CLGPNM.Betacon12.m.sd   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,2]
CLGPNM.Betacon12.hpd.sd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,2]
CLGPNM.Betacon22.m.sd   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,2]
CLGPNM.Betacon22.hpd.sd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,2]
CLGPNM.Betacon13.m.sd   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[1,3]
CLGPNM.Betacon13.hpd.sd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,1,3]
CLGPNM.Betacon23.m.sd   <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), mode_est), 1:2, sd)[2,3]
CLGPNM.Betacon23.hpd.sd <- apply(apply(Betacon[,,5000:20000,], c(1,2,4), hpd_est), 1:3, sd)[,2,3]

CLGPNM.lambda.m   <- mean(apply(lambda[5000:20000,,], 2, mode_est))
CLGPNM.lambda.m.sd   <- sd(apply(lambda[5000:20000,,], 2, mode_est))
CLGPNM.lambda.hpd <- rowMeans(apply(lambda[5000:20000,,], 2, hpd_est))
CLGPNM.lambda.hpd.sd <- apply(apply(lambda[5000:20000,,], 2, hpd_est), 1, sd)

#Compute a cross-validation variance/standard deviation

#Put estimates in correct format for tables
modes.GPN  <- c(CLGPNM.Beta11.m, CLGPNM.Beta12.m, CLGPNM.Beta13.m,
                CLGPNM.Beta21.m, CLGPNM.Beta22.m, CLGPNM.Beta23.m,
                CLGPNM.Sigma11.m, CLGPNM.Sigma22.m, CLGPNM.Sigma33.m,
                CLGPNM.Sigma12.m, CLGPNM.Sigma13.m, CLGPNM.Sigma23.m,
                CLGPNM.lambda.m)
hpd.LB.GPN <- c(CLGPNM.Beta11.hpd[1], CLGPNM.Beta12.hpd[1], CLGPNM.Beta13.hpd[1],
                CLGPNM.Beta21.hpd[1], CLGPNM.Beta22.hpd[1], CLGPNM.Beta23.hpd[1],
                CLGPNM.Sigma11.hpd[1], CLGPNM.Sigma22.hpd[1], CLGPNM.Sigma33.hpd[1],
                CLGPNM.Sigma12.hpd[1], CLGPNM.Sigma13.hpd[1], CLGPNM.Sigma23.hpd[1],
                CLGPNM.lambda.hpd[1])
hpd.UB.GPN <- c(CLGPNM.Beta11.hpd[2], CLGPNM.Beta12.hpd[2], CLGPNM.Beta13.hpd[2],
                CLGPNM.Beta21.hpd[2], CLGPNM.Beta22.hpd[2], CLGPNM.Beta23.hpd[2],
                CLGPNM.Sigma11.hpd[2], CLGPNM.Sigma22.hpd[2], CLGPNM.Sigma33.hpd[2],
                CLGPNM.Sigma12.hpd[2], CLGPNM.Sigma13.hpd[2], CLGPNM.Sigma23.hpd[2],
                CLGPNM.lambda.hpd[2])

modescon.GPN  <- c(CLGPNM.Betacon11.m, CLGPNM.Betacon12.m, CLGPNM.Betacon13.m,
                   CLGPNM.Betacon21.m, CLGPNM.Betacon22.m, CLGPNM.Betacon23.m,
                   CLGPNM.Sigmacon11.m, CLGPNM.Sigmacon22.m, CLGPNM.Sigmacon33.m,
                   CLGPNM.Sigmacon12.m, CLGPNM.Sigmacon13.m, CLGPNM.Sigmacon23.m,
                   CLGPNM.lambda.m)
hpdcon.LB.GPN <- c(CLGPNM.Betacon11.hpd[1], CLGPNM.Betacon12.hpd[1], CLGPNM.Betacon13.hpd[1],
                   CLGPNM.Betacon21.hpd[1], CLGPNM.Betacon22.hpd[1], CLGPNM.Betacon23.hpd[1],
                   CLGPNM.Sigmacon11.hpd[1], CLGPNM.Sigmacon22.hpd[1], CLGPNM.Sigmacon33.hpd[1],
                   CLGPNM.Sigmacon12.hpd[1], CLGPNM.Sigmacon13.hpd[1], CLGPNM.Sigmacon23.hpd[1],
                   CLGPNM.lambda.hpd[1])
hpdcon.UB.GPN <- c(CLGPNM.Betacon11.hpd[2], CLGPNM.Betacon12.hpd[2], CLGPNM.Betacon13.hpd[2],
                   CLGPNM.Betacon21.hpd[2], CLGPNM.Betacon22.hpd[2], CLGPNM.Betacon23.hpd[2],
                   CLGPNM.Sigmacon11.hpd[2], CLGPNM.Sigmacon22.hpd[2], CLGPNM.Sigmacon33.hpd[2],
                   CLGPNM.Sigmacon12.hpd[2], CLGPNM.Sigmacon13.hpd[2], CLGPNM.Sigmacon23.hpd[2],
                   CLGPNM.lambda.hpd[2])

modes.GPN.sd  <- c(CLGPNM.Beta11.m.sd, CLGPNM.Beta12.m.sd, CLGPNM.Beta13.m.sd,
                   CLGPNM.Beta21.m.sd, CLGPNM.Beta22.m.sd, CLGPNM.Beta23.m.sd,
                   CLGPNM.Sigma11.m.sd, CLGPNM.Sigma22.m.sd, CLGPNM.Sigma33.m.sd,
                   CLGPNM.Sigma12.m.sd, CLGPNM.Sigma13.m.sd, CLGPNM.Sigma23.m.sd,
                  CLGPNM.lambda.m.sd)
hpd.LB.GPN.sd <- c(CLGPNM.Beta11.hpd.sd[1], CLGPNM.Beta12.hpd.sd[1], CLGPNM.Beta13.hpd[1],
                   CLGPNM.Beta21.hpd.sd[1], CLGPNM.Beta22.hpd.sd[1], CLGPNM.Beta23.hpd[1],
                   CLGPNM.Sigma11.hpd.sd[1], CLGPNM.Sigma22.hpd.sd[1], CLGPNM.Sigma33.hpd[1],
                   CLGPNM.Sigma12.hpd.sd[1], CLGPNM.Sigma13.hpd.sd[1], CLGPNM.Sigma23.hpd[1],
                   CLGPNM.lambda.hpd.sd[1])
hpd.UB.GPN.sd <- c(CLGPNM.Beta11.hpd.sd[2], CLGPNM.Beta12.hpd.sd[2], CLGPNM.Beta13.hpd.sd[2],
                   CLGPNM.Beta21.hpd.sd[2], CLGPNM.Beta22.hpd.sd[2], CLGPNM.Beta23.hpd.sd[2],
                   CLGPNM.Sigma11.hpd.sd[2], CLGPNM.Sigma22.hpd.sd[2], CLGPNM.Sigma33.hpd.sd[2],
                   CLGPNM.Sigma12.hpd.sd[2], CLGPNM.Sigma13.hpd.sd[2], CLGPNM.Sigma23.hpd.sd[2],
                   CLGPNM.lambda.hpd.sd[2])

modescon.GPN.sd  <- c(CLGPNM.Betacon11.m.sd, CLGPNM.Betacon12.m.sd, CLGPNM.Betacon13.m.sd,
                      CLGPNM.Betacon21.m.sd, CLGPNM.Betacon22.m.sd, CLGPNM.Betacon23.m.sd,
                      CLGPNM.Sigmacon11.m.sd, CLGPNM.Sigmacon22.m.sd, CLGPNM.Sigmacon33.m.sd,
                      CLGPNM.Sigmacon12.m.sd, CLGPNM.Sigmacon13.m.sd, CLGPNM.Sigmacon23.m.sd,
                      CLGPNM.lambda.m.sd)
hpdcon.LB.GPN.sd <- c(CLGPNM.Betacon11.hpd.sd[1], CLGPNM.Betacon12.hpd.sd[1], CLGPNM.Betacon13.hpd.sd[1],
                      CLGPNM.Betacon21.hpd.sd[1], CLGPNM.Betacon22.hpd.sd[1], CLGPNM.Betacon23.hpd.sd[1],
                      CLGPNM.Sigmacon11.hpd.sd[1], CLGPNM.Sigmacon22.hpd.sd[1], CLGPNM.Sigmacon33.hpd.sd[1],
                      CLGPNM.Sigmacon12.hpd.sd[1], CLGPNM.Sigmacon13.hpd.sd[1], CLGPNM.Sigmacon23.hpd.sd[1],
                      CLGPNM.lambda.hpd.sd[1])
hpdcon.UB.GPN.sd <- c(CLGPNM.Betacon11.hpd.sd[2], CLGPNM.Betacon12.hpd.sd[2], CLGPNM.Betacon13.hpd.sd[2],
                      CLGPNM.Betacon21.hpd.sd[2], CLGPNM.Betacon22.hpd.sd[2], CLGPNM.Betacon23.hpd.sd[2],
                      CLGPNM.Sigmacon11.hpd.sd[2], CLGPNM.Sigmacon22.hpd.sd[2], CLGPNM.Sigmacon33.hpd.sd[2],
                      CLGPNM.Sigmacon12.hpd.sd[2], CLGPNM.Sigmacon13.hpd.sd[2], CLGPNM.Sigmacon23.hpd.sd[2],
                      CLGPNM.lambda.hpd.sd[2])

#Create table
CLGPNMres.tab <- cbind(modes.GPN, modes.GPN.sd, hpd.LB.GPN, hpd.LB.GPN.sd,
                       hpd.UB.GPN, hpd.UB.GPN.sd, modescon.GPN, modescon.GPN.sd, 
                       hpdcon.LB.GPN, hpdcon.LB.GPN.sd, hpdcon.UB.GPN, hpdcon.UB.GPN.sd)

CLGPNMres.tab.new = cbind(apply(CLGPNMres.tab[,1:2], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                           sprintf("%.2f", round(x[2], 2)), ")")),
                          apply(CLGPNMres.tab[,3:4], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                           sprintf("%.2f", round(x[2], 2)), ")")),
                          apply(CLGPNMres.tab[,5:6], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                           sprintf("%.2f", round(x[2], 2)), ")")),
                          apply(CLGPNMres.tab[,7:8], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                           sprintf("%.2f", round(x[2], 2)), ")")),
                          apply(CLGPNMres.tab[,9:10], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                            sprintf("%.2f", round(x[2], 2)), ")")),
                          apply(CLGPNMres.tab[,11:12], 1, function(x) paste0(sprintf("%.2f", round(x[1], 2)), " (", 
                                                                             sprintf("%.2f", round(x[2], 2)), ")")))

rownames(CLGPNMres.tab.new) <- c("$\\beta_{0_s^{I}}$", "$\\beta_{0_s^{II}}$",
                                 "$\\beta_{0_y}$", "$\\beta_{1_s^{I}}$",
                                 "$\\beta_{1_s^{II}}$", "$\\beta_{1_y}$",
                                 "$\\sum_{s_{1,1}}$", "$\\sum_{s_{2,2}}$",
                                 "$\\sum_{y_{3,3}}$", "$\\sum_{s_{1,2}}$",
                                 "$\\sum_{sy_{1,3}}$", "$\\sum_{sy_{2,3}}$",
                                 "$\\lambda$")
colnames(CLGPNMres.tab.new) <- c("Mode", "HPD LB", "HPD UB", "Mode", "HPD LB", "HPD UB")

#kable(CLGPNMres.tab.new, "latex", booktabs = TRUE, escape = FALSE, digits = 2,
#      caption = "Results, cross-validation mean and standard deviation, for the GPN-SSN model")%>%
#add_header_above(c("Parameter" = 1, "Unconstrained" = 3, "Constrained" = 3))

```

```{r ppreggpnm, echo = FALSE, results = FALSE, eval = FALSE}
theta_pred_GPNM.min <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred.min"))
theta_pred_GPNM.median <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred.median"))
theta_pred_GPNM.max <- simplify2array(lapply(res_CLGPNM, "[[", "theta_pred.max"))

mean_circ(apply(theta_pred_GPNM.min[5000:20000,], 2, mode_est_circ))*(180/pi)
mean_circ(apply(theta_pred_GPNM.median[5000:20000,], 2, mode_est_circ))*(180/pi)
mean_circ(apply(theta_pred_GPNM.max[5000:20000,], 2, mode_est_circ))*(180/pi)

dim(theta_pred_GPNM.median)

apply(apply(theta_pred_GPNM.min[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
apply(apply(theta_pred_GPNM.median[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)
apply(apply(theta_pred_GPNM.max[5000:20000,], 2, hpd_est_circ), 1, mean_circ)*(180/pi)

```

\begin{table}
\caption{\label{tab:estCLGPNM}Results, cross-validation mean and standard deviation, for the GPN-SSN model}
\centering
\begin{tabular}[t]{lllllll}
\toprule
\multicolumn{1}{c}{Parameter} & \multicolumn{3}{c}{Unconstrained} & \multicolumn{3}{c}{Constrained} \\
\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-4} \cmidrule(l{2pt}r{2pt}){5-7}
  & Mode & HPD LB & HPD UB & Mode & HPD LB & HPD UB\\
\midrule
$\beta_{0_s^{I}}$ & 0.30 (0.01) & 0.26 (0.01) & 0.34 (0.01) & 2.11 (0.10) & 1.75 (0.09) & 2.50 (0.11)\\
$\beta_{0_s^{II}}$ & 0.19 (0.00) & 0.17 (0.01) & 0.21 (0.00) & 1.33 (0.07) & 1.10 (0.05) & 1.57 (0.06)\\
$\beta_{0_y}$ & 0.33 (0.01) & 0.30 (0.30) & 0.36 (0.01) & 0.33 (0.01) & 0.30 (0.01) & 0.36 (0.01)\\
$\beta_{1_s^{I}}$ & 0.09 (0.01) & 0.05 (0.01) & 0.13 (0.01) & 0.60 (0.06) & 0.33 (0.05) & 0.90 (0.06)\\
$\beta_{1_s^{II}}$ & 0.07 (0.00) & 0.04 (0.00) & 0.09 (0.01) & 0.48 (0.03) & 0.30 (0.03) & 0.66 (0.04)\\
\addlinespace
$\beta_{1_y}$ & 0.09 (0.01) & 0.06 (0.06) & 0.12 (0.01) & 0.09 (0.01) & 0.06 (0.01) & 0.12 (0.01)\\
$\sum_{s_{1,1}}$ & 0.05 (0.00) & 0.04 (0.00) & 0.06 (0.00) & 2.43 (0.14) & 1.72 (0.07) & 3.46 (0.13)\\
$\sum_{s_{2,2}}$ & 0.02 (0.00) & 0.02 (0.00) & 0.03 (0.00) & 1.00 (0.00) & 1.00 (0.00) & 1.00 (0.00)\\
$\sum_{y_{3,3}}$ & 0.03 (0.00) & 0.02 (0.02) & 0.04 (0.00) & 0.03 (0.00) & 0.02 (0.00) & 0.04 (0.00)\\
$\sum_{s_{1,2}}$ & 0.00 (0.00) & -0.00 (0.00) & 0.01 (0.00) & 0.07 (0.06) & -0.20 (0.06) & 0.35 (0.06)\\
\addlinespace
$\sum_{sy_{1,3}}$ & 0.03 (0.00) & 0.02 (0.00) & 0.04 (0.00) & 0.23 (0.01) & 0.17 (0.01) & 0.32 (0.01)\\
$\sum_{sy_{2,3}}$ & 0.01 (0.00) & 0.01 (0.01) & 0.02 (0.00) & 0.09 (0.01) & 0.06 (0.01) & 0.12 (0.01)\\
$\lambda$ & 0.16 (0.01) & 0.14 (0.01) & 0.18 (0.01) & 0.16 (0.01) & 0.14 (0.01) & 0.18 (0.01)\\
\bottomrule
\end{tabular}
\end{table}

First recall the regression equation predicting the circular and linear
outcome:

\[\hat{\boldsymbol{\mu}}_{i} = \boldsymbol{\beta}_0 +\boldsymbol{\beta}_1\text{SE}_i,\]

\noindent where
\(\boldsymbol{\mu}_i = (\boldsymbol{\mu}_{s_i}, \boldsymbol{\mu}_{y_i})^t\),
\(\boldsymbol{\beta}_0 = (\beta_{0_{s^{I}}}, \beta_{0_{s^{II}}},\beta_{0_y})^t\)
and
\(\boldsymbol{\beta}_1 = (\beta_{1_{s^{I}}}, \beta_{1_{s^{II}}},\beta_{1_y})^t\).
The parameters \(\beta_{0_{s^{I}}}\), \(\beta_{0_{s^{II}}}\),
\(\beta_{1_{s^{I}}}\) and \(\beta_{1_{s^{II}}}\) are the intercepts and
regression coefficients for the circular outcome and \(\beta_{0_y}\) and
\(\beta_{1_y}\) are the intercept and regression coefficient for the
linear outcome.\newline
\indent Table \ref{tab:estCLGPNM} shows the results for the modified GPN-SSN model fit to
the teacher dataset. The estimates in this table are the averages and
standard deviations of the 10 cross-validation estimates. We show both
the estimated posterior mode and the 95\% highest posterior density
(HPD) interval for each parameter. The predicted circular means can be
computed from \(\beta_{0_{s^{I}}}\) and \(\beta_{0_{s^{II}}}\) in a
similar fashion as for the CL-PN and CL-GPN models. Table \ref{tab:means} shows the
posterior mode of the estimated circular mean, which equals 35.30\(^\circ\) and hence
is very similar to those of the CL-PN and CL-GPN models.\newline
\indent For the same reason as in the CL-GPN model we cannot compute circular
regression coefficients for the effect of self-efficacy on the type of
interpersonal behavior such as the \(b_c\) and \(SAM\). Instead, we will again
compute posterior predictive distributions for the predicted circular outcome of
individuals scoring the minimum, maximum and median self-efficacy. The modes
and 95\% HPD intervals of these posterior predictive distributions are
\(\hat{\theta}_{SE_{min}} = 206.87^\circ (117.11^\circ, \: 72.02^\circ)\),
\(\hat{\theta}_{SE_{median}} = 24.68^\circ (334.73^\circ, \: 128.27^\circ)\),
\(\hat{\theta}_{SE_{max}} = 29.81^\circ (0.74^\circ, \: 80.61^\circ)\). On a
circle the HPD intervals of the three posterior predictive distributions
overlap. Had they not overlapped we could have concluded that as the
self-efficacy increases, the score of the teacher on the IPC moves
counterclockwise. The effect of self-efficacy on the strength of interpersonal
behavior is quantified by \(\beta_{1_y}\), and we learn that for a 1 unit
increase in self-efficacy the strength of interpersonal behavior increases by
0.09. The average strength is quantified by \(\beta_{0_y}\) and equals
0.33.\newline
\indent To investigate the association between the linear and circular outcome
we look at the covariances between the linear outcome and the sine and cosine of
the circular outcome \(\sum_{sy_{2,3}}\) and \(\sum_{sy_{1,3}}\). Both
covariances, \(\sum_{sy_{2,3}} = 0.09\) and \(\sum_{sy_{1,3}} = 0.23\), are
different from zero, but the one of the cosine component is larger. This means
the correlation with the Communion component is larger and that teachers scoring
both high on Communion and Agency show stronger behavior. This is a slightly
different conclusion from the one in the CL-PN and CL-GPN models.

\subsubsection{Model fit}

In this section we will assess the overall fit of the cylindrical models to the
data via the PLSL criterion described in Section \ref{Modelfit}. Table
\ref{tab:ModelFit} shows the values of this criterion for the linear and
circular outcomes of the four models.\newline
\indent The CL-PN and CL-GPN models have the best out-of-sample predictive
performance for the linear outcome. They show roughly the same performance because
they model the linear outcome in the same way. Only the value of \(r\) in
\eqref{circlinlink} differs. We should note that even though the predictive
performance of the Abe-Ley model for the linear outcome is worst on average, the
standard deviation of the cross-validation estimates is rather large. This means
that in some samples, the Abe-Ley model shows a lower PLSL value than the
average of 25.49\newline
\indent The Abe-Ley model has the best out-of-sample predictive performance for
the circular outcome. This would suggest that for the circular variable a
slightly skewed distribution fits best. However, both the GPN-SSN  and the
CL-GPN models fit much worse even though the distribution for the circular
outcome in these models can also take a skewed shape. It should be noted that
the  standard deviation of the cross-validation estimates is rather large for
the Abe-Ley and the CL-GPN model. It is possible that these large standard
deviations for the PLSL are caused by the fact that they are computed for a
relatively small sample size, but this does not explain why the PLSL has a large
standard deviation for only a few cylindrical models and not for all.

```{r ModelFit, cache = TRUE, echo = FALSE}

CLGPNM.ll.circ <- simplify2array(lapply(res_CLGPNM, "[[", "ll.circ"))
CLGPNM.ll.lin  <- simplify2array(lapply(res_CLGPNM, "[[", "ll.lin"))
CLGPN.ll.circ  <- simplify2array(lapply(res_CLGPN, "[[", "ll.circ"))
CLGPN.ll.lin   <- simplify2array(lapply(res_CLGPN, "[[", "ll.lin"))
CLPN.ll.circ   <- simplify2array(lapply(res_CLPN, "[[", "ll.circ"))
CLPN.ll.lin    <- simplify2array(lapply(res_CLPN, "[[", "ll.lin"))

CLGPNM.ll.circ.m <- mean(-2*apply(CLGPNM.ll.circ[5:20,], 2, mean))
CLGPNM.ll.lin.m  <- mean(-2*apply(CLGPNM.ll.lin[5:20,], 2, mean))
CLGPN.ll.circ.m  <- mean(-2*apply(CLGPN.ll.circ[5:20,], 2, mean))
CLGPN.ll.lin.m   <- mean(-2*apply(CLGPN.ll.lin[5:20,], 2, mean))
CLPN.ll.circ.m   <- mean(-2*apply(CLPN.ll.circ[5:20,], 2, mean))
CLPN.ll.lin.m    <- mean(-2*apply(CLPN.ll.lin[5:20,], 2, mean))

CLGPNM.ll.circ.sd <- sd(-2*apply(CLGPNM.ll.circ[5:20,], 2, mean))
CLGPNM.ll.lin.sd  <- sd(-2*apply(CLGPNM.ll.lin[5:20,], 2, mean))
CLGPN.ll.circ.sd  <- sd(-2*apply(CLGPN.ll.circ[5:20,], 2, mean))
CLGPN.ll.lin.sd   <- sd(-2*apply(CLGPN.ll.lin[5:20,], 2, mean))
CLPN.ll.circ.sd   <- sd(-2*apply(CLPN.ll.circ[5:20,], 2, mean))
CLPN.ll.lin.sd    <- sd(-2*apply(CLPN.ll.lin[5:20,], 2, mean))

AL.ll.circ <- simplify2array(lapply(resAL, "[[", "ll.circ"))
AL.ll.lin <- simplify2array(lapply(resAL, "[[", "ll.lin"))

AL.ll.circ.m <- mean(AL.ll.circ)
AL.ll.lin.m  <- mean(AL.ll.lin)

AL.ll.circ.sd <- sd(AL.ll.circ)
AL.ll.lin.sd  <- sd(AL.ll.lin)

PLSL.m <- round(rbind(c(CLPN.ll.circ.m, CLGPN.ll.circ.m, AL.ll.circ.m, CLGPNM.ll.circ.m),
                      c(CLPN.ll.lin.m, CLGPN.ll.lin.m, AL.ll.lin.m, CLGPNM.ll.lin.m)), 2)
PLSL.sd <- round(rbind(c(CLPN.ll.circ.sd, CLGPN.ll.circ.sd, AL.ll.circ.sd, CLGPNM.ll.circ.sd),
                      c(CLPN.ll.lin.sd, CLGPN.ll.lin.sd, AL.ll.lin.sd, CLGPNM.ll.lin.sd)), 2)
PLSL.sd <- apply(PLSL.sd, 1:2, function(x) paste0("(", sprintf("%.2f", round(x, 2)), ")"))

PLSL <- data.frame(PLSL.m[1,], PLSL.sd[1,], PLSL.m[2,], PLSL.sd[2,])

rownames(PLSL) <- c("CL-PN", "CL-GPN", "Abe-Ley", "GPN-SSN")
colnames(PLSL) <- c("mean", "sd", "mean", "sd")

#kable(PLSL, "latex", booktabs = T, escape = FALSE,
#      caption = "PLSL criteria, cross-validation mean and standard deviation, for the circular and linear outcome in the four #cylindrical models")%>%
#add_header_above(c("Model" = 1, "Circular" = 2, "Linear" = 2))

```

\begin{table}

\caption{\label{tab:ModelFit}PLSL criteria, cross-validation mean and standard deviation, for the circular and linear outcome in the four cylindrical models}
\centering
\begin{tabular}[t]{lrlrl}
\toprule
\multicolumn{1}{c}{Model} & \multicolumn{2}{c}{Circular} & \multicolumn{2}{c}{Linear} \\
\cmidrule(l{2pt}r{2pt}){1-1} \cmidrule(l{2pt}r{2pt}){2-3} \cmidrule(l{2pt}r{2pt}){4-5}
  & mean & sd & mean & sd\\
\midrule
CL-PN & 82.96 & (9.47) & -17.65 & (3.70)\\
CL-GPN & 85.22 & (18.12) & -18.12 & (3.57)\\
Abe-Ley & 31.97 & (22.07) & 25.49 & (17.46)\\
GPN-SSN & 106.38 & (8.84) & -2.14 & (6.78)\\
\bottomrule
\end{tabular}
\end{table}

\section{Discussion}\label{Discussion}

In this paper we modified four models for cylindrical data in such a way that
they include a regression of both the linear and circular outcome onto a
set of covariates. Subsequently we have shown how these four methods can
be used to analyze a dataset on the interpersonal behavior of teachers.
In this final section we will comment on the differences between these models, the
results from the analysis of the teacher data and how cylindrical models
improve the analysis of such cylindrical data.\newline
\indent In terms of interpretability, the CL-PN and Abe-Ley models perform
best.  In the
CL-GPN and GPN-SSN models the interpretation of the parameters of the
circular outcome component is not straightforward, if at all possible.
This is caused by the fact that in addition to the mean vector the
covariance matrix of the GPN distribution affects the location of the
circular data, making it difficult to compute regression coefficients on
the circle. @wang2012directional state that Monte Carlo integration
can be used to compute a circular mean and variance for the GPN
distribution. In future research, this solution might be applied to the
methods of @CremersMulderKlugkist2017 in order to compute circular
coefficients for GPN models.\newline
\indent In terms of flexibility the GPN-SSN model scores best. Multiple
linear and circular outcomes can be included and we can thus apply the
model to multivariate cylindrical data. In addition the GPN-SSN, the
CL-GPN and CL-PN models are extendable to a mixed-effects structure and
can thus also be fit to longitudinal data (see @nunez2014bayesian and @hernandez2016general for
hierarchical/mixed-effects models for the PN and GPN distributions
respectively). For the Abe-Ley model this may also be possible but has
not been done in previous research for the conditional distribution of
its circular outcome (sine-skewed von
Mises). Concerning asymmetry, both the
GPN-SSN as well as the Abe-Ley model allow for non-symmetrical shapes of
the distributions of both the linear and circular outcome, while the CL-GPN
model permits an asymmetric  circular outcome.\newline
\indent To investigate model fit for the teacher data, we assessed
out-of-sample predictive performance for both the linear and circular
outcome. The CL-PN and CL-GPN models have the best fit for the linear
outcome while the Abe-Ley model fits best  the circular outcome.
Differences in fit, in addition to being a result of different
distributional assumptions, may also be caused by the way in which the
relation between the linear and circular outcome is modeled. Whereas in
both the Abe-Ley and GPN-SSN models the distribution of the linear
outcome is conditioned on the circular outcome and vice versa, the
distribution of the circular outcome in the CL-PN and CL-GPN models is
independent of the linear outcome. In these models the circular outcome
is regressed onto the linear outcome.\newline
\indent The four cylindrical models that were modified to the regression
context in this paper are not the only cylindrical distributions
available from the literature. Other interesting cylindrical distributions have been introduced by
@fernandez2007models, @kato2008dependent and @sugasawa2015flexible (for more references we
refer to Chapter 2 of @ley2017modern). In the present study we
have decided not to include these distributions for reasons of space,
complexity of the models and ease of implementing a regression
structure. In future research however it would be interesting to
investigate other types of cylindrical distributions as well in order to
compare the interpretability, flexibility and model fit to the models
developed in the present study.\newline
\indent We conclude the paper on a general note regarding cylindrical models.
They  offer new insights into data of a cylindrical nature in psychology.
Concerning the example used in this paper, the advantage of cylindrical data
analysis is that we were able to analyze all circular and linear information in
the data simultaneously. In previous research, the two components of the
interpersonal circumplex (\emph{i.e.}, Agency and Communion) were analyzed
separately. Such an approach also provides information about the strength of
teachers' score on Agency and Communion, yet a large portion of information
about the combination of Agency and Communion, which describes the kind of
behavior that is observed,  gets lost. A first solution to include both
dimensions as a circular variable in data analysis was described by
@Cremers2018Assessing. A downside of that analysis was that information about
the strength of the specific type of interpersonal behavior could not be
retained. In the present study, we have shown how using cylindrical models can
simultaneously model the information about the type of and strength of
interpersonal behavior and how these are influenced by teachers' self-efficacy
in classroom management. The results of the present study therefore provide an
answer to an often stated problem in data analysis of interpersonal circumplex
data.

\section*{References}
<div id="refs"></div>