### 2019-01-28 Manuscript/SupplementaryMaterial.Rmd

---
title: " "
author: |
bibliography: CircularData.bib
csl: apa.csl
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: no
fontsize: 12pt
geometry: margin=1in
header-includes:
  - \usepackage{multirow}
  - \usepackage{appendix}
  - \usepackage{color}
  - \usepackage{hyperref}
  - \usepackage{subcaption}
  - \setlength\parindent{24pt}
  - \usepackage{setspace}\doublespacing
documentclass: article
pandoc_args: --natbib
---

\section{Supplementary Material}\label{Appendix}

In this Supplementary Material we give a more detailed description of the cylindrical regression models and outline the MCMC procedures to fit them. R-code for the MCMC sampler and the analysis of the teacher data can be found here: \url{https://github.com/joliencremers/CylindricalComparisonCircumplex}. Note that the dimensions of the objects (design matrices, mean vectors, etc.) are those that were used in the analysis of the teacher data where we have 1 circular outcome, 1 linear outcome and estimate an intercept and regression coefficient for the covariate self-efficacy. Note that for the regression of the linear component in the CL-PN and CL-GPN models we also have the sine and cosine of the circular outcome in the regression equation, this makes the vector with regression coefficients, $\boldsymbol{\gamma}$, four-dimensional.

\section{Four cylindrical regression models}\label{Models}

\subsection{The modified CL-PN and modified CL-GPN  models}\label{CL-(G)PN}

Following @mastrantonio2015bayesian we consider in this section two models where
the relation between \(\Theta \in [0, 2\pi)\) and \(Y\in (-\infty, + \infty)\)
and $q$ covariates is specified as
\begin{equation}\label{circlinlink}
Y = \gamma_0 + \gamma_{cos}*\cos(\Theta)*R + \gamma_{sin}*\sin(\Theta)*R + \gamma_1*x_1 + \dots + \gamma_q*x_q +  \epsilon,
\end{equation}
\noindent where the random variable \(R\geq0\) will be introduced below, the
error term \(\epsilon \sim N(0, \sigma^2)\) with variance \(\sigma^2>0\),
\(\gamma_0, \gamma_{cos}, \gamma_{sin}, \gamma_1, \dots, \gamma_q\) are the
intercept and regression coefficients and \(x_1, \dots, x_q\) are the \(q\)
covariates. In both of these models the
conditional distribution of \(Y\) given \(\Theta=\theta\) and \(R = r\) is given
by
\begin{equation}\label{ycondtheta}
f(y \mid \theta, r) = \frac{1}{\sqrt{2\pi\sigma^2}}\exp\left[-\frac{(y - (\gamma_0 + \gamma_1x_1 + \dots + \gamma_qx_q+c))^{2}}{2\sigma^2}\right],\nonumber
\end{equation}
\noindent where \(c = \begin{bmatrix} r \cos(\theta) \\ r\sin(\theta)
\end{bmatrix}^t \begin{bmatrix} \gamma_{cos} \\ \gamma_{sin} \end{bmatrix}\),
\(r \geq 0\). The linear outcome thus has a normal distribution conditional on
\(\Theta\) and \(R\) and contains already linear covariates \(x_1, \dots, x_q\)
in its location part.\newline
\indent For the circular outcome we assume either a projected normal (PN) or a general
projected normal (GPN) distribution. These distributions arise from the radial
projection of a distribution defined on the plane onto the circle. The relation
between a bivariate vector \(\boldsymbol{S}\) in the plane and the circular
outcome \(\Theta\) is defined as follows
\begin{equation}\label{projection}
\boldsymbol{S} = \begin{bmatrix} S^{I} \\ S^{II} \end{bmatrix} = R\boldsymbol{u} = \begin{bmatrix} R \cos (\Theta) \\  R\sin (\Theta) \end{bmatrix},
\end{equation}
\noindent where \(R = \mid\mid \boldsymbol{S} \mid\mid\), the Euclidean norm of
the bivariate vector \(\boldsymbol{S}\). In the PN distribution we assume
\(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{I})\) and in the GPN we
assume \(\boldsymbol{S} \sim N_2(\boldsymbol{\mu}, \boldsymbol{\Sigma})\) where
$\boldsymbol{\mu} \in \mathbb{R}^2$, \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2 & \xi\\ \xi & 1
\end{bmatrix}\), and \(\xi,\tau \in (-\infty, +\infty)\) (as in
@hernandez2016general). This leads to the
circular-linear PN (CL-PN) and circular-linear GPN (CL-GPN) distributions. We
will now detail how we modify both cylindrical distributions to also incorporate
covariates for the circular part.

\subsubsection{The modified CL-PN distribution}

Following @nunez2011bayesian, the joint density of \(\Theta\) and \(R\) for the
PN distribution equals
\begin{equation}\label{pnreg}
f(\theta,r \mid \boldsymbol{\mu}, \boldsymbol{I}) = \frac{r}{2\pi} \exp\left[- \frac{(r\boldsymbol{u} - \boldsymbol{\mu})^t(r\boldsymbol{u} - \boldsymbol{\mu})}{2}\right],
\end{equation}
\noindent where \(\boldsymbol{u}= \begin{bmatrix} \cos (\theta) \\ \sin (\theta)
\end{bmatrix}\) and $r$ is the same as in \eqref{circlinlink} and
\eqref{ycondtheta} and is defined in \eqref{projection}. In a regression setup
the outcomes \(\theta_i,r_i\) for each individual \(i = 1, \dots, n\), where
\(n\) is the sample size, are generated independently from the distribution with
density \eqref{pnreg}. The mean vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\)
is then defined as \(\boldsymbol{\mu}_i = \boldsymbol{B}^t\boldsymbol{z}_i\)
where the vector \(\boldsymbol{z}_i\) is a vector of dimension $p + 1$ that
contains the covariate values and the value 1 to estimate an intercept and
\(\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})\) contains
the regression coefficients and intercepts.

\subsubsection{The modified CL-GPN distribution}

Following @wang2012directional and @hernandez2016general the joint density of
\(R\) and \(\Theta\) for the GPN distribution equals
\begin{equation}\label{gpnreg}
f(\theta, r \mid \boldsymbol{\mu}, \boldsymbol{\Sigma}) = \frac{r}{2\pi\tau} \exp\left[ -\frac{(r\boldsymbol{u}-\boldsymbol{\mu})^{t}\boldsymbol{\Sigma}^{-1}(r\boldsymbol{u}-\boldsymbol{\mu})}{2}\right],
\end{equation}
\noindent where we recall that \(\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 +
\xi^2 & \xi\\ \xi & 1 \end{bmatrix}\). In a regression setup the outcomes
\(\theta_i\) and \(r_i\) for each individual are generated independently from
\eqref{gpnreg}. The mean vector \(\boldsymbol{\mu}_i \in \mathbb{R}^2\) is
defined in the same way via covariates as for the modified CL-PN distribution.

\subsubsection{Parameter estimation}

Both cylindrical models introduced here are estimated using Markov Chain Monte
Carlo (MCMC) methods based on @nunez2011bayesian, @wang2012directional and
@hernandez2016general for the regression of the circular outcome. A detailed
description of the Bayesian estimation and MCMC samplers can be found in the
Supplementary Material.

\subsection{The modified Abe-Ley model}\label{WeiSSVM}

This model is an extension of the cylindrical model introduced in
@abe2017tractable to the regression context. The joint density of \(\Theta\) and
\(Y\), in this model defined only on the positive real half-line \([0, +
\infty)\), reads
\begin{equation}\label{WeiSSVMdensity}
f(\theta, y) = \frac{\alpha\nu^\alpha}{2\pi\cosh(\kappa)}
                 (1 +\lambda\sin(\theta - \mu))
                 y^{\alpha-1}
                 \exp[-(\nu y)^{\alpha}(1-\tanh(\kappa)\cos(\theta - \mu))],
\end{equation}
\noindent where \(\alpha > 0\) is a linear shape parameter, \(\kappa > 0\) and
\(\lambda \in [-1, 1]\) are circular concentration and skewness parameters with
\(\kappa\) also regulating the circular-linear dependence. Our modification
occurs at the level of the linear scale parameter \(\nu>0\) and circular
location parameter \(\mu\in [0, 2\pi)\), both of which we express in terms of
covariates: \(\nu_i = \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma}) > 0\) and
\(\mu_i = \beta_0 + 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta})\). The
parameter \(\boldsymbol{\gamma}\) is a vector of \(q\) regression coefficients
\(\gamma_j \in (-\infty, +\infty)\) for the prediction of \(y\) where \(j = 0,
\dots, q\) and \(\nu_0\) is the intercept. The parameter \(\beta_0 \in [0,
2\pi)\) is the intercept and \(\boldsymbol{\beta}\) is a vector of \(p\)
regression coefficients \(\beta_j \in (-\infty, +\infty)\) for the prediction of
\(\theta\) where \(j = 1, \dots, p\). The vector \(\boldsymbol{x}_i\) is a
vector of predictor values for the prediction of \(y\) and \(\boldsymbol{z}_i\)
is a vector of predictor values for the prediction of \(\theta\). In a
regression setup the outcome vector \((\theta_i, y_i)^t\) for each individual is
generated independently from the modified
density \eqref{WeiSSVMdensity}.\newline
\indent As in @abe2017tractable, the conditional distribution of \(Y\) given
\(\Theta=\theta\) is a Weibull distribution with shape \(\alpha\) and scale
\(\nu(1-\tanh(\kappa)\cos(\theta - \mu))^{1/\alpha}\) and the conditional
distribution of \(\Theta\) given \(Y=y\) is a sine skewed von Mises distribution
with location parameter \(\mu\) and concentration parameter \((\nu
y)^\alpha\tanh(\kappa)\). The log-likelihood for this model equals
\begin{align}\label{WeiSSVMLikelihood}
l(\alpha, \boldsymbol{\gamma}, \lambda, \kappa, \boldsymbol{\beta}) 
   &= n[\ln(\alpha) - \ln(2\pi\cosh(\kappa))] + \alpha \sum^{n}_{i = 1} \boldsymbol{x}_i^t\boldsymbol{\gamma} \nonumber\\
   &\:\:\:\:+\sum^{n}_{i = 1} \ln(1 +\lambda\sin(\theta_i - \beta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta}))) 
   +(\alpha-1)\sum^{n}_{i = 1} \ln(y_i) \nonumber\\
   &\:\:\:\:-\sum^{n}_{i = 1}( \exp(\boldsymbol{x}_i^t\boldsymbol{\gamma})y_i)^{\alpha}(1-\tanh(\kappa)\cos(\theta_i - \beta_0 - 2\tan^{-1}(\boldsymbol{z}_i^t\boldsymbol{\beta}))).\nonumber
\end{align}
\noindent We can use numerical optimization (Nelder-Mead) to find solutions for
the maximum likelihood (ML) estimates for the parameters of the model.

\subsection{Modified joint projected and skew normal (GPN-SSN)}\label{CL-GPN_multivariate}

This model is an extension of the cylindrical model introduced by
@mastrantonio2018joint to the regression context. Both models contain \(m\) independent circular
  outcomes and \(w\) independent linear outcomes. The circular outcomes
  \(\boldsymbol{\Theta} = (\boldsymbol{\Theta}_1, \dots,
  \boldsymbol{\Theta}_m)\) are modelled together by a multivariate GPN
  distribution. The joint distribution of $\boldsymbol{\Theta}$ and $\boldsymbol{R}$ can thus be modeled as the product of (\ref{gpnreg}) for each of the $m$ circular outcomes. The linear outcomes \(\boldsymbol{Y} = (\boldsymbol{Y}_1,
  \dots, \boldsymbol{Y}_w)\) are modelled together by a multivariate skew normal
  distribution [@sahu2003new]. Because the GPN distribution is modelled using a
  so-called augmented representation (as in \eqref{projection} and
  \eqref{gpnreg}) it is convenient to use a similar tactic for modelling the
  multivariate skew normal distribution. Following @mastrantonio2018joint the
  linear outcomes are represented as
\[\boldsymbol{Y} = \boldsymbol{M}_y + \boldsymbol{\Lambda}\boldsymbol{D} + \boldsymbol{H},\]
\noindent where \(\boldsymbol{M}_y\) is a mean vector for the linear outcome
\(\boldsymbol{Y}\), \(\boldsymbol{\Lambda} = \text{diag}(\boldsymbol{\lambda})\)
is a \(w \times w\) diagonal matrix with diagonal elements
\(\lambda_1, \dots, \lambda_w\) (skewness parameters),
\(\boldsymbol{D} \sim HN_w(\boldsymbol{0}_w, \boldsymbol{I}_w)\), a
$w$-dimensional half normal distribution [@olmos2012extension], and
\(\boldsymbol{H} \sim N_w(\boldsymbol{0}_w, \boldsymbol{\Sigma}_y)\). This means
that, conditional on the auxiliary data \(\boldsymbol{D}\), \(\boldsymbol{Y}\)
is normally distributed with mean \(\boldsymbol{M}_y +
\boldsymbol{\Lambda}\boldsymbol{D}\) and covariance matrix
\(\boldsymbol{\Sigma}_y\). The joint density for \((\boldsymbol{Y}^t,
\boldsymbol{D}^t)^t\) is defined as:
\begin{equation}\label{YDjoint}
f(\boldsymbol{y}, \boldsymbol{d}) = 2^w\phi_w(\boldsymbol{y} \mid \boldsymbol{M}_y + \boldsymbol{\Lambda}\boldsymbol{d}, \boldsymbol{\Sigma}_y) \phi_w(\boldsymbol{d} \mid \boldsymbol{0}_w, \boldsymbol{I}_w),\nonumber
\end{equation}
\noindent where
\(\phi_\ell(\cdot|\boldsymbol{M}_\ell,\boldsymbol{\Sigma}_\ell)\) stands for
the \(\ell\)-dimensional normal density with mean vector
\(\boldsymbol{M}_\ell\) and
covariance \(\boldsymbol{\Sigma}_\ell\). As in @mastrantonio2018joint dependence between the linear and circular outcome is created by
modelling the augmented representations of \(\boldsymbol{\Theta}\) and
\(\boldsymbol{Y}\) together in a \(2m + w\) dimensional normal
distribution. The joint density of the model is then represented by:
\begin{equation}\label{YDThetarjoint} 
f(\boldsymbol{\theta}, \boldsymbol{r},
\boldsymbol{y}, \boldsymbol{d}) = 2^w\phi_{2m+w}((\boldsymbol{s}^t,
\boldsymbol{y}^t)^t \mid \boldsymbol{M} + (\boldsymbol{0}_{2m}^t, ({\rm
diag}(\boldsymbol{\lambda})\boldsymbol{d})^t)^t, \boldsymbol{\Sigma})
\phi_w(\boldsymbol{d} \mid \boldsymbol{0}_w, \boldsymbol{I}_w) \prod_{j =
1}^{m}r_j, 
\end{equation}
\noindent where \(\boldsymbol{s} = (r_1(\cos(\theta_1), \sin(\theta_1)), \dots,
r_m(\cos(\theta_m), \sin(\theta_m)))^t\), the mean vector \(\boldsymbol{M} =
(\boldsymbol{M}_s^t, \boldsymbol{M}_y^t)^t\) and \(\boldsymbol{\Sigma} =
\left ( \begin{matrix} \boldsymbol{\Sigma}_s & \boldsymbol{\Sigma}_{sy} \\
\boldsymbol{\Sigma}_{sy}^t & \boldsymbol{\Sigma}_y \\ \end{matrix} \right )\).
The matrix \(\boldsymbol{\Sigma}_s\) is the covariance matrix for the variances
of and covariances between the augmented representations of the circular outcome
and the matrix \(\boldsymbol{\Sigma}_{sy}\) contains covariances between the
augmented representations of the circular outcome and the linear outcome.
\newline
\indent In our regression extension we have \(i = 1, \dots, n\) observations of
\(m\) circular outcomes, \(w\) linear outcomes and \(g\) covariates. The mean in
the density in \eqref{YDThetarjoint} then becomes \(\boldsymbol{M}_i =
\boldsymbol{B}^t\boldsymbol{x}_i\) where \(\boldsymbol{B}\) is a \((g + 1)
\times (2m + w)\) matrix with regression coefficients and intercepts and $\boldsymbol{x}_i$
is a $g + 1$ dimensional vector containing the value 1 to estimate an intercept
and the $g$ covariate values. \newline
\indent We estimate the model using MCMC methods. A detailed description of these
methods is given in the Supplementary Material.

\section{Model fit}

\indent We use the loglikelihoods of the following conditional densities for the
computation of the PLSL in the teacher data:

\begin{itemize}
\item For the modified CL-PN model:

$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{pnreg}.

\item For the modified CL-GPN model:

$y_i \mid \mu_i, \sigma^2 \sim N(\mu_i, \sigma^2)$, where $\mu_i = \hat{y}_i$ and for $\theta_i$ we use \eqref{gpnreg}.

\item For the modified Abe-Ley model:

$y_i \mid \theta_i, \nu_i, \mu_i, \kappa, \alpha \sim W\left(\nu_i(1-\tanh(\kappa)\cos(\theta_i - \mu_i))^{1/\alpha}, \alpha\right)$, a Weibull distribution.

$\theta_i \mid y_i, \nu_i, \mu_i, \kappa, \alpha, \lambda \sim SSVM\left(\mu_i, (\nu_iy_i)^{\alpha}(\tanh{\kappa})\right)$, a sine-skewed von Mises distribution.

\item For the modified joint projected and skew normal model:

$y_i \mid \boldsymbol{M}_i, \boldsymbol{\Sigma}, \theta_i, r_i \sim SSN(M_{i_y} + \lambda d_i + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_s^{-1}(\boldsymbol{s}_i - \boldsymbol{M}_{i_s}), \sigma^2_y + \boldsymbol{\Sigma}_{sy}^t\boldsymbol{\Sigma}_s^{-1}\boldsymbol{\Sigma}_{sy}),$

$\theta_i \mid \boldsymbol{M}_i, \boldsymbol{\Sigma}, y_i, d_i \sim GPN(\boldsymbol{M}_{i_s} + \boldsymbol{\Sigma}_{sy}\sigma^{-2}_y(y_i - M_{i_y} - \lambda d_i), \boldsymbol{\Sigma}_s + \boldsymbol{\Sigma}_{sy}\sigma_y^{-2}\boldsymbol{\Sigma}_{sy}^t)$

where $SSN$ is the skew normal distribution.
\end{itemize}

\section{MCMC procedures}
\subsection{Bayesian Model and MCMC procedure for the modified CL-PN model}\label{A1}

We use the following algorithm to obtain posterior estimates from the model:

\begin{enumerate}
\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_1, \dots, \theta_n$ and the linear outcome $\boldsymbol{y} = y_1, \dots, y_n$ where $n$ is the sample size, and the design matrices $\boldsymbol{Z}^k_{n \times 2}$ (for $k \in \{I,II\}$) and $\boldsymbol{X}_{n \times 4}$ of the circular and the linear outcome respectively, in a training (90\%) and holdout (10\%) set. 
\item Define the prior parameters for the training set. In this paper we use:

\begin{itemize}
\item Prior for $\boldsymbol{\gamma}$: $N_4(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_{0})$, with  $\boldsymbol{\mu}_{0} = (0,0,0,0)^t$ and  $\boldsymbol{\Lambda}_{0} = 10^{-4}\boldsymbol{I}_4$.
\item Prior for $\sigma^2$: $IG(\alpha_{0}, \beta_{0})$, an inverse gamma prior with $\alpha_{0} = 0.001$ and  $\beta_{0} = 0.001$.
\item Prior for $\boldsymbol{\beta^{k}}$: $N_2(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_{0})$, with $\boldsymbol{\mu}_{0} = (0,0)^t$ and  $\boldsymbol{\Lambda}_{0} = 10^{-4}\boldsymbol{I}_2$ for $k \in \{I,II\}$.
\end{itemize}

\item Set starting values $\boldsymbol{\gamma} = (0,0,0,0)^t$, $\sigma^2 = 1$ and $\boldsymbol{\beta^{k}} = (0,0)^t$ for $k \in \{I,II\}$. Also set starting values $r_i = 1$ in the training and holdout set. 
\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:
$$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos (\theta_i)\\  r_i\sin (\theta_i)\end{bmatrix}.$$
\item Sample $\boldsymbol{\gamma}$, $\sigma^2$ and $\boldsymbol{\beta^{k}}$ for $k \in \{I,II\}$ for the training dataset from their conditional posteriors:

\begin{itemize}
\item Posterior for $\boldsymbol{\gamma}$: $N_4(\boldsymbol{\mu}_n, \sigma^2\boldsymbol{\Lambda}^{-1}_n)$, with $\boldsymbol{\mu}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)^{-1}(\boldsymbol{\Lambda}_0\boldsymbol{\mu}_0 + \boldsymbol{X}^t\boldsymbol{y})$ and $\boldsymbol{\Lambda}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)$.
\item Posterior for $\sigma^2$: $IG(\alpha_{n}, \beta_{n})$, an inverse gamma posterior with $\alpha_{n} = \alpha_0 + n/2$ and $\beta_{n} = \beta_0 + \frac{1}{2}(\boldsymbol{y}^t\boldsymbol{y} + \boldsymbol{\mu}_{0}^t\boldsymbol{\Lambda}_0\boldsymbol{\mu}_{0} + \boldsymbol{\mu}_{n}^t\boldsymbol{\Lambda}_n\boldsymbol{\mu}_{n})$.
\item Posterior for $\boldsymbol{\beta^{k}}$: $N_2(\boldsymbol{\mu}_n, \boldsymbol{\Lambda}_n)$, with $\boldsymbol{\mu}_n = ((\boldsymbol{Z}^k)^t\boldsymbol{Z}^k + \boldsymbol{\Lambda}_0)^{-1}(\boldsymbol{\Lambda}_0\boldsymbol{\mu}_0 + (\boldsymbol{Z}^k)^t\boldsymbol{s}^k)$ and $\boldsymbol{\Lambda}_n = ((\boldsymbol{Z}^k)^t\boldsymbol{Z}^k + \boldsymbol{\Lambda}_0)$.
\end{itemize}

\item Sample new $r_i$ for the training and holdout dataset from the following posterior:
$$f(r_i \mid \theta_i, \boldsymbol{\mu}_i) \propto r_i \exp{\left(-\frac{1}{2}(r_i)^2 + b_ir_i\right)}$$ 
where $b_i = \begin{bmatrix} \cos (\theta_i) \\ \sin (\theta_i)\end{bmatrix}^t\boldsymbol{\mu}_i$, $\boldsymbol{\mu}_i = \boldsymbol{B}^t\boldsymbol{z_i}$ and $\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$. 
\noindent We can sample from this posterior using a slice sampling technique (Cremers et al., 2018):

\begin{itemize}
\item In a slice sampler the joint density for an auxiliary variable $v_{i}$ with $r_{i}$ is
$$p(r_{i}, v_{i}\mid \theta_{i}, \boldsymbol{\mu}_{i}=\boldsymbol{B}^t\boldsymbol{z}_{i}) \propto r_{i} \textbf{I}\left(0 < v_i < \exp\left\{ -\frac{1}{2}(r_{i} - b_{i})^2\right\}\right)\textbf{I}(r_i > 0).$$
\noindent The full conditional for $v_{i}$, $p(v_{i} \mid r_{i},\boldsymbol{\mu}_{i}, \theta_{i})$, is
$$U\left(0, \exp\left\{-\frac{1}{2}(r_{i} -  b _{i})^2\right\}\right)$$
and the full conditional for $r_i$, $p(r_{i} \mid v_{i},\boldsymbol{\mu}_{i}, \theta_{i})$, is proportional to
$$r_{i} \textbf{I}\left(b_{i} + \max\left\{-b_{i}, -\sqrt{-2\ln v_{i}}\right\} < r_{i} < b_{i} + \sqrt{-2\ln v_{i}}\right).$$
\noindent We thus sample $v_{i}$ from the uniform distribution specified above. Independently we sample a value $m$ from $U(0,1)$. We obtain a new value for $r_{i}$ by computing $ r_{i} = \sqrt{(r_{i_{2}}^{2}-r_{i_{1}}^{2})m + r_{i_{1}}^{2}}$ where $r_{i_{1}}=b_{i} +\max\left\{-b_{i}, -\sqrt{-2\ln v_{i}}\right\}$ and $ r_{i_{2}}= b_{i} + \sqrt{-2\ln v_{i}}$.
\end{itemize}
\item Compute the PLSL for the circular and linear outcome on the holdout set using the estimates of $\boldsymbol{\gamma}$, $\sigma^2$ and $\boldsymbol{\beta^{k}}$ for $k \in \{I,II\}$ for the training dataset.
\item Repeat steps 4 to 7 until the sampled parameter estimates have converged. We assess convergence visually using traceplots.
\end{enumerate}

\newpage
\subsection{Bayesian Model and MCMC procedure for the modified CL-GPN model}\label{A2}

We use the following algorithm to obtain posterior estimates from the model:

\begin{enumerate}
\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_1, \dots, \theta_n$ and the linear outcome $\boldsymbol{y} = y_1, \dots, y_n$ where $n$ is the sample size, and the design matrices $\boldsymbol{Z}_{n \times 2}$  and $\boldsymbol{X}_{n \times 4}$ of the circular and the linear outcome respectively, in a training (90\%) and holdout (10\%) set. 
\item Define the prior parameters for the training set. In this paper we use:

\begin{itemize}
\item Prior for $\boldsymbol{\gamma}$: $N_4(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_{0})$, with $\boldsymbol{\mu}_{0} = (0,0,0,0)^t$ and $\boldsymbol{\Lambda}_{0} = 10^{-4}\boldsymbol{I}_4$.
\item Prior for $\sigma^2$: $IG(\alpha_{0}, \beta_{0})$, an inverse gamma prior with $\alpha_{0} = 0.001$ and $\beta_{0} = 0.001$.
\item Prior for $\boldsymbol{\beta}_{j}$: $N_2(\boldsymbol{\mu}_{0}, \boldsymbol{\Lambda}_0)$, with $\boldsymbol{\mu}_{0} = (0,0)^t$ and  $\boldsymbol{\Sigma}_{0} = 10^{5}\boldsymbol{I}_2$ for $j \in \{0, \dots, p\}$ where $p$ is the number of covariates, 1, in $\boldsymbol{Z}$.
\item Prior for $\xi$: $N(\mu_0, \sigma^2)$, with $\mu_0 = 0$ and $\sigma^2 = 10^{4}$.
\item Prior for $\tau$: $IG(\alpha_{0}, \beta_{0})$, an inverse gamma prior with $\alpha_{0} = 0.01$ and $\beta_{0} = 0.01$.
\end{itemize}

\item Set starting values $\boldsymbol{\gamma} = (0,0,0,0)^t$, $\sigma^2 = 1$, $\boldsymbol{\beta}_j = (0,0)^t$ for $j \in \{0,1\}$, $\xi = 0$, $\tau = 1$ and $\boldsymbol{\Sigma} = \begin{bmatrix} \tau^2 + \xi^2 & \xi\\ \xi & 1 \end{bmatrix}$. Also set starting values $r_i = 1$ in the training and holdout set. 
\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:
$$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos (\theta_i) \\  r_i\sin (\theta_i)\end{bmatrix}.$$
\item Sample $\boldsymbol{\gamma}$, $\sigma^2$, $\boldsymbol{\beta}_j$ for $j \in \{0,1\}$, $\xi$ and $\tau$ for the training dataset from their conditional posteriors:

\begin{itemize}
\item Posterior for $\boldsymbol{\gamma}$: $N_4(\boldsymbol{\mu}_n, \sigma^2\boldsymbol{\Lambda}^{-1}_n)$, with $\boldsymbol{\mu}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)^{-1}(\boldsymbol{\Lambda}_0\boldsymbol{\mu}_0 + \boldsymbol{X}^t\boldsymbol{y})$ and $\boldsymbol{\Lambda}_n = (\boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\Lambda}_0)$.
\item Posterior for $\sigma^2$: $IG(\alpha_{n}, \beta_{n})$, an inverse gamma posterior where $\alpha_{n} = \alpha_0 + n/2$ and $\beta_{n} = \beta_0 + \frac{1}{2}(\boldsymbol{y}^t\boldsymbol{y} + \boldsymbol{\mu}_{0}^t\boldsymbol{\Lambda}_0\boldsymbol{\mu}_{0} + \boldsymbol{\mu}_{n}^t\boldsymbol{\Lambda}_n\boldsymbol{\mu}_{n})$.
\item Posterior for $\boldsymbol{\beta}_j$: $N_2(\boldsymbol{\mu}_{j_{n}}, \boldsymbol{\Sigma}_{j_{n}})$, with $\boldsymbol{\mu}_{j_{n}} = \boldsymbol{\Sigma}_{j_{n}}\boldsymbol{\Sigma}^{-1}\Bigg(-\sum_{i=1}^{n}z_{i,j-1}\sum_{l\neq j}z_{i,l-1}\boldsymbol{\beta}_l + \sum_{i=1}^{n}z_{i,j-1}r_i\begin{bmatrix} \cos (\theta_i) \\ \sin (\theta_i)\end{bmatrix}\Bigg)$ and  $\boldsymbol{\Sigma}_{j_{n}} = \Big(\sum_{i=1}^{n}z_{i,j-1}^2\boldsymbol{\Sigma}^{-1}+\boldsymbol{\Lambda}_0\Big)^{-1}$ for $j \in \{0, \dots, p\}$ where $p$ is the number of covariates, 1, in $\boldsymbol{Z}$.
\item Posterior for $\xi$: $N(\mu_n, \sigma^2_n)$, with $\mu_n = \frac{\tau^{-2} \sum_{i=1}^{n}(s^{I}_{i} - \mu_i^{I})(s^{II}_{i} - \mu_i^{II}) + \mu_0\sigma_0^{-2}}{\tau^{-2}\sum_{i=1}^{n}(s^{II}_{i} - \mu_i^{II})^2 + \sigma_0^{-2}}$ and $\sigma_n^2 = \frac{1}{\tau^{-2}\sum_{i=1}^{n}(s^{II}_{i} - \mu_i^{II})^2 + \sigma_0^{-2}}$ where $\mu_i^{I} = (\boldsymbol{\beta}^{I})^t\boldsymbol{z}_i$ and $\mu_i^{II} = (\boldsymbol{\beta}^{II})^t\boldsymbol{z}_i$. 
\item Posterior for $\tau$: $IG(\alpha_n, \beta_n)$, an inverse gamma posterior with $\alpha_n = \frac{n}{2} + \alpha_0$ and $\beta_n = \sum\limits_{i = 1}^{n}(s^{I}_{i} - \{\mu_i^{I} + \xi(s^{II}_{i} - \mu_i^{II})\})^2 + \beta_0$
\end{itemize}

\item Sample new $r_i$ for the training and holdout dataset from the following posterior:
$$f(r_i \mid \theta_i, \boldsymbol{\mu}_i) \propto r_i \exp{\left\{-\frac{1}{2}A_i\bigg(r_i-\frac{B_i}{A_i}\bigg)^2\right\}}$$ 
where $B_i = \begin{bmatrix} \cos (\theta_i) \\ \sin (\theta_i)\end{bmatrix}^t\boldsymbol{\Sigma}^{-1}\boldsymbol{\mu}_i$, $\boldsymbol{\mu}_i = \boldsymbol{B}^t\boldsymbol{z_i}$, $\boldsymbol{B} = (\boldsymbol{\beta}^{I}, \boldsymbol{\beta}^{II})$ and $A_i = \begin{bmatrix} \cos (\theta_i) \\ \sin (\theta_i)\end{bmatrix}^t\boldsymbol{\Sigma}^{-1}\begin{bmatrix} \cos (\theta_i) \\ \sin (\theta_i)\end{bmatrix}$.
\noindent We can sample from this posterior using a slice sampling technique (Hernandez-Stumpfhauser et al. 2018):

\begin{itemize}
\item In a slice sampler the joint density for an auxiliary variable $v_{i}$ with $r_{i}$ is
$$p(r_{i}, v_{i}\mid \theta_{i}, \boldsymbol{\mu}_{i}=\boldsymbol{B}^t\boldsymbol{z}_{i}) \propto r_{i} \textbf{I}\bigg(0 < v_i < \exp\left\{ -\frac{1}{2} A_i\left(r_{i} - \frac{B_i}{A_i}\right)^2\right\}\bigg)\textbf{I}(r_i > 0).$$
\item The full conditional for $v_{i}$, $p(v_{i} \mid r_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$, is
$$U\Bigg(0, \exp\left\{-\frac{1}{2}A_i\bigg(r_i -  \frac{B_{i}}{A_i}\bigg)^2\right\}\Bigg)$$
and the full conditional for $r_i$, $p(r_{i} \mid v_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$, is proportional to
$$r_{i} \textbf{I}\left(\frac{B_i}{A_i} + \max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\} < r_{i} < \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}\right).$$
\item We thus sample $v_{i}$ from the uniform distribution specified above. Independently we sample a value $m$ from $U(0,1)$. We obtain a new value for $r_{i}$ by computing $r_{i} = \sqrt{(r_{i_{2}}^{2}-r_{i_{1}}^{2})m + r_{i_{1}}^{2}}$ where $r_{i_{1}}=\frac{B_i}{A_i} +\max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\}$ and $ r_{i_{2}}= \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}$.
\end{itemize}

\item Compute the PLSL for the circular and linear outcome on the holdout set using the estimates of $\boldsymbol{\gamma}$, $\sigma^2$, $\boldsymbol{\beta^{k}}$ for $k \in \{I,II\}$, $\xi$ and $\tau$ for the training dataset. 
\item Repeat steps 4 to 7 until the sampled parameter estimates have converged. We visually assess convergence using traceplots.
\end{enumerate}

\newpage
\subsection{Bayesian Model and MCMC procedure for the modified GPN-SSN model}\label{A3}

\begin{enumerate}
\item Split the data, with the circular outcome $\boldsymbol{\theta} = \theta_1, \dots, \theta_n$ and the linear outcome $\boldsymbol{y} = y_1, \dots, y_n$ where $n$ is the sample size, and the design matrix $\boldsymbol{X}_{n \times 2}$ in a training (90\%) and holdout (10\%) set. Note that in this paper we have only one circular outcome and one linear outcome and the MCMC procedure outlined here is specified for this situation. It can however be generalized to a situation with multiple circular and linear outcomes without too much effort. 
\item Define the prior parameters for the training set. Since we have only one circular outcome, one linear outcome and one covariate, we have $m = 1$, $w = 1$ and $g = 1$. In this paper we use the following priors:

\begin{itemize}
\item Prior for $\boldsymbol{\Sigma}$: $IW(\boldsymbol{\Psi}_0, \nu_0)$, an inverse Wishart with $\boldsymbol{\Psi}_0 = 10^{-4}\boldsymbol{I}_{2m + w}$ and $\nu_0 = 1$.   
\item Prior for $\boldsymbol{B}$ in vectorized form: $N_{(g + 1)(2m + w)}(\boldsymbol{\beta}_0, \boldsymbol{\Sigma}  \otimes \boldsymbol{\kappa}_0)$, where $\otimes$ stands for the Kronecker product, $\boldsymbol{\beta}_0 = \text{vec}(\boldsymbol{B}_0)$, the matrix with prior values for the regression coefficients. We choose $\boldsymbol{\beta}_0 = \boldsymbol{0}_{(g + 1)(2m + w)}$, $\boldsymbol{B}_0 = \boldsymbol{0}_{(g + 1) \times (2m + w)}$ and $\boldsymbol{\kappa}_0 = 10^{-4}\boldsymbol{I}_{g + 1}$.
\item Prior for $\lambda$: $N(\gamma_0, \omega_0)$, with $\gamma_0 = 0$ and $\omega_0 = 10000$.
\end{itemize}

\item Set starting values $\boldsymbol{\beta} = (0,0,0,0,0,0)^t$, $\boldsymbol{\Sigma} = \boldsymbol{I}_3$ and $\lambda = 0$. Also set starting values $r_i = 1$ and $d_i = 1$ in the training and holdout set. 
\item Compute the latent bivariate outcome $\boldsymbol{s}_i = (s_i^{I}, s_i^{II})^t$ underlying the circular outcome for the holdout and training dataset as follows:
$$\begin{bmatrix} s^{I}_{i} \\ s^{II}_{i} \end{bmatrix} = \begin{bmatrix} r_i \cos (\theta_i) \\  r_i\sin (\theta_i)\end{bmatrix}.$$
\item Compute the latent outcomes $\tilde{y}_i$ underlying the linear outcome for the holdout and training dataset as follows:
$$\tilde{y}_i = \lambda d_i. $$
\item Compute $\boldsymbol{\eta}_i$ defined as follows for each individual $i$:
$$\boldsymbol{\eta}_i = (\boldsymbol{s}_i^t,y_i)^t - (\boldsymbol{0}_{2m}^t, \lambda d_i)^{t}.$$

\item Sample $\boldsymbol{B}$, $\boldsymbol{\Sigma}$ and $\lambda$ for the training dataset from their conditional posteriors:

\begin{itemize}
\item Posterior for $\boldsymbol{\Sigma}$: $IW(\boldsymbol{\Psi}_n, \nu_n)$, an inverse Wishart with $\boldsymbol{\Psi}_n = \boldsymbol{\Psi}_0 + (\boldsymbol{\eta} - \boldsymbol{X}^t\boldsymbol{B})^t(\boldsymbol{\eta} - \boldsymbol{X}^t\boldsymbol{B}) + (\boldsymbol{B} - \boldsymbol{B}_0)^t\boldsymbol{\kappa}_0(\boldsymbol{B} - \boldsymbol{B}_0)$ and $\nu_n = \nu_0 + n$ where $n$ is the sample size.
\item Posterior for $\boldsymbol{B}$ in matrix form: $MN(\boldsymbol{B}_n,   \boldsymbol{\kappa}_n, \boldsymbol{\Sigma})$, with $\boldsymbol{B}_n = \boldsymbol{\kappa}_n^{-1}\boldsymbol{X}^t\boldsymbol{\eta} + \boldsymbol{\kappa}_0\boldsymbol{B}_0$ and $\boldsymbol{\kappa}_n = \boldsymbol{X}^t\boldsymbol{X} + \boldsymbol{\kappa}_0$.
\item Posterior for $\lambda$: $N(\gamma_n, \omega_n)$, with $\omega_n = \big(\sum_{i = 1}^{n}d_i^2\sigma^{-2}_{y|s} + \omega_0^{-1}\big)^{-1}$ and $\gamma_n = \omega_n \big(\sum_{i = 1}^{n}d_i\sigma^{-2}_{y|s}(y_i - \mu_{y_i|s_i}) + \omega_0^{-1}\gamma_0 \big)$ where $\mu_{y_i|s_i} = \mu_y + \boldsymbol{\Sigma}_{sy}^{t}\boldsymbol{\Sigma}_{s}^{-1}(\boldsymbol{s_i - \boldsymbol{\mu}_s)}$ and $\sigma^2_{y|s} = \sigma^2_{y} - \boldsymbol{\Sigma}_{sy}^{t}\boldsymbol{\Sigma}_{s}^{-1}\boldsymbol{\Sigma}_{sy}$.
\end{itemize}

\item Sample new $d_i$ for the training and holdout dataset from the following posterior:
$$f(d_i) \propto \phi(y_i|\mu_{y_i|s_i} + \lambda d_i, \sigma^2_{y|s})\phi(d_i|0, 1),$$
where $\mu_{y_i|s_i} = \boldsymbol{B}_{y_i|s_i}^t\boldsymbol{x}_i$. We can see each $d_i$ as a positive regressor with $\lambda$ as covariate and $\phi(d_i|0, 1)$ as prior (Mastrantonio, 2018). The full conditional is then truncated normal with support $\mathbb{R}^{+}$ as follows:
$$N(m_{d_i}, v),$$ 
\noindent where $v = \big(\lambda^2\sigma^{-2}_{y|s} + 1\big)$ and $m_{d_i} = v\lambda\sigma^{-2}_{y|s}\big(y_i - \mu_{y_i|s_i}\big)$. 
\item Sample new $r_i$ for the training and holdout dataset from the following posterior
$$f(r_i \mid \theta_i, \boldsymbol{\mu}_i) \propto r_i \exp{\left\{-0.5A_i\bigg(r_i-\frac{B_i}{A_i}\bigg)^2\right\}}$$ 
where $B_i = \begin{bmatrix} \cos (\theta_i) \\ \sin (\theta_i)\end{bmatrix}^t\boldsymbol{\Sigma}_{s_i|y_i}^{-1}\boldsymbol{\mu}_{s_i|y_i}$, $\boldsymbol{\mu}_{s_i|y_i} = \boldsymbol{B}_{s_i|y_i}^t\boldsymbol{x}_i$ and $A_i = \begin{bmatrix} \cos (\theta_i) \\ \sin (\theta_i)\end{bmatrix}^t\boldsymbol{\Sigma}_{s_i| y_i}^{-1}\begin{bmatrix} \cos (\theta_i) \\ \sin (\theta_i)\end{bmatrix}$. The parameters $\boldsymbol{\mu}_{s_i|y_i}$ and $\boldsymbol{\Sigma}_{s_i| y_i}$ are the conditional mean and covariance matrix of $\boldsymbol{s}_i$ assuming that $(\boldsymbol{s}_i^t, y_i)^t \sim N_{2m+w}(\boldsymbol{\mu} + (\boldsymbol{0}_{2m}^t, \lambda d_i)^t, \boldsymbol{\Sigma})$. 
Because in this paper $\boldsymbol{\theta}$ originates from a bivariate variable that is known we can in this model (where the variance-covariance matrix of the circular outcome is not constrained in the estimation procedure) simply define the $r_i$ as the Euclidean norm of the bivariate datapoints. However, for didactic purposes we continue with the explanation of the sampling procedure. We
can sample from the posterior for $r_i$ using a slice sampling technique (Hernandez-Stumpfhauser et al. 2018): 
\begin{itemize}
\item In a slice sampler the joint density for an auxiliary variable $v_{i}$ with $r_{i}$ is
$$p(r_{i}, v_{i}\mid \theta_{i}, \boldsymbol{\mu}_{i}=\boldsymbol{B}^{t}\boldsymbol{x}_{i}) \propto r_{i} \textbf{I}\bigg(0 < v_i < \exp\left\{ -\frac{1}{2}A_i\left(r_{i} - \frac{B_i}{A_i}\right)^2\right\}\bigg)\textbf{I}(r_i > 0).$$
\item The full conditional for $v_{i}$, $p(v_{i} \mid r_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$, is
$$U\Bigg(0, \exp\left\{-\frac{1}{2}A_i\bigg(r_i -  \frac{B_{i}}{A_i}\bigg)^2\right\}\Bigg)$$
and the full conditional for $r_i$, $p(r_{i} \mid v_{i},\boldsymbol{\mu}_{i}, \boldsymbol{\Sigma}, \theta_{i})$, is proportional to 
$$r_{i} \textbf{I}\left(\frac{B_i}{A_i} + \max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\} < r_{i} < \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}\right)$$
\item We thus sample $v_{i}$ from the uniform distribution specified above. Independently we sample a value $m$ from $U(0,1)$. We obtain a new value for $r_{i}$ by computing $r_{i} = \sqrt{(r_{i_{2}}^{2}-r_{i_{1}}^{2})m + r_{i_{1}}^{2}}$ where $r_{i_{1}}=\frac{B_i}{A_i} +\max\left\{-\frac{B_i}{A_i}, -\sqrt{\frac{-2\ln v_{i}}{A_i}}\right\}$ and $ r_{i_{2}}= \frac{B_i}{A_i} + \sqrt{\frac{-2\ln v_{i}}{A_i}}$.

\end{itemize}

\item Compute the PLSL for the circular and linear outcome on the holdout set using the estimates of $\boldsymbol{B}$, $\boldsymbol{\Sigma}$ and $\lambda$ for the training dataset.

\item Repeat steps 4 to 10 until the sampled parameter estimates have converged.

\item In the MCMC sampler we have estimated an unconstrained $\boldsymbol{\Sigma}$. However, for identification of the model we need to apply  constraints to both $\boldsymbol{\Sigma}$ and $\boldsymbol{\mu}$. Therefore we need the matrix
$$\boldsymbol{C} = \begin{bmatrix} \boldsymbol{C}_s & \boldsymbol{0}_{2m \times w} \\ \boldsymbol{0}_{2m \times w}^t & \boldsymbol{I}_w \end{bmatrix}$$
where $\boldsymbol{C}_s$ is a $2m \times 2m$ diagonal matrix with every $(2(j-1) + k)^{th}$ entry $> 0$ where $k \in \{1,2\}$ and $j = 1, \dots, m$ (Mastrantonio, 2018). The estimates $\boldsymbol{\Sigma}$ and $\boldsymbol{\mu}$ can then be related to their constrained versions $\tilde{\boldsymbol{\Sigma}}$ and $\tilde{\boldsymbol{\mu}}$ as follows:
$$\boldsymbol{\mu} = \boldsymbol{C}\tilde{\boldsymbol{\mu}}$$
$$\boldsymbol{\Sigma} = \boldsymbol{C}\tilde{\boldsymbol{\Sigma}}\boldsymbol{C}.$$

\end{enumerate}

\section*{References}
<div id="refs"></div>