### 2019-08-12 Manuscript/Titlepage.Rmd
Update manuscript + comments heleen

---
title: "Regression models for Cylindrical data in Psychology"
author: |
  | Jolien Cremers^[Corresponding author: jcre@sund.ku.dk]$^{\:\:1}$, Helena J.M. Pennings$^{2,3}$,  Christophe Ley $^{4}$
  | $^1$Department of Biostatistics, Copenhagen University
  | $^2$TNO
  | $^3$Department of Education, Utrecht University
  | $^4$Department of Applied Mathematics, Computer Science and Statistics, Ghent University
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    latex_engine: pdflatex
    number_sections: yes
  html_document: default
  word_document: default
fontsize: 12pt
geometry: margin=1in
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\section*{Acknowledgements}
We thank Irene Klugkist for insightful remarks and comments on the present
paper. Jolien Cremers acknowledges financial support from a Vidi grant awarded
to I. Klugkist from the Dutch Organization for Scientific Research (NWO
452‐12‐010).